### 2018-07-08 paper/draft/elsarticle-template.tex
cleanup

\documentclass[review, times]{elsarticle}

\usepackage{lineno,hyperref}
\modulolinenumbers[5]

\journal{Journal of \LaTeX\ Templates}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

% personal packages
\usepackage{amsmath,amssymb}
\usepackage{subfig}
\usepackage{tabu}
\usepackage{booktabs}
\usepackage{float}

% personal macros
\newtheorem{thm}{Theorem} \newtheorem{lem}[thm]{Lemma} \newdefinition{rmk}{Remark} \newproof{pf}{Proof}
\newproof{pot}{Proof of Theorem \ref{thm2}}

\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}

\begin{document}

\begin{frontmatter}

\title{A fast spectral method for inelastic collision operator and the heated granular flow\tnoteref{mytitlenote}}
% \tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}

%% Group authors per affiliation:
% \author{Jingwei Hu, Zheng Ma\fnref{myfootnote}}
% \address{Department of Mathematics, Purdue University, 150 N. University Street, West Lafayette, IN, USA}
% \fntext[myfootnote]{}

%% or include affiliations in footnotes:
\author[mymainaddress]{Jingwei Hu}
\ead{jingweihu@purdue.edu}

\author[mymainaddress]{Zheng Ma}
% \cortext[mycorrespondingauthor]{Corresponding author}
\ead{ma531@purdue.edu}

\address[mymainaddress]{150 N. University Street, West Lafayette, IN, USA}
% \address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
In this paper, we proposed a fast spectral algorithm of the inelastic operator, with its application to one of the widely used model of granular gases, the heated Enskog-Boltzmann equation. Comparing to the direct spectral method, our fast algorithm reduces the computational complexity from $O\left(N^6\right)$ to $O\left(MN^4\log(N) \right)$ and the storage from $O(N^6)$ to $O\left( MN^4\right)$, where $N$ is the number of discretization points in velocity dimension and $M \ll N^2$ is the number of numerical quadrature points. We test the numerical accuracy and efficiency in both two dimensional and three dimensional cases, where the famous Haff's cooling law is recovered in the 3D example.
\end{abstract}

% \begin{keyword}
% \texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
% \MSC[2010] 00-01\sep  99-00
% \end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction}

It has been found in the past few decays that the granular gases behave fundamentally different from the usual molecular gases modelled as elastically colliding spheres. The rich phenomenology of such systems, such as the formation of clusters and shear instability, draws a lot of attention from both theoretical and industrial application point of view. Different from their molecular counterparts, granular gases allow inelastic collision, in other words, break the time-reversible symmetry because of the energy dissipation. Despite this dissipative nature and its resulting nontrivial properties, the basic equation of kinetic theory, the Boltzmann equation, can be still extended to describe the granular gases with a different collision operator, namely,
\begin{equation}
\partial_t f + v \cdot \nabla_x f = Q_{\text{in}}(f,f),
\label{boltz}
\end{equation}
where $f(t,x,v)$ is the one-particle distribution function depending on the time $t \geq 0$, position $x \in \mathbb{R}^d$ and velocity $v \in \mathbb{R}^d$ with the dimension $d \geq 1$, and $Q_{\text{in}}$ is the so-called inelastic (or granular) collision operator, whose exact expression is presented in later discussions. For convenience, we use $Q$ instead of $Q_{\text{in}}$ in the rest of this paper, whose notation should not be confused with the usual elastic collision operator. Another widely used model, first introduced by van Noije and Ernst \cite{noije}, is the spatial homogeneous inelastic Boltzmann equation based on Enskog-Boltzmann model with a heat source:
\begin{equation} \label{inBoltz}
\partial_tf-\varepsilon \Delta_vf=Q(f,f),
\end{equation}
where the distribution function $f$ depends only on the time $t$ and velocity $v$, and the term $\varepsilon \Delta_vf$ represents the diffusion effects with the diffusion coefficient $\epsilon \ll 1$, incurred by a heat bath of infinite temperature.

The numerical difficulty and cost, of course, lie in the computation of the collision operator. At each time step, the construction of the inelastic collision operator requires $O(N^{2d})$ operations and $O(N^{2d})$ storage in a direct numerical scheme. As is pointed out in \cite{filbet2005}, although the loss term in the inelastic collision can be evaluated only in $O(N^d \log N)$ operations thanks to its convolution structure, the cost of the gain part is still rather formidable. A natural question of course remains to reduce the computational cost for the entire collision operator, as well as the numerical storage -- in order words, to fully exploit the structure of the granular collision operator. In this paper, we propose a fast spectral algorithm for the inelastic collision operator, inspired by a sequence of studies on elastic Boltzmann operator \cite{fast1999, mouhot2006fast, hu2012fast, gamba_hu}. To be specific, in contrast to a direct spectral solver, in 3D this algorithm reduces the computational cost from $O(N^{6})$ to $O\left(MN^{4}\log(N)\right)$ and the storage from $O(N^{6})$ to $O(MN^{4})$, where $M \ll N^{2}$ is the number of quadrature points on $\mathbb{S}^{2}$.

%Dynamics of granular gases has received a lot of attention in the past few decays due to the distinct phenomenology displayed by such systems, with rich applications from the industrial point of view. Different from molecular gases, modelled as elastically colliding spheres, granular gases allow inelastic collision, which in turn implies energy dissipation. As was first pointed out by Goldhirsch and Zanetti in 1993, this dissipative nature is responsible for the nontrivial behaviours, such as formation of clusters and shear instability, comparing to their molecular counterparts. Despite their differences, both gases Fortunately, the basic equation of kinetic theory, the Boltzmann equation, can be easily extended to describe the granular gases.

The rest of the paper is organized as follows: Section 2 provides a brief overview of the inelastic collision operator, and the inelastic Boltzmann equation widely used in practice for the study of granular flow. The numerical algorithms of the inelastic collision operator is present in Section 3. Starting from a naive trial and discussions of its limitations, we henceforth propose a fast spectral method of the operator taking full advantage of its convolution structure in both the two dimensional and three dimensional cases. Finally in Section 4, a number of numerical tests are performed to test both accuracy and efficiency of our fast algorithm in 2D and 3D cases. Also, as can be in the 3D numerical example, the Haff's cooling law is verified.

\section{Inelastic collision operator and its Enskog-Boltzmann equation}
\subsection{Inelastic collision }
To describe the inelastic binary collision, a reconstitution coefficient $e$ is introduced. Specifically, assuming two particles with velocities $v$ and $v_*$ are going to collide, after the collision, the velocities denoted by $v'$ and $v_*'$ are given by the so-called $\omega$-representation \cite{Villani2006} 
\begin{align}\label{omega}
\left\{
\begin{array}{l}
\displaystyle v'=v-\frac{1+e}{2}[(v-v_*)\cdot \omega ]\omega, \\[8pt]
\displaystyle v_*'=v_*+\frac{1+e}{2}[(v-v_*)\cdot \omega]\omega,
\end{array}\right.
\end{align}
where $\omega \in S^{d-1}$ is the impact direction, and $0 \leq e \leq 1$ is the restitution coefficient (with $e=1$ recovering the elastic case). It follows that
\begin{equation}
(v'-v_*')\cdot \omega=-e [(v-v_*)\cdot \omega].
\end{equation}
Furthermore, instead of the conservation of both momentum and energy in the usual molecular gases, here one has the conservation of momentum and the loss of energy:
\begin{equation} \label{lossenergy}
v'+v_*'=v+v_*; \quad v'^2+v_*'^2=v^2+v_*^2-\frac{1-e^2}{2}[(v-v_*)\cdot \omega]^2.
\end{equation}
Now we are ready to define the inelastic collision operator in its weak form
\begin{equation} \label{col_omega}
\int Q(f,f)(v)\phi(v)\diff{v}= \frac{1}{2} \int_{\mathbb{R}^d} \int_{\mathbb{R}^d} \int_{\mathbb{S}^{d-1}} B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_*\left(\phi'+\phi_*'-\phi-\phi_* \right)\diff{\omega} \diff{v} \diff{v_*},
\end{equation}
where $g := v - v_\ast$, $\hat{g}$ denotes the unit vector in the direction of $g$, $\phi(t,v)$ a regular test function, $B_{\omega}$ the binary collision kernel in the $\omega$-representation, and we use the shorthands $f$ for $f(t,v)$, $f_\ast$ for $f_\ast(t, v_\ast)$ and so on. This operator clearly has collisional invariants $1$ and $v_i$ ($i = 1, 2, 3$), namely,
\begin{equation}
  \int_{\mathbb{R}^3} Q(f, f)(v) \diff v = \int_{\mathbb{R}^3} Q(f,f)(v)v_i\diff v = 0,
\end{equation}
where $v_i$ denotes the $i$-th component of the velocity $v$.

Note that \eqref{omega} can be also written in the so-called $\sigma$-representation, where $\sigma$ is related to $\omega$ in the following way:
\begin{equation} \label{relation}
(g\cdot \omega)\omega =\frac{1}{2}(g-|g|\sigma), \quad g:=v-v_*.
\end{equation}
In $\sigma$-representation, the parametrization (\ref{omega}) becomes
\begin{align*} 
\left\{
\begin{array}{l}
\displaystyle v'=\frac{v+v_*}{2}+\frac{1-e}{4}(v-v_*)+\frac{1+e}{4}|v-v_*|\sigma, \\[8pt]
\displaystyle v_*'=\frac{v+v_*}{2}- \frac{1-e}{4}(v-v_*)-\frac{1+e}{4}|v-v_*|\sigma.
\end{array}\right.
\end{align*}
The transformation between $\omega$ and $\sigma$ representations is given as follows:
\begin{align*}
& B_{\omega}(|g|, |\omega\cdot \hat{g}|) \diff{\omega} \diff{v_*}  = B_{\sigma}(|g|,\sigma\cdot \hat{g})\diff{\sigma} \diff{v_*}  , \\
& B_{\omega}(|g|, |\omega\cdot \hat{g}|) =|2(\omega\cdot \hat{g})|^{d-2}B_{\sigma}(|g|,1-2(\omega\cdot\hat{g})^2),
\end{align*}
where $B_\sigma$ denotes the collision kernels in $\sigma$-representations. Two common cases to keep in mind are the 2D pseudo (no angular dependence) Maxwell molecule, where $B_{\sigma}=B_{\omega}=1$; and the 3D hard sphere, with $B_{\sigma}=|g|$, $B_{\omega}=2|g\cdot\omega|$. Hereafter, we shall use the collision operator in its $\sigma$-representation
\begin{equation} \label{col_sigma}
\int Q(f,f)(v)\phi(v)\diff{v}= \frac{1}{2} \int_{\mathbb{R}^d} \int_{\mathbb{R}^d} \int_{\mathbb{S}^{d-1}}
B_{\sigma}(|g|,\sigma\cdot \hat{g})ff_*\left(\phi'+\phi_*'-\phi-\phi_* \right)\diff{\sigma}\diff{v}\diff{v_*}.
\end{equation}
since its convenience for numerical purposes, which will be discussed with more details later.

\subsection{Inelastic Enskog-Boltzmann equation with heating sources}

Consider the inelastic Boltzmann equation with a heating source \eqref{inBoltz}. We shall revisit a few properties of this equation. First, for a solution $f$ to \eqref{inBoltz}, the conservation of mass and momentum becomes straightfoward taking $\phi = 1$ and $\phi = v$ respectively in \eqref{col_sigma}. In other words,
\begin{equation}
\rho := \int f\diff{v}\equiv \rho_0, \quad  u :=\frac{1}{\rho}\int fv \diff{v}\equiv u_0.
\end{equation}
where $\rho$ is the density and $u$ the bulk velocity. Similarly, define the kinetic energy as
\begin{equation}
  E := \frac{1}{2}\int f |v|^2\diff v
\end{equation}
and take $\phi = |v|^2$, one obtains the relation of energy dissipation
\begin{equation}\label{ki_energy}
\partial_t E -\varepsilon d\rho_0=-\frac{1-e^2}{16}\iiint B_{\sigma}(|g|,\sigma\cdot \hat{g})|g|^2(1-\sigma\cdot \hat{g})ff_*\diff{\sigma}\diff{v}\diff{v_*},
\end{equation}
where we used (\ref{lossenergy}) and (\ref{relation}). Consider the following collision kernel, namely, VHS (variable hard sphere) model
\begin{equation} \label{kernel}
B_{\sigma}(|g|,\sigma\cdot \hat{g})=C_{\lambda}|g|^{\lambda}b_{\lambda}(\sigma\cdot \hat{g}),
\end{equation}
where $C_{\lambda}$ is some constant and $b_{\lambda}$ is some function. For Maxwell molecule, i.e., $\lambda=0$ in (\ref{kernel}), equation \eqref{ki_energy} becomes
\begin{equation} 
\partial_t E -\varepsilon d\rho_0=-\frac{1-e^2}{16}C_0\iiint |g|^{2}b_0(\sigma\cdot \hat{g})(1-\sigma\cdot \hat{g})ff_*\,\diff{\sigma}\,\diff{v}\,\diff{v_*}.
\end{equation}
Note that
\begin{equation}
\int C_0b_0(\sigma\cdot \hat{g})(1-\sigma\cdot \hat{g})\,\diff{\sigma}
\end{equation}
is a constant regardless of $\hat{g}$ and, without loss of generality, is assumed to be $1$. Then
\begin{align} 
  \partial_t E -\varepsilon d\rho_0 &=-\frac{1-e^2}{16}\iint |g|^{2}ff_*\,\diff{v}\,\diff{v_*} \notag\\
   & =-\frac{1-e^2}{16}\iint (v^2+v_*^2-2v\cdot v_*)ff_*\,\diff{v}\,\diff{v_*} \notag \\
   &=-\frac{1-e^2}{8}\left(2\rho_0E - (\rho_0 u_0)^2\right).
\end{align}

Finally, the temperature defined by
\begin{equation}
  T = \frac{1}{\rho d}\int_{\mathbb{R}^d}f|u - v|^2\diff v
\end{equation}
goes to $0$ as time approaches infinity. In particular, if the initial condition is $\rho_0=1$, $u_0=0$, then $T=\frac{2}{d}E$ satisfies
\begin{equation}
\partial_t T-2\varepsilon =-\frac{1-e^2}{4}T,
\end{equation}
whose solution is
\begin{equation}\label{soln:T}
T=\left(T_0-\frac{8\varepsilon}{1-e^2}\right)\exp{\left(-\frac{1-e^2}{4}t\right)}+\frac{8\varepsilon}{1-e^2}.
\end{equation}
\begin{rmk}
This explicit expression serves as the reference quantity for numerical tests, since in inelastic cases it is notably harder to find a special $f$ such that the collision operator has an analytical expression in contrast to elastic cases.
\end{rmk}

\section{A fast spectral algorithm of the inelastic collision operator}

In this section, we construct the fast spectral algorithm. Before that, let us first revisit the rather conventional spectral method and its direct solver. The idea of the spectral method is rather straightfoward, that is, to exploit the structure of the collision operator in its Fourier space. Specificallly, by choosing the test function $\phi(v)=e^{-i\frac{\pi}{L}k\cdot v}$ in the weak form \eqref{col_sigma}, one can obtain the expression of the $k$-th mode in the Fourier expansion of $Q$:
\begin{equation} \label{sum1}
\hat{Q}_k=\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}G(l,m)\hat{f}_l\hat{f}_m,
\end{equation} 
where the weight $G(l,m)$ is given by
\begin{equation*}  \label{G}
G(l,m)=\int_{\mathbb{R}^d}e^{-i\frac{\pi}{L}m \cdot g}\left[\int_{S^{d-1}}B_{\sigma}(|g|,\sigma\cdot \hat{g})\left(e^{-i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot (|g|\sigma-g)}-1\right)\diff{\sigma}\right]\diff{g},
\end{equation*}
here $g$ needs to be truncated properly as was done for the elastic case \cite{gamba_hu}.

For the three-dimensional VHS model \eqref{kernel}, it follows by a straightfoward calculation that the formula for $G$ reduces to
\begin{equation} \label{direct_G}
  G(l,m)=16\pi^2C_{\gamma}\int_0^R \rho^{\gamma+2}\left[\text{Sinc}\left(\frac{\pi}{L}\left|\frac{1+e}{4}(l+m)-m\right|\rho\right)\text{Sinc}\left(\frac{\pi}{L}\frac{1+e}{4}|l+m|\rho\right)-\text{Sinc}\left(\frac{\pi}{L}\rho|m|\right)\right]\diff\rho.
\end{equation}
Hence, one can construct a direct spectral solver by evaluating the numerical quadrature of \eqref{direct_G} and proceed the summation of \eqref{sum1}. However, as is mentioned before, this approach results in a high computational cost of $O(N^6)$ and storage of $O(N^6)$, apparently because there is no convolution structure in the summation to follow. In applications, the scale of problems that one can compute using this approach may become quickly bottlenecked due to the computational limitations, which motivates the study of a fast spectral algorithm.

We begin this exploration by asking ourselves whether there are any underlying convolution structures to be unveiled. While it becomes unclear once using the final analytical expression \eqref{direct_G} of $G(l,m)$, there are in fact hidden convolution structures in equation \eqref{G}, namely, to consider a partial numerical quadrature of this intermidiate form instead of integrating it analytically. In particular, the idea of our fast algorithm is to separate the weight $G$ as $G(l,m)\approx\sum_{t=1}^T\alpha_t(l+m)\beta_t(m)$ using quadrature rules.

\subsection{2D case}

In 2D VHS  case $B_{\sigma}(|g|,\sigma\cdot \hat{g})=C_{\gamma}|g|^{\gamma}$:
\begin{equation}
  \int_{S^1}B_{\sigma}(|g|,\sigma\cdot \hat{g})\left(e^{-i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot (|g|\sigma-g)}-1\right)\diff{\sigma}=2\pi C_{\gamma}|g|^{\gamma}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot g}\text{J}_0\left(\frac{\pi}{L}\frac{1+e}{4}|l+m||g|\right)-1\right],
\end{equation}
then,
\begin{equation}
  G(l,m)=\sum_{\rho,\sigma}w_{\rho}w_{\sigma}2\pi C_{\gamma}\rho^{\gamma+1}e^{-i\frac{\pi}{L}\rho m \cdot \sigma}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}\rho (l+m)\cdot \sigma}\text{J}_0\left(\frac{\pi}{L}\frac{1+e}{4}|l+m|\rho\right)-1\right],
\end{equation}
where $w_\rho$ and $w_\sigma$ are the weights of quadratures for $\rho$ and $\sigma$ respectively. $J_0$ is $0$-th order Bessel function. Therefore,
\begin{equation} 
  \hat{Q}_k=\sum_{\rho,\sigma}w_{\rho}w_{\sigma}2\pi C_{\gamma}\rho^{\gamma+1}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}\rho k\cdot \sigma}\text{J}_0\left(\frac{\pi}{L}\frac{1+e}{4}\rho |k|\right)-1\right]\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[e^{-i\frac{\pi}{L}\rho m \cdot \sigma}\hat{f}_m\right],
\end{equation} 
or the loss term can be computed separately as
\begin{equation} 
\hat{Q}_k^-=\sum_{\rho,\sigma}w_{\rho}w_{\sigma}2\pi C_{\gamma}\rho^{\gamma+1}\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[e^{-i\frac{\pi}{L}\rho m \cdot \sigma}\hat{f}_m\right]=\sum_{\rho}w_{\rho}4\pi^2 C_{\gamma}\rho^{\gamma+1}\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[\text{J}_0\left(\frac{\pi}{L}\rho |m|\right)\hat{f}_m\right].
\end{equation}
We refer the later method as ``separate'' method and previous one as ``full'' method correspondingly.

\subsection{3D case}

In 3D VHS case $B_{\sigma}(|g|,\sigma\cdot \hat{g})=C_{\gamma}|g|^{\gamma}$:
\begin{equation}
\int_{S^2}B_{\sigma}(|g|,\sigma\cdot \hat{g})\left(e^{-i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot (|g|\sigma-g)}-1\right)\,\diff{\sigma}=4\pi C_{\gamma}|g|^{\gamma}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot g}\text{Sinc}\left(\frac{\pi}{L}\frac{1+e}{4}|l+m||g|\right)-1\right],
\end{equation}
then,
\begin{equation}
G(l,m)=\sum_{\rho,\sigma}w_{\rho}w_{\sigma}4\pi C_{\gamma}\rho^{\gamma+2}e^{-i\frac{\pi}{L}\rho m \cdot \sigma}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}\rho (l+m)\cdot \sigma}\text{Sinc}\left(\frac{\pi}{L}\frac{1+e}{4}|l+m|\rho\right)-1\right],
\end{equation}
therefore,
\begin{equation} 
\hat{Q}_k=\sum_{\rho,\sigma}w_{\rho}w_{\sigma}4\pi C_{\gamma}\rho^{\gamma+2}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}\rho k\cdot \sigma}\text{Sinc}\left(\frac{\pi}{L}\frac{1+e}{4}\rho |k|\right)-1\right]\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[e^{-i\frac{\pi}{L}\rho m \cdot \sigma}\hat{f}_m\right],
\end{equation} 
or the loss term can be computed separately as
\begin{equation} 
\hat{Q}_k^-=\sum_{\rho,\sigma}w_{\rho}w_{\sigma}4\pi C_{\gamma}\rho^{\gamma+2}\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[e^{-i\frac{\pi}{L}\rho m \cdot \sigma}\hat{f}_m\right]=\sum_{\rho}w_{\rho}16\pi^2 C_{\gamma}\rho^{\gamma+2}\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[\text{Sinc}\left(\frac{\pi}{L}\rho |m|\right)\hat{f}_m\right],
\end{equation}
which are the ``full" and ``separate" methods in 3D case.

\begin{rmk}
	In our numerical implementation, we use the Gauss-Legendre quadrature in the radial direction $\rho$, while for the integration in $\sigma$ we propose to use the spherical design in 3D case which is the near optimal quadrature on the sphere and requires fewer quadrature points than tensor product based Gauss quadratures for a large class of functions (2D case is easy to treat by using traditional quadratures). To gain a better idea of how this quadrature works, a error test of this quadrature will be given in next section.
\end{rmk}

\section{Numerical examples}

In this section, we first verify the accuracy and efficiency of the proposed method with extensive numerical studies in two dimensional cases, and then provide three dimensional numerical results as well. Note that unlike the elastic case where the analytical results of the collision kernel are accessible by choosing a suitable function $f$, the inelastic case, however, only has the analytical formula for its macro quantities, such as the temperature $T$ as is shown in \eqref{soln:T}. This means in order to provide a trustworthy verification of our solver, one needs a numerical scheme to solve the inelastic Boltzmann equation \eqref{inBoltz}.

In the following numerical simulations, we use our fast spectral solver to evaluate the collision operator, paired with a standard fourier spectral method to treat the diffusion term $\varepsilon\Delta f$ and the SSP-RK3 (the 3rd order Strong Stability-Preserving Runge-Kutta method) \cite{shuTadmor} for time discretizations.

\subsection{Maxwell molecules - 2D example 1}

In this part, we perform several 2D examples to verify the accuracy and efficiency of our method. We consider equation \eqref{inBoltz} with $\varepsilon = 10^{-6}$ and Maxwell molecule, i.e., $\lambda = 0$. The initial condition for $f$ is given as the the 2D BKW solution
\begin{equation} \label{ext1}
  f(0,v) = \frac{1}{2\pi K^2}\exp\left(-\frac{v^2}{2K}\right)\left(2K-1+\frac{1-K}{2K}v^2\right),
\end{equation}
where $K=1-\exp(-1/16)/2$. One can easily check that $\rho_0 = 1$, $u_0 = 0$ and $T_0 = E_0$ in this case.

The macro quantity we compute is temperature $T$ at some given final time $T_\text{final}$. The numercial result $T_\text{num}$ is obtained by taking the moments of the numerical solution $f_\text{num}$, which is computed by RK3 and our fast spectral method with $N = 64$ and $M = 30$. The reference solution $T_\text{ref}$ is obtained by using the exact form \eqref{soln:T}. Finally, the error is measured by $|T_\text{num} - T_\text{ref}|$.

% (\ref{ext1}) satisfies exactly the spatially homogeneous Boltzmann equation
% \begin{equation} \label{homo}
%   \frac{\partial f}{\partial t} = \mathcal{Q}(f),
% \end{equation}

\paragraph{\bf Convergence in time} In order to suppress the error due to time discretization, we first perform a convergence test of 3rd Runge-Kutta SSP method used in the simulation:
\begin{align}
  k_1 &= L(f^n), \notag \\
  k_2 &= L(f^n + \frac{1}{2}k_1\Delta t), \notag \\
  k_3 &= L(f^n - k_1\Delta t + 2k_2\Delta t), \notag \\
  f^{n+1} &= f^n + \frac{1}{6}(k_1 + 4k_2 + k_3)\Delta t,
\end{align}
where $L$ is spatial discretization.

% \begin{figure}[htb]
%   \centering
%   \includegraphics[width = .48\linewidth]{figs/sep/dt_2d_bkw_e=02}\hfill
%   \includegraphics[width = .48\linewidth]{figs/full/dt_2d_bkw_e_full=02}
%   \caption{Convergence of RK3 with $e = 0.2$. Left: separate. Right: full.}
%   \label{dt_bkw_e02}
% \end{figure}
% \begin{figure}[htb]
%   \centering
%   \includegraphics[width=0.5\textwidth]{figs/sep/dt_2d_bkw_e=05}\hfill
%   \includegraphics[width=0.5\textwidth]{figs/full/dt_2d_bkw_e_full=05}
%   \caption{Convergence of RK3 with $e = 0.5$. Left: separate. Right: full.}
%   \label{dt_bkw_e05}
% \end{figure}
% \begin{figure}[htb]
%   \centering
%   \includegraphics[width=0.5\textwidth]{figs/sep/dt_2d_bkw_e=08}\hfill
%   \includegraphics[width=0.5\textwidth]{figs/full/dt_2d_bkw_e_full=08}
%   \caption{Convergence of RK3 with $e = 0.8$. Left: separate. Right: full.}
%   \label{dt_bkw_e08}
% \end{figure}

\begin{figure}[H]
  \centering
  \subfloat[$e = 0.2$, separate]{\includegraphics[width = 0.5\linewidth]{figs/sep/dt_2d_bkw_e=02}}\hfill
  \subfloat[$e = 0.2$, full]{\includegraphics[width = 0.5\linewidth]{figs/full/dt_2d_bkw_e_full=02}} \\
  \subfloat[$e = 0.5$, separate]{\includegraphics[width = 0.5\linewidth]{figs/sep/dt_2d_bkw_e=05}}\hfill
  \subfloat[$e = 0.5$, full]{\includegraphics[width = 0.5\linewidth]{figs/full/dt_2d_bkw_e_full=05}} \\
  \subfloat[$e = 0.8$, separate]{\includegraphics[width = 0.5\linewidth]{figs/sep/dt_2d_bkw_e=08}}\hfill
  \subfloat[$e = 0.8$, full]{\includegraphics[width = 0.5\linewidth]{figs/full/dt_2d_bkw_e_full=08}}
  \caption{Convergence test in time of the 2D example 1 with $N = 32$ for various values of $e=0.2$, $e=0.5$ and $e=0.8$.}
  \label{dt_conv_1}
\end{figure}

In Figure~\ref{dt_conv_1} we plot the relation between the errors versus different $\Delta t$s for $e = 0.2, 0.5$ and $0.8$, $T_\text{final}=2$. A third order convergence rate can be seen very easily. Since in this test we fix all the parameters in our spectral method as $\Delta t$ changes, these figures also imply that the error of the spectral method computing the collision is really small, or at least around $O(10^{-9})$.

This test shows that our RK3 method in time direction is indeed 3rd-order and in the following convergence tests of $N$ we will set $\Delta t = 0.01$

\paragraph{\bf Convergence in $N$} We then perform the convergence test of the spectral method. As mentioned previously we set $\Delta t = 0.01$ and $M = 30$, also $T_\text{final} = 2$. As $N$ increases from $8$ to $128$, we calculate the error for different $e$s using both ``separate" and ``full" method as shown in Table~\ref{table 1}.

\begin{table}[H]
  \centering
  \subfloat[$e = 0.2$]{
  \begin{tabu} to 0.6\linewidth {X[1, c] X[3, c] X[3, c]}
    \toprule
    $N$ & Separate & Full \\
    \midrule
    8 & 9.21116565e-01 & 9.21116565e-01 \\
    16 & 1.27634481e-02 & 1.27640374e-02 \\
    32 & 6.79544555e-06 & 6.79745658e-06 \\
    64 & 2.34851361e-10 & 2.36438646e-10 \\
    128 & 6.30565600e-11 & 6.13890050e-11 \\
    \bottomrule
  \end{tabu}
  } \\
  \subfloat[$e = 0.5$]{
  \begin{tabu} to 0.6\linewidth {X[1, c] X[3, c] X[3, c]}
    \toprule
    $N$ & Separate & Full \\
    \midrule
    8 & 7.98706096e-01 & 7.98706096e-01 \\
    16 & 6.42641236e-03 & 6.42644165e-03 \\
    32 & 4.55713801e-06 & 4.55730861e-06 \\
    64 & 4.93595165e-11 & 4.93770580e-11 \\
    128 & 3.13873372e-11 & 3.14279713e-11 \\
    \bottomrule
  \end{tabu}
  } \\
  \subfloat[$e = 0.8$]{
  \begin{tabu} to 0.6\linewidth {X[1, c] X[3, c] X[3, c]}
    \toprule
    $N$ & Separate & Full \\
    \midrule
    8 & 5.52666966e-01 & 5.52666966e-01 \\
    16 & 4.88821204e-04 & 4.88586534e-04 \\
    32 & 1.14430393e-07 & 1.13897201e-07 \\
    64 & 9.82359749e-11 & 9.82520731e-11 \\
    128 & 1.00099595e-10 & 1.00117470e-10 \\
    \bottomrule
  \end{tabu}
  }
  \caption{Convergence test in $N$ of the 2D example 1 with $\Delta t = 0.01$ for various values of $e=0.2$, $e=0.5$ and $e=0.8$.}
  \label{table 1}
\end{table}

These results show that our method indeed can achieve spectral accuracy.

\subsection{Maxwell molecules - 2D example 2}

Follow the set up in previous subsection, we change the initial condition from 2D BKW solution to a exponential double-well function
\begin{equation}
  f(0, v) = \frac{0.8}{\pi}\left(e^{-4|v-2|^2} + e^{-|v+0.5|^2}\right).
\end{equation}

Again both tests of convergence in time and $N$ are performed and the results are shown in the following Figure~\ref{conv_dt_2} and Table~\ref{table 2}.

\begin{figure}[H]
  \centering
  \subfloat[$e = 0.2$, separate]{\includegraphics[width = 0.5\linewidth]{figs/sep/dt_2d_e=02}} \\
  \subfloat[$e = 0.5$, separate]{\includegraphics[width = 0.5\linewidth]{figs/sep/dt_2d_e=05}} \\
  \subfloat[$e = 0.8$, separate]{\includegraphics[width = 0.5\linewidth]{figs/sep/dt_2d_e=08}}
  \caption{Convergence test in time of the 2D example 2 with $N = 32$ for various values of $e=0.2$, $e=0.5$ and $e=0.8$.}
  \label{conv_dt_2}
\end{figure}

\begin{table}[H]
  \centering
  \subfloat[$e = 0.2$]{
  \begin{tabu} to 0.6\linewidth {X[1, c] X[3, c] X[3, c]}
    \toprule
    $N$ & Separate & Full \\
    \midrule
    8 & 1.25303916e-01 & 1.25303916e-01 \\
    16 & 1.41601818e-02 & 1.42811856e-02 \\
    32 & 1.21162093e-04 & 8.50383206e-05 \\
    64 & 8.65618628e-08 & 5.75217760e-05 \\
    128 & 2.64749862e-08 & 5.74603408e-05 \\
    \bottomrule
  \end{tabu}
  } \\
  \subfloat[$e = 0.5$]{
  \begin{tabu} to 0.6\linewidth {X[1, c] X[3, c] X[3, c]}
    \toprule
    $N$ & Separate & Full \\
    \midrule
    8 & 9.06935081e-02 & 9.06935081e-02 \\
    16 & 2.06153352e-02 & 2.07345865e-02 \\
    32 & 1.08598123e-04 & 7.68575010e-05 \\
    64 & 3.61540865e-08 & 4.58166915e-05 \\
    128 & 4.84827622e-09 & 4.57852460e-05 \\
    \bottomrule
  \end{tabu}
  } \\
  \subfloat[$e = 0.8$]{
  \begin{tabu} to 0.6\linewidth {X[1, c] X[3, c] X[3, c]}
    \toprule
    $N$ & Separate & Full \\
    \midrule
    8 & 4.21932177e-02 & 4.21932177e-02 \\
    16 & 2.19257970e-02 & 2.17989934e-02 \\
    32 & 1.25546782e-04 & 1.17000137e-04 \\
    64 & 1.55334599e-08 & 2.42867607e-05 \\
    128 &  5.91159477e-09 & 2.42767854e-05 \\
    \bottomrule
  \end{tabu}
  }
  \caption{Convergence test in $N$ of the 2D example 2 with $\Delta t = 0.01$ for various values of $e=0.2$, $e=0.5$ and $e=0.8$.}
  \label{table 2}
\end{table}

One can find the error becomes worse compared to BKW initial condition, especially for the ``full'' method (see Table~\ref{table 2}). This is also why no time convergence test for ``full'' method is done for this example.

\subsection{3D example}

For 3D examples, we first test the accuracy of spherical design. In order to do this we still consider the Maxwell molecules. After this we will verify the physical Haff's cooling law by considering a hard sphere molecule.

\paragraph{\bf Accuracy of spherical design}
Consider the collision kernel $B = \frac{1}{4\pi}$ is a constant and take the initial condition as
\begin{equation}\label{3Dbkw}
  f(t_0, v) = \frac{1}{2(2\pi K(t_0))^{3/2}}\exp\left(\frac{v^2}{2K(t_0)}\right)\left(\frac{5K(t_0)-3}{K(t_0}+\frac{1-K(t_0)}{K^2(t_0)}v^2\right),
\end{equation}
where $K(t) = 1 - \exp(-t_0/6)$ and we take $t_0 = 6.5$.

We set $e=0.2$ and $T_\text{final} = 1$, $\Delta t = 0.01$ with $N = 32$. By comparing the numerical temperature with analytical expression for different spherical design points, we obtain the following Table~\ref{sph_design}

\begin{table}[H]
  \centering
  \begin{tabu} to 0.8\linewidth {X[1,c] X[2,c]}
    \toprule
    Spherical Design & $e = 0.2$ \\
    \midrule
    6 & 0.0013957739849754791 \\
    12 & 9.9706271716293315e-05 \\
    32 & 2.2499901350947482e-06 \\
    48 & 2.4272557155313734e-06 \\
    70 & 2.4703481364962698e-06 \\
    94 & 2.4703481364962698e-06 \\
    120 & 2.453380453903975e-06 \\
    \bottomrule
  \end{tabu}
  \caption{Errrors of spherical design.}
  \label{sph_design}
\end{table}

\paragraph{\bf Haff's cooling law - hard sphere molecule}
In this example, we try to observe the famous Haff's cooling law \cite{haff1983}. In the first stage of the evolution from a flux-free spatial uniform state, the granular gases remain homogeneous. This stage is called the homogeneous cooling state, in which the granular temperature follows Haff's law, first derived by Haff in 1983 \cite{haff1983, oxfordbook}, in which the temperature decays in the order of $t^{-2}$.
In this test, we use the same example as in the previous subsection, but for hard sphere molecules. That is, consider the collision kernel
\begin{equation}
  B = \frac{1}{4\pi}|g| = \frac{1}{4\pi}|v-v_\ast|.
\end{equation}
with the reconstitution coefficient $ e = 0.2$. The initial datum is chosen as \eqref{3Dbkw}, and we compute till the final time $T_\text{final} = 3$ with numerical discretizations $\Delta t = 0.01$ and $N = 32$.
\begin{figure}[H]
  \centering
  \includegraphics[width = .8\linewidth]{figs/Haff's_cooling}
  \caption{Haff's cooling law}
  \label{Haff_cooling}
\end{figure}
As can be seen in Figure~\ref{Haff_cooling}, our numerical temperature decays as 
$$ T(t) = \frac{2.25}{ (1+0.36t)^2},$$
which agrees exactly with the Haff's formula $$T(t) = \frac{T_0}{(1+C_0 t)^2}.$$

\section*{References}

\bibliography{mybibfile}

\end{document}

### 2018-07-08 paper/draft/elsarticle_v2.tex
cleanup

--- 2018-07-08 paper/draft/elsarticle-template.tex
@@
 \usepackage{booktabs}
-\usepackage{float}
 
@@
 \title{A fast spectral method for inelastic collision operator and the heated granular flow\tnoteref{mytitlenote}}
-% \tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}
+\tnotetext[mytitlenote]{Fully documented templates are available in the elsarticle package on \href{http://www.ctan.org/tex-archive/macros/latex/contrib/elsarticle}{CTAN}.}
 
 %% Group authors per affiliation:
-% \author{Jingwei Hu, Zheng Ma\fnref{myfootnote}}
-% \address{Department of Mathematics, Purdue University, 150 N. University Street, West Lafayette, IN, USA}
-% \fntext[myfootnote]{}
+\author{Elsevier\fnref{myfootnote}}
+\address{Radarweg 29, Amsterdam}
+\fntext[myfootnote]{Since 1880.}
 
 %% or include affiliations in footnotes:
-\author[mymainaddress]{Jingwei Hu}
-\ead{jingweihu@purdue.edu}
+\author[mymainaddress,mysecondaryaddress]{Elsevier Inc}
+\ead[url]{www.elsevier.com}
 
-\author[mymainaddress]{Zheng Ma}
-% \cortext[mycorrespondingauthor]{Corresponding author}
-\ead{ma531@purdue.edu}
+\author[mysecondaryaddress]{Global Customer Service\corref{mycorrespondingauthor}}
+\cortext[mycorrespondingauthor]{Corresponding author}
+\ead{support@elsevier.com}
 
-\address[mymainaddress]{150 N. University Street, West Lafayette, IN, USA}
-% \address[mysecondaryaddress]{360 Park Avenue South, New York}
+\address[mymainaddress]{1600 John F Kennedy Boulevard, Philadelphia}
+\address[mysecondaryaddress]{360 Park Avenue South, New York}
 
@@
 
-% \begin{keyword}
-% \texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
-% \MSC[2010] 00-01\sep  99-00
-% \end{keyword}
+\begin{keyword}
+\texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
+\MSC[2010] 00-01\sep  99-00
+\end{keyword}
 
@@
 For convenience, we use $Q$ instead of $Q_{\text{in}}$ in the rest of this paper, whose notation should not be confused with the usual elastic collision operator.
-Another widely used model, first introduced by van Noije and Ernst \cite{noije}, is the spatial homogeneous inelastic Boltzmann equation based on Enskog-Boltzmann model with a heat source:
+Another widely used model, first introduced by van Noije and Ernst \cite{}, is the spatial homogeneous inelastic Boltzmann equation based on Enskog-Boltzmann model with a heat source:
 \begin{equation} \label{inBoltz}
@@
 At each time step, the construction of the inelastic collision operator requires $O(N^{2d})$ operations and $O(N^{2d})$ storage in a direct numerical scheme.
-As is pointed out in \cite{filbet2005}, although the loss term in the inelastic collision can be evaluated only in $O(N^d \log N)$ operations thanks to its convolution structure, the cost of the gain part is still rather formidable.
+As is pointed out in \cite{filbet}, although the loss term in the inelastic collision can be evaluated only in $O(N^d \log N)$ operations thanks to its convolution structure, the cost of the gain part is still rather formidable.
 A natural question of course remains to reduce the computational cost for the entire collision operator, as well as the numerical storage -- in order words, to fully exploit the structure of the granular collision operator.
-In this paper, we propose a fast spectral algorithm for the inelastic collision operator, inspired by a sequence of studies on elastic Boltzmann operator \cite{fast1999, mouhot2006fast, hu2012fast, gamba_hu}.
-To be specific, in contrast to a direct spectral solver, in 3D this algorithm reduces the computational cost from $O(N^{6})$ to $O\left(MN^{4}\log(N)\right)$ and the storage from $O(N^{6})$ to $O(MN^{4})$, where $M \ll N^{2}$ is the number of quadrature points on $\mathbb{S}^{2}$.
+In this paper, we propose a fast spectral algorithm for the inelastic collision operator, inspired by a sequence of studies on elastic Boltzmann operator \cite{gambaHu, lexingHu, , }.
+To be specific, in contrast to a direct spectral solver, in 3D this algorithm reduces the computational cost from $O(N^{6})$ to $O\left(MN^{4}\log(N)\right)$ and the storage from $O(N^{6}$ to $O(MN^{4})$, where $M \ll N^{2}$ is the number of quadrature points on $\mathbb{S}^{2}$.
 
@@
 To describe the inelastic binary collision, a reconstitution coefficient $e$ is introduced.
-Specifically, assuming two particles with velocities $v$ and $v_*$ are going to collide, after the collision, the velocities denoted by $v'$ and $v_*'$ are given by the so-called $\omega$-representation \cite{Villani2006}
+Specifically, assuming two particles with velocities $v$ and $v_*$ are going to collide, after the collision, the velocities denoted by $v'$ and $v_*'$ are given by the so-called $\omega$-representation \cite{vilani}
 \begin{align}\label{omega}
@@
 \end{equation}
-where we used (\ref{lossenergy}) and (\ref{relation}). Consider the following collision kernel, namely, VHS (variable hard sphere) model
+where we used (\ref{lossenergy}) and (\ref{relation}). Consider the following collision kernel
 \begin{equation} \label{kernel}
@@
 
-In 2D VHS  case $B_{\sigma}(|g|,\sigma\cdot \hat{g})=C_{\gamma}|g|^{\gamma}$:
+In 2D, and VHS case $B_{\sigma}(|g|,\sigma\cdot \hat{g})=C_{\gamma}|g|^{\gamma}$:
 \begin{equation}
-  \int_{S^1}B_{\sigma}(|g|,\sigma\cdot \hat{g})\left(e^{-i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot (|g|\sigma-g)}-1\right)\diff{\sigma}=2\pi C_{\gamma}|g|^{\gamma}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot g}\text{J}_0\left(\frac{\pi}{L}\frac{1+e}{4}|l+m||g|\right)-1\right],
+  \int_{S^1}B_{\sigma}(|g|,\sigma\cdot \hat{g})\left(e^{-i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot (|g|\sigma-g)}-1\right)\,\diff{\sigma}=2\pi C_{\gamma}|g|^{\gamma}\left[ e^{i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot g}\text{J}_0\left(\frac{\pi}{L}\frac{1+e}{4}|l+m||g|\right)-1\right],
 \end{equation}
-then,
+then let $\rho=|g|$, $\sigma=\hat{g}$,
 \begin{equation}
@@
 \end{equation}
-where $w_\rho$ and $w_\sigma$ are the weights of quadratures for $\rho$ and $\sigma$ respectively. $J_0$ is $0$-th order Bessel function. Therefore,
+therefore,
 \begin{equation} 
@@
 \end{equation}
-We refer the later method as ``separate'' method and previous one as ``full'' method correspondingly.
+We refer the later method as "seperate" method and previous one as "full" method correspondingly.
 
@@
 
-In 3D VHS case $B_{\sigma}(|g|,\sigma\cdot \hat{g})=C_{\gamma}|g|^{\gamma}$:
+In 3D, and VHS case $B_{\sigma}(|g|,\sigma\cdot \hat{g})=C_{\gamma}|g|^{\gamma}$:
 \begin{equation}
@@
 \end{equation}
-then,
+then let $\rho=|g|$, $\sigma=\hat{g}$,
 \begin{equation}
@@
 \end{equation}
-which are the ``full" and ``separate" methods in 3D case.
-
-\begin{rmk}
-	In our numerical implementation, we use the Gauss-Legendre quadrature in the radial direction $\rho$, while for the integration in $\sigma$ we propose to use the spherical design in 3D case which is the near optimal quadrature on the sphere and requires fewer quadrature points than tensor product based Gauss quadratures for a large class of functions (2D case is easy to treat by using traditional quadratures).
-To gain a better idea of how this quadrature works, a error test of this quadrature will be given in next section.
-\end{rmk}
+which are the "full" and "separate" methods in 3D case.
 
@@
 
-In the following numerical simulations, we use our fast spectral solver to evaluate the collision operator, paired with a standard fourier spectral method to treat the diffusion term $\varepsilon\Delta f$ and the SSP-RK3 (the 3rd order Strong Stability-Preserving Runge-Kutta method) \cite{shuTadmor} for time discretizations.
-
-\subsection{Maxwell molecules - 2D example 1}
+In the following numerical simulations, we use our fast spectral solver to evaluate the collision operator, paired with a standard fourier spectral method to treat the diffusion term $\varepsilon\Delta f$ and the SSP-RK3 (the 3rd order Strong Stability-Preserving Runge-Kutta method) \cite{ShuTadmor} for time discretizations.
+\subsection{2D examples - Maxwell molecules}
 
@@
 
+In Figure~\ref{dt_conv_1} we plot the relation between the errors versus different $\Delta t$s for $e = 0.2, 0.5$ and $0.8$, $T_\text{final}=2$.
+A third order convergence rate can be seen very easily.
+Since in this test we fix all the parameters in our spectral method as $\Delta t$ changes, these figures also imply that the error of the spectral method computing the collision is really small, or at least around $O(10^{-9})$.
 % \begin{figure}[htb]
@@
 
-\begin{figure}[H]
+\begin{figure}[htb]
   \centering
@@
   \subfloat[$e = 0.8$, full]{\includegraphics[width = 0.5\linewidth]{figs/full/dt_2d_bkw_e_full=08}}
-  \caption{Convergence test in time of the 2D example 1 with $N = 32$ for various values of $e=0.2$, $e=0.5$ and $e=0.8$.}
+  \caption{convergence in time 1}
   \label{dt_conv_1}
@@
 
-In Figure~\ref{dt_conv_1} we plot the relation between the errors versus different $\Delta t$s for $e = 0.2, 0.5$ and $0.8$, $T_\text{final}=2$.
-A third order convergence rate can be seen very easily.
-Since in this test we fix all the parameters in our spectral method as $\Delta t$ changes, these figures also imply that the error of the spectral method computing the collision is really small, or at least around $O(10^{-9})$.
+\begin{figure}[htb]
+  \centering
+  \subfloat[$e = 0.2$, separate]{\includegraphics[width = 0.5\linewidth]{figs/sep/dt_2d_e=02}} \\
+  \subfloat[$e = 0.5$, separate]{\includegraphics[width = 0.5\linewidth]{figs/sep/dt_2d_e=05}} \\
+  \subfloat[$e = 0.8$, separate]{\includegraphics[width = 0.5\linewidth]{figs/sep/dt_2d_e=08}}
+  \caption{convergence in time 2}
+\end{figure}
 
@@
 As mentioned previously we set $\Delta t = 0.01$ and $M = 30$, also $T_\text{final} = 2$.
-As $N$ increases from $8$ to $128$, we calculate the error for different $e$s using both ``separate" and ``full" method as shown in Table~\ref{table 1}.
+As $N$ increases from $8$ to $128$, we calculate the error for different $e$s using both "separate" and "full" method as shown in Table~\ref{N_bkw_e02}, Table~\ref{N_bkw_e05} and Table~\ref{N_bkw_e08}.
 
-\begin{table}[H]
+\begin{table}[htb]
   \centering
@@
   }
-  \caption{Convergence test in $N$ of the 2D example 1 with $\Delta t = 0.01$ for various values of $e=0.2$, $e=0.5$ and $e=0.8$.}
+  \caption{convergence in N 1}
   \label{table 1}
@@
 
-These results show that our method indeed can achieve spectral accuracy.
-
-\subsection{Maxwell molecules - 2D example 2}
-
-Follow the set up in previous subsection, we change the initial condition from 2D BKW solution to a exponential double-well function
-\begin{equation}
-  f(0, v) = \frac{0.8}{\pi}\left(e^{-4|v-2|^2} + e^{-|v+0.5|^2}\right).
-\end{equation}
-
-Again both tests of convergence in time and $N$ are performed and the results are shown in the following Figure~\ref{conv_dt_2} and Table~\ref{table 2}.
-
-\begin{figure}[H]
-  \centering
-  \subfloat[$e = 0.2$, separate]{\includegraphics[width = 0.5\linewidth]{figs/sep/dt_2d_e=02}} \\
-  \subfloat[$e = 0.5$, separate]{\includegraphics[width = 0.5\linewidth]{figs/sep/dt_2d_e=05}} \\
-  \subfloat[$e = 0.8$, separate]{\includegraphics[width = 0.5\linewidth]{figs/sep/dt_2d_e=08}}
-  \caption{Convergence test in time of the 2D example 2 with $N = 32$ for various values of $e=0.2$, $e=0.5$ and $e=0.8$.}
-  \label{conv_dt_2}
-\end{figure}
-
-\begin{table}[H]
+\begin{table}[htb]
   \centering
@@
   }
-  \caption{Convergence test in $N$ of the 2D example 2 with $\Delta t = 0.01$ for various values of $e=0.2$, $e=0.5$ and $e=0.8$.}
+  \caption{convergence in N 2}
   \label{table 2}
@@
 
-One can find the error becomes worse compared to BKW initial condition, especially for the ``full'' method (see Table~\ref{table 2}).
-This is also why no time convergence test for ``full'' method is done for this example.
+These results show that our method indeed can achieve spectral accuracy.
 
@@
 
-For 3D examples, we first test the accuracy of spherical design.
-In order to do this we still consider the Maxwell molecules.
-After this we will verify the physical Haff's cooling law by considering a hard sphere molecule.
-
 \paragraph{\bf Accuracy of spherical design}
-Consider the collision kernel $B = \frac{1}{4\pi}$ is a constant and take the initial condition as
-\begin{equation}\label{3Dbkw}
-  f(t_0, v) = \frac{1}{2(2\pi K(t_0))^{3/2}}\exp\left(\frac{v^2}{2K(t_0)}\right)\left(\frac{5K(t_0)-3}{K(t_0}+\frac{1-K(t_0)}{K^2(t_0)}v^2\right),
-\end{equation}
-where $K(t) = 1 - \exp(-t_0/6)$ and we take $t_0 = 6.5$.
-
-We set $e=0.2$ and $T_\text{final} = 1$, $\Delta t = 0.01$ with $N = 32$.
-By comparing the numerical temperature with analytical expression for different spherical design points, we obtain the following Table~\ref{sph_design}
-
-\begin{table}[H]
-  \centering
-  \begin{tabu} to 0.8\linewidth {X[1,c] X[2,c]}
-    \toprule
-    Spherical Design & $e = 0.2$ \\
-    \midrule
-    6 & 0.0013957739849754791 \\
-    12 & 9.9706271716293315e-05 \\
-    32 & 2.2499901350947482e-06 \\
-    48 & 2.4272557155313734e-06 \\
-    70 & 2.4703481364962698e-06 \\
-    94 & 2.4703481364962698e-06 \\
-    120 & 2.453380453903975e-06 \\
-    \bottomrule
-  \end{tabu}
-  \caption{Errrors of spherical design.}
-  \label{sph_design}
-\end{table}
-
-\paragraph{\bf Haff's cooling law - hard sphere molecule}
-In this example, we try to observe the famous Haff's cooling law \cite{haff1983}.
-In the first stage of the evolution from a flux-free spatial uniform state, the granular gases remain homogeneous.
-This stage is called the homogeneous cooling state, in which the granular temperature follows Haff's law, first derived by Haff in 1983 \cite{haff1983, oxfordbook}, in which the temperature decays in the order of $t^{-2}$.
-In this test, we use the same example as in the previous subsection, but for hard sphere molecules. That is, consider the collision kernel
-\begin{equation}
-  B = \frac{1}{4\pi}|g| = \frac{1}{4\pi}|v-v_\ast|.
-\end{equation}
-with the reconstitution coefficient $ e = 0.2$.
-The initial datum is chosen as \eqref{3Dbkw}, and we compute till the final time $T_\text{final} = 3$ with numerical discretizations $\Delta t = 0.01$ and $N = 32$.
-\begin{figure}[H]
+For 3D
+\paragraph{\bf Haff's cooling law}
+\begin{figure}[htb]
   \centering
@@
 \end{figure}
-As can be seen in Figure~\ref{Haff_cooling}, our numerical temperature decays as 
-$$ T(t) = \frac{2.25}{ (1+0.36t)^2},$$
-which agrees exactly with the Haff's formula $$T(t) = \frac{T_0}{(1+C_0 t)^2}.$$
 
@@
 
-\bibliography{mybibfile}
+% \bibliography{mybibfile}
 

### 2018-07-08 paper/paper_submit_ver/fast_spec_inelastic.tex
cleanup

\documentclass[review, times]{elsarticle}

\usepackage{lineno,hyperref}

\usepackage{color}
\modulolinenumbers[5]

\journal{Journal of Computational Physics}

%%%%%%%%%%%%%%%%%%%%%%%
%% Elsevier bibliography styles
%%%%%%%%%%%%%%%%%%%%%%%
%% To change the style, put a % in front of the second line of the current style and
%% remove the % from the second line of the style you would like to use.
%%%%%%%%%%%%%%%%%%%%%%%

%% Numbered
%\bibliographystyle{model1-num-names}

%% Numbered without titles
%\bibliographystyle{model1a-num-names}

%% Harvard
%\bibliographystyle{model2-names.bst}\biboptions{authoryear}

%% Vancouver numbered
%\usepackage{numcompress}\bibliographystyle{model3-num-names}

%% Vancouver name/year
%\usepackage{numcompress}\bibliographystyle{model4-names}\biboptions{authoryear}

%% APA style
%\bibliographystyle{model5-names}\biboptions{authoryear}

%% AMA style
%\usepackage{numcompress}\bibliographystyle{model6-num-names}

%% `Elsevier LaTeX' style
\bibliographystyle{elsarticle-num}
%%%%%%%%%%%%%%%%%%%%%%%

% personal packages
\usepackage{amsmath,amssymb}
\usepackage{subfig}
\usepackage{tabu}
\usepackage{booktabs}
\usepackage{float}

\newcommand{\rd}{\,\mathrm{d}}
\newcommand{\tv}{\,\tilde{v}}
\newcommand{\tg}{\,\tilde{g}}
\newcommand{\tf}{\,\tilde{f}}

% personal macros
\newtheorem{thm}{Theorem} \newtheorem{lem}[thm]{Lemma} \newdefinition{rmk}{Remark} \newproof{pf}{Proof}
\newproof{pot}{Proof of Theorem \ref{thm2}}

\newcommand*\diff{\mathop{}\!\mathrm{d}}
\newcommand*\Diff[1]{\mathop{}\!\mathrm{d^#1}}

\begin{document}

\begin{frontmatter}

\title{A fast spectral method for the inelastic Boltzmann collision operator and application to heated granular gases\tnoteref{ack}}
\tnotetext[ack]{This work was partially supported by NSF grant DMS-1620250 and NSF CAREER grant DMS-1654152. Support from DMS-1107291: RNMS KI-Net is also gratefully acknowledged.}

%% Group authors per affiliation:
% \author{Jingwei Hu, Zheng Ma\fnref{myfootnote}}
% \address{Department of Mathematics, Purdue University, 150 N. University Street, West Lafayette, IN, USA}
% \fntext[myfootnote]{}

%% or include affiliations in footnotes:
\author[mymainaddress]{Jingwei Hu}
\ead{jingweihu@purdue.edu}

\author[mymainaddress]{Zheng Ma}
% \cortext[mycorrespondingauthor]{Corresponding author}
\ead{ma531@purdue.edu}

\address[mymainaddress]{Department of Mathematics, Purdue University\\150 N. University Street, West Lafayette, IN 47907, USA}
% \address[mysecondaryaddress]{360 Park Avenue South, New York}

\begin{abstract}
In this paper, we propose a simple fast Fourier spectral method for the inelastic Boltzmann collision operator, with its application to one of the widely used models of granular gases, the inelastic Boltzmann equation with a heating source. Compared to the direct Fourier spectral method, our fast algorithm reduces the computational complexity from $O\left(N^6\right)$ to $O\left(MN^4\log N \right)$ per evaluation of the collision operator in three dimensions, where $N$ is the number of discretization points in each velocity dimension and $M \ll N^2$ is the number of quadrature points used on the unit sphere. We test the numerical accuracy and efficiency of the proposed method in both two dimensional and three dimensional examples, where in the latter case the famous Haff's cooling law for granular flows is successfully recovered.
\end{abstract}

\begin{keyword}
inelastic collision operator \sep inelastic Boltzmann equation with a heating source  \sep granular gas \sep  Haff's cooling law \sep fast Fourier spectral method \sep spherical design 
% \texttt{elsarticle.cls}\sep \LaTeX\sep Elsevier \sep template
% \MSC[2010] 00-01\sep  99-00
\end{keyword}

\end{frontmatter}

\linenumbers

\section{Introduction}

It has been found in the past few decades that the granular gases behave fundamentally different from the usual molecular gases modeled as elastically colliding spheres. The rich phenomenon of such systems, such as the formation of clusters and shear instability, draws a lot of attention from both theoretical and industrial application point of view. Different from their molecular counterparts, granular gases allow inelastic collisions, in other words, break the time-reversible symmetry because of the energy dissipation. Despite this dissipative nature and its resulting nontrivial properties, the basic equation of kinetic theory, the Boltzmann equation, can still be extended to describe the granular gases with a different collision operator \cite{BP04}, namely,
\begin{equation}
\partial_t f + v \cdot \nabla_x f = Q_{\text{in}}(f,f),
\label{boltz}
\end{equation}
where $f=f(t,x,v)$ is the one-particle distribution function depending on time $t \geq 0$, position $x \in \mathbb{R}^d$, and particle velocity $v \in \mathbb{R}^d$ with the dimension $d \geq 1$ ($d$ should be 3 in practice, but for simplicity $d=1$ or $2$ are also considered), and $Q_{\text{in}}$ is the so-called inelastic or granular collision operator --- a nonlinear (quadratic) integral operator modeling binary collisions among particles whose exact expression will be presented in later discussions. For convenience, we use $Q$ instead of $Q_{\text{in}}$ in the rest of this paper, and the notation should not be confused with the usual elastic collision operator \cite{Cercignani}. Another widely used model, first introduced by van Noije and Ernst \cite{noije}, is the spatially homogeneous inelastic Boltzmann equation with a heating source:
\begin{equation} \label{inBoltz}
\partial_tf-\varepsilon \Delta_vf=Q(f,f),
\end{equation}
where the distribution function $f$ depends only on the time $t$ and velocity $v$, and the term $\varepsilon \Delta_vf$ represents the diffusion effects with the diffusion coefficient $\varepsilon \ll 1$, incurred by a heat bath of infinite temperature. The rationale to consider this model is that many experiments about granular material include shaking, as a way to input energy into the system, counterbalancing the freezing due to energy loss.

In spite of the wide engineering and industrial applications of granular gases, the mathematical theory behind the inelastic Boltzmann equation is still at an early stage and most of the theoretical studies are retained to the spatially homogeneous case. We refer to the recent review by Villani \cite{Villani2006} for a collection of rigorous mathematical results and open questions on this topic. Numerically, mainly two classes of methods have been considered for the inelastic Boltzmann equation, one stochastic and one deterministic. In \cite{MS00, GRW05, RW07}, the direct simulation Monte Carlo (DSMC) method, initially proposed for solving the classical Boltzmann equation \cite{Bird}, is modified and extended to the inelastic case. Due to its simplicity and efficiency, DSMC still remains the most viable choice for large scale kinetic simulations. On the other hand, the deterministic method, in particular, the spectral method based on Fourier series/transform has become quite popular in the past decade thanks to the advances in computing power. Without being exhaustive, we refer to \cite{DP14} and references therein for a general discussion of such methods. For inelastic collision operators, the first attempt was made in \cite{NPT03} for the one dimensional model; later in \cite{FPT05, GT09, FR13}, both two and three dimensional cases were considered by directly generalizing the spectral method for the classical Boltzmann equation. Although the spectral method can produce very accurate results in comparison to DSMC, its enormous computational cost and excessive storage requirement present a huge challenge in practice. Indeed, all the above mentioned methods require $O(N^{2d})$ complexity per evaluation of the collision operator and the same amount of memory to store the precomputed weights, where $N$ is the number of discretization points in each velocity dimension. This quickly becomes a bottleneck as $N$ increases especially for physically relevant three dimensional problems.

{\bf Our contribution.} In this paper, we propose a simple fast algorithm for the inelastic Boltzmann collision operator. The method is based on the same framework as the direct Fourier spectral method \cite{FPT05}. By exploiting the low rank and convolutional structure in the collision integral, we are able to reduce the computational cost from $O(N^{2d})$ to $O(MN^{d+1}\log N)$, where $M \ll N^{d-1}$ is the number of quadrature points used on the unit circle (in 2D) or sphere (in 3D). No precomputation is needed for the variable hard sphere model (a commonly used collision kernel including Maxwell molecules and hard spheres as subcases), and only $O(MN^{d+1})$ memory is required to store precomputed weights for more general kernels. The idea is largely inspired by the recent work \cite{GHHH17}, where instead of precomputing all the weights we approximate them partially ``on the fly" using a quadrature rule. We carefully validate the accuracy and efficiency of the proposed method using a series of examples. In particular, we find that separate approximation of the gain and loss terms of the collision operator yields better accuracy compared to approximating them as a whole, although the latter seems quite natural starting from the weak formulation of the equation; we also numerically verify the famous Haff's cooling law for 3D inelastic hard spheres, which says that the temperature in a granular gas decays like $O(t^{-2})$ with time, as predicted in Haff's seminal work \cite{Haff83}.

{\bf Related work.} A closely related work is the fast Fourier spectral method introduced in \cite{WZR15} for the generalized Enskog equation (the Boltzmann and Enskog collision operators reduce to the same thing in the spatially homogeneous case). This method is based on a different representation of the collision operator, the so-called Carleman form, and relies on an extra approximation of the integral in the radial direction. As a result, the final computational cost is comparable to our method. However, we point out that the method in \cite{WZR15} only readily works for 2D Maxwell molecules and 3D hard spheres. To deal with other types of interactions (even 3D Maxwell molecules, say), additional modification of the kernel or parameter fitting is necessary, as was done in \cite{WWSRZ13}. In contrast, our method makes no assumption in the collision kernel and is very easy to implement. Further, we propose to use the spherical design \cite{Womersley} to approximate the integral on the unit sphere, which is the optimal quadrature on the sphere up to date and shows better performance (require much fewer points to attain the same level of accuracy) than the tensor product Gauss quadrature as used in \cite{WZR15}.

The rest of the paper is organized as follows: Section 2 provides a brief overview of the inelastic collision operator, and the heated inelastic Boltzmann equation widely used in practice for modeling granular flows. Section 3 presents the details of the fast Fourier spectral method for the inelastic collision operator. In Section 4, we perform a series of numerical tests in both 2D and 3D to validate the accuracy and efficiency of the proposed method. The paper is concluded in Section 5.

\section{The inelastic collision operator and the inelastic Boltzmann equation with a heating source}

\subsection{The inelastic collision operator -- strong and weak formulations}

Unlike the classical Boltzmann collision operator \cite{Cercignani} where both the momentum and energy are conservative, the mathematical expression of the inelastic collision operator $Q(f,f)$ is more complicated due to the energy dissipation, and there is a lot of ambiguity in the existing literature especially related to its strong formulation. To clarify and fix the notation, we introduce in this subsection both the strong and weak forms of the inelastic collision operator. Our presentation is consistent to \cite{CCC09} where a more thorough derivation can be found.

Assume two particles with velocities $v$ and $v_*$ are going to collide. During the collision, there is some loss of momentum in the impact direction $\omega \in S^{d-1}$ ($S^{d-1}$ is the unit sphere in $\mathbb{R}^d$), resulting in the post-collisional velocities $v'$ and $v_*'$. Let $e$ stand for the restitution coefficient or inelasticity parameter, then
\begin{equation} \label{IO}
(v'-v_*')\cdot \omega=-e[(v-v_*)\cdot \omega], \quad 0 \leq e \leq 1.
\end{equation}
Note that the degree of inelasticity is indicated by $e$, where $e=1$ corresponds to perfectly elastic collisions (no loss of energy) and $e=0$ to sticky collisions (particles travel together after a collision). $e$ could be a constant, or may depend on the impact velocity $|(v-v_*)\cdot \omega|$ for viscoelastic particles \cite{BP}.

Using (\ref{IO}), $v'$ and $v_*'$ can be represented as
\begin{align}\label{omega}
\left\{
\begin{array}{l}
\displaystyle v'=v-\frac{1+e}{2}[(v-v_*)\cdot \omega ]\omega, \\[8pt]
\displaystyle v_*'=v_*+\frac{1+e}{2}[(v-v_*)\cdot \omega]\omega.
\end{array}\right.
\end{align}
Following Villani \cite{Villani02}, we shall call this $\omega$-representation. From (\ref{omega}) one can easily verify the conservation of momentum and dissipation of energy:
\begin{equation} \label{momentum}
v'+v_*'=v+v_*; \quad |v'|^2+|v_*'|^2-|v|^2-|v_*|^2=-\frac{1-e^2}{2}[(v-v_*)\cdot \omega]^2\leq 0.
\end{equation}

To write down the strong form of the collision operator, one needs to define the pre-collisional velocities $\tv$ and $\tv_*$ that produce velocities $v$ and $v_*$ after the collision. They follow the same rule as given in (\ref{omega}), with $(v,v_*)$ replaced by $(\tv, \tv_*)$ and $(v',v_*')$ replaced by $(v,v_*)$. Note that $\tv$ and $\tv_*$ do not coincide with $v'$ and $v_*'$ unlike the classical (elastic) case; in other words, the collisions are not reversible.

Now the inelastic Boltzmann collision operator (in strong form) can be written as
\begin{align} \label{strong}
Q(f,f)(v)&=\int_{\mathbb{R}^d}\int_{S^{d-1}}B_{\omega}(|\tg|,|\omega\cdot  \hat{\tg}|) \tf\tf_*\,J\,\rd{\omega}\rd{v_*}-\int_{\mathbb{R}^d}\int_{S^{d-1}}B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_*\,\rd{\omega}\rd{v_*}\nonumber\\
:&=Q^+(f,f)-Q^-(f,f),
\end{align}
which naturally splits into a gain term $Q^+$ and a loss term $Q^-$. In (\ref{strong}), $g:=v-v_*$ is the relative velocity, $\hat{g}$ is the unit vector along $g$, and we used the shorthand notation $\tg=\tv-\tv_*$, $\tf=f(\tv)$, $\tf_*=f(\tv_*)$ and so on. $J$ is the determinant of the Jacobian from $(v,v_*)$ to $(\tv,\tv_*)$, i.e.,
\begin{equation}
J=\left |  \frac{\partial (\tv,\tv_*)}{\partial (v,v_*)}     \right|.
\end{equation}
$B_{\omega}=B_{\omega}(|g|,|\omega\cdot \hat{g}|)$ is the collision kernel depending only on $|g|$ and $|\omega\cdot \hat{g}|$, whose specific form can be determined from the interaction potential using scattering theory. %In particular, $B_{\omega}=2|\omega \cdot g|$ for 3D hard spheres.

Starting from the strong form (\ref{strong}) of the collision operator, one can derive various weak forms that are more often used when studying $f$ via physical observables. For a test function $\varphi(v)$, we have
\begin{align}
\int_{\mathbb{R}^d} Q(f,f)(v)\,\varphi(v)\,\rd{v}=&\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\int_{S^{d-1}}B_{\omega}(|\tg|,|\omega\cdot \hat{\tg}|)\tf \tf_* \,J\,\varphi\,\rd{\omega}\rd{v_*}\rd{v}\nonumber\\
&  -\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\int_{S^{d-1}}B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_*\varphi\,\rd{\omega}\rd{v_*}\rd{v}\nonumber\\
=&\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\int_{S^{d-1}}B_{\omega}(|\tg|,|\omega\cdot \hat{\tg}|)\tf \tf_* \varphi\,\rd{\omega}\rd{\tv_*}\rd{\tv}\nonumber\\
& -\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\int_{S^{d-1}}B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_*\varphi\,\rd{\omega}\rd{v_*}\rd{v}\nonumber\\
=&\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\int_{S^{d-1}}B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_* \varphi'\,\rd{\omega}\rd{v_*}\rd{v}\nonumber\\
& -\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\int_{S^{d-1}}B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_*\varphi\,\rd{\omega}\rd{v_*}\rd{v},
\end{align}
where in the second equality we changed $\rd{v_*}\rd{v}$ to $\rd{\tv_*}\rd{\tv}$ in the gain term and in the third equality we changed $(\tv,\tv_*)$ to $(v,v_*)$ for a fixed $\omega$, hence $(v,v_*)$ becomes $(v',v_*')$. Therefore, we have
\begin{align} \label{weak1}
\int_{\mathbb{R}^d} Q(f,f)(v)\,\varphi(v)\,\rd{v}&= \int_{\mathbb{R}^d} \int_{\mathbb{R}^d} \int_{S^{d-1}} B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_*  \left(\varphi'-\varphi\right)\,\rd{\omega} \rd{v} \rd{v_*}.
\end{align}
In (\ref{weak1}), if we swap $v$ and $v_*$ for a fixed $\omega$, then $v'$ is swapped with $v_*'$ due to (\ref{omega}):
\begin{align}
\int_{\mathbb{R}^d} Q(f,f)(v)\,\varphi(v)\,\rd{v}&= \int_{\mathbb{R}^d} \int_{\mathbb{R}^d} \int_{S^{d-1}} B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_*  \left(\varphi_*'-\varphi_*\right)\,\rd{\omega} \rd{v} \rd{v_*}.
\end{align}
Averaging the previous two steps, we obtain another weak form
\begin{align} \label{weak2}
\int_{\mathbb{R}^d} Q(f,f)(v)\,\varphi(v)\,\rd{v}&=\frac{1}{2} \int_{\mathbb{R}^d} \int_{\mathbb{R}^d} \int_{S^{d-1}} B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_* \left(\varphi'+\varphi_*'-\varphi-\varphi_* \right)\,\rd{\omega}\rd{v} \rd{v_*}.
\end{align}

As we have seen above, the $\omega$-representation is very easy to perform the change of variables, hence to obtain the weak forms. For numerical purpose, it is convenient to consider a different parametrization, the so-called $\sigma$-representation (again following the terminology in \cite{Villani02}):
\begin{align} \label{sigma}
\left\{
\begin{array}{l}
\displaystyle v'=\frac{v+v_*}{2}+\frac{1-e}{4}(v-v_*)+\frac{1+e}{4}|v-v_*|\sigma, \\[8pt]
\displaystyle v_*'=\frac{v+v_*}{2}- \frac{1-e}{4}(v-v_*)-\frac{1+e}{4}|v-v_*|\sigma,
\end{array}\right.
\end{align}
where $\sigma$ is another unit vector on $S^{d-1}$ and is related to $\omega$ as
\begin{equation} \label{relation}
(g\cdot \omega)\omega=\frac{1}{2}(g-|g|\sigma).
\end{equation}

It can be verified that the transformation between $\omega$ and $\sigma$ representations in the integral form is given by
\begin{align}
\int_{\mathbb{R}^d}\int_{S^{d-1}}B_{\omega}(|g|,|\omega\cdot \hat{g}|)\cdot \rd{\omega}\rd{v_*}=\int_{\mathbb{R}^d}\int_{S^{d-1}}B_{\sigma}(|g|,\sigma\cdot \hat{g})\cdot \rd{\sigma}\rd{v_*},
\end{align}
with 
\begin{align}
B_{\omega}(|g|, |\omega\cdot \hat{g}|) =|2(\omega\cdot \hat{g})|^{d-2}B_{\sigma}(|g|,1-2(\omega\cdot\hat{g})^2).
\end{align}
Therefore, the two weak forms (\ref{weak1}) and (\ref{weak2}) written in the $\sigma$-representation read as
{\bf Weak form 1:}
\begin{align} \label{weak11}
\int_{\mathbb{R}^d} Q(f,f)(v)\,\varphi(v)\,\rd{v}&= \int_{\mathbb{R}^d} \int_{\mathbb{R}^d} \int_{S^{d-1}} B_{\sigma}(|g|,\sigma\cdot \hat{g})ff_*  \left(\varphi'-\varphi\right)\,\rd{\sigma} \rd{v} \rd{v_*},
\end{align}
{\bf Weak form 2:}
\begin{align} \label{weak22}
\int_{\mathbb{R}^d} Q(f,f)(v)\,\varphi(v)\,\rd{v}&=\frac{1}{2} \int_{\mathbb{R}^d} \int_{\mathbb{R}^d} \int_{S^{d-1}} B_{\sigma}(|g|,\sigma\cdot \hat{g})ff_*\left(\varphi'+\varphi_*'-\varphi-\varphi_* \right)\,\rd{\sigma} \rd{v} \rd{v_*},
\end{align}
where $g=v-v_*$, and $(v',v_*')$ are given by (\ref{sigma}). The collision kernel $B_{\sigma}$ may take various forms depending on the types of interactions. The most commonly used form is the variable hard sphere (VHS) model \cite{Bird}:
\begin{equation} \label{VHS}
B_{\sigma}(|g|,\sigma\cdot \hat{g})=C_{\lambda}|g|^{\lambda}, \quad 0\leq \lambda \leq 1,
\end{equation}
where $C_{\lambda}> 0$ is some constant. Two special cases are: Maxwell molecules ($\lambda=0$) and hard spheres ($\lambda=1$).

In the following discussions, we shall exclusively use the weak forms (\ref{weak11}) and (\ref{weak22}).

\subsection{The inelastic Boltzmann equation with a heating source}

In this subsection, we consider the inelastic Boltzmann equation with a heating source \eqref{inBoltz}. We shall revisit a few properties of this equation.

First of all, we define the following moments of $f$:
\begin{equation}
\rho=\int_{\mathbb{R}^d}f\,\rd{v}, \quad m=\int_{\mathbb{R}^d}fv\,\rd{v}, \quad E=\frac{1}{2}\int_{\mathbb{R}^d}f |v|^2\,\rd{v},
\end{equation}
where $\rho$, $m$ and $E$ are, respectively, the density, momentum and energy. We further define the bulk velocity $u$ and temperature $T$ as
\begin{equation}
u=\frac{1}{\rho}\int_{\mathbb{R}^d} fv \,\rd{v}, \quad  T = \frac{1}{d\rho }\int_{\mathbb{R}^d}f|v - u|^2 \,\rd{v},
\end{equation}
then
\begin{equation}
m=\rho u, \quad E=\frac{1}{2}\rho u^2+\frac{d}{2}\rho T.
\end{equation}

Now assuming $f$ is a solution to equation \eqref{inBoltz}, the conservation of mass and momentum is straightforward by taking $\varphi = 1$ and $v$ respectively in the weak form \eqref{weak22}, i.e.,
\begin{equation}
\rho \equiv \rho_0, \quad  m\equiv m_0,
\end{equation}
with $\rho_0$ and $m_0$ being the initial density and momentum. By taking $\varphi=|v|^2/2$ in \eqref{weak22}, one obtains the following energy equation
\begin{equation}\label{ki_energy}
\partial_t E -\varepsilon d\rho_0=-\frac{1}{16}\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\int_{S^{d-1}} (1-e^2)B_{\sigma}(|g|,\sigma\cdot \hat{g})|g|^2(1-\sigma\cdot \hat{g})ff_*\,\rd{\sigma}\rd{v}\rd{v_*},
\end{equation}
where we used (\ref{momentum}) and (\ref{relation}).

For constant restitution coefficient $e$ and Maxwell molecules ($\lambda=0$ in (\ref{VHS})), equation \eqref{ki_energy} becomes
\begin{equation} 
\partial_t E -\varepsilon d\rho_0=-\frac{1-e^2}{16}C_0\int_{\mathbb{R}^d}\int_{\mathbb{R}^d}\int_{S^{d-1}}  |g|^{2}(1-\sigma\cdot \hat{g})ff_*\,\rd{\sigma}\rd{v}\rd{v_*}.
\end{equation}
Note that
\begin{equation}
\int_{S^{d-1}} C_0(1-\sigma\cdot \hat{g})\,\rd{\sigma}
\end{equation}
is a positive constant regardless of $\hat{g}$ and, without loss of generality, is assumed to be $1$. Then
\begin{align} 
  \partial_t E -\varepsilon d\rho_0 &=-\frac{1-e^2}{16}\int_{\mathbb{R}^d}\int_{\mathbb{R}^d} |g|^{2}ff_*\,\rd{v}\rd{v_*} \notag\\
   & =-\frac{1-e^2}{16}\int_{\mathbb{R}^d}\int_{\mathbb{R}^d} (|v|^2+|v_*|^2-2v\cdot v_*)ff_*\,\rd{v}\rd{v_*} \notag \\
   &=-\frac{1-e^2}{8}\left(2\rho_0E - m_0^2\right).
\end{align}

For simplicity, if we take the initial condition such that $\rho_0=1$, $u_0=0$, then $T=\frac{2}{d}E$ satisfies
\begin{equation}
\partial_t T-2\varepsilon =-\frac{1-e^2}{4}T,
\end{equation}
whose solution is
\begin{equation}\label{soln:T}
T=\left(T_0-\frac{8\varepsilon}{1-e^2}\right)\exp{\left(-\frac{1-e^2}{4}t\right)}+\frac{8\varepsilon}{1-e^2}.
\end{equation}

\begin{rmk}
This explicit expression will serve as the reference quantity in our later numerical tests, since in the inelastic case it is notably harder to find an analytical $f$ satisfying the Boltzmann equation (in fact there is no such $f$ to the best of our knowledge) in contrast to the elastic case where a BKW solution is available.
\end{rmk}

\section{A fast spectral method for the inelastic Boltzmann collision operator}

In this section, we describe in detail the fast spectral method for the inelastic Boltzmann collision operator. Since our method aims to directly accelerate the Fourier-Galerkin spectral method proposed in \cite{FPT05}, we first lay out the basic framework of this original method and then introduce our new idea.

The starting point is the weak form (\ref{weak11}) with a change of integration variable $v_* \rightarrow g=v-v_*$:
\begin{align} 
\int_{\mathbb{R}^d} Q(f,f)(v)\,\varphi(v)\,\rd{v}&= \int_{\mathbb{R}^d} \int_{\mathbb{R}^d} \int_{S^{d-1}} B_{\sigma}(|g|,\sigma\cdot \hat{g})f(v)f(v-g)  \left(\varphi(v')-\varphi(v)\right)\,\rd{\sigma} \rd{g} \rd{v},
\end{align}
where
\begin{equation}
v'=v-\frac{1+e}{4}(g-|g|\sigma).
\end{equation}
It is then assumed that $f$ has a compact support: $\text{Supp}_v(f) \approx B_S$, where $B_S$ is a ball centered at the origin with radius $S$ (in practice, the support can be chosen as some multiple of the initial thermal speed $v_{\text{th}}=\sqrt{T_0}$). Hence it suffices to truncate the infinite integral in $g$ to a larger ball $B_R$ with radius $R=2S$:
\begin{align}  \label{weak}
\int_{\mathbb{R}^d} Q(f,f)(v)\,\varphi(v)\,\rd{v}&= \int_{\mathbb{R}^d} \int_{B_R} \int_{S^{d-1}} B_{\sigma}(|g|,\sigma\cdot \hat{g})f(v)f(v-g)  \left(\varphi(v')-\varphi(v)\right)\,\rd{\sigma} \rd{g} \rd{v}.
\end{align}
Next, we restrict $v$ to a cubic computational domain $D_L=[-L,L]^d$, and approximate $f$ by a truncated Fourier series:
\begin{equation}
f(v)\approx f_N(v)=\sum_{k=-\frac{N}{2}}^{\frac{N}{2}-1}\hat{f}_k e^{i\frac{\pi}{L}k\cdot v}, \quad \hat{f}_k=\frac{1}{(2L)^d}\int_{D_L}f(v)e^{-i\frac{\pi}{L}k\cdot v}\,\rd{v}.
\end{equation}
Note that $k=(k_1,\dots,k_d)$ is a multidimensional index, and $\sum_{k=-N/2}^{N/2-1}:=\sum_{k_1,\dots,k_d=-N/2}^{N/2-1}$. The choice of $L$ can be made similarly as in the classical case $L\geq (3+\sqrt{2})S/2$, see \cite{PR00} for more details. Now substituting $f_N$ into (\ref{weak}) and choosing $\varphi(v)=e^{-i\frac{\pi}{L}k\cdot v}$, we can obtain the $k$-th mode of the collision operator as
\begin{align} \label{QQ}
\hat{Q}_k:&=\frac{1}{(2L)^d}\int_{D_L}Q(f_N,f_N)(v)e^{-i\frac{\pi}{L}k\cdot v}\,\rd{v}\nonumber\\
&=\frac{1}{(2L)^d}\int_{D_L} \int_{B_R} \int_{S^{d-1}} B_{\sigma}(|g|,\sigma\cdot \hat{g})f_N(v)f_N(v-g)  \left(e^{-i\frac{\pi}{L}k\cdot v'}-e^{-i\frac{\pi}{L}k\cdot v}\right)\,\rd{\sigma} \rd{g} \rd{v}\nonumber\\
&=\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\int_{B_R}\int_{S^{d-1}}B_{\sigma}(|g|,\sigma\cdot \hat{g})e^{-i\frac{\pi}{L}m\cdot g}\left(e^{i\frac{\pi}{L}\frac{1+e}{4}k\cdot (g-|g|\sigma)}-1\right)\hat{f}_l\hat{f}_m\,\rd{\sigma} \rd{g},
\end{align}
where in the last equality we used the orthogonality of the Fourier basis.

Define the weight
\begin{equation} \label{weight}
G(l,m)=\int_{B_R}e^{-i\frac{\pi}{L}m \cdot g}\left[\int_{S^{d-1}}B_{\sigma}(|g|,\sigma\cdot \hat{g})\left(e^{i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot (g-|g|\sigma)}-1\right)\,\rd{\sigma}\right]\,\rd{g},
\end{equation}
then (\ref{QQ}) can be written as
\begin{equation} \label{sum}
\hat{Q}_k=\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}G(l,m)\hat{f}_l\hat{f}_m.
\end{equation} 
In the original spectral method \cite{FPT05}, the weight $G(l,m)$ is precomputed and stored as it is independent of the solution $f$ (memory requirement $O(N^{2d})$). During the online computation, the weighted sum (\ref{sum}) is directly evaluated whose complexity is $O(N^{2d})$.

To reduce the computational cost, we would like to render the weighted convolution (\ref{sum}) into a pure convolution so that it can be computed efficiently by the fast Fourier transform (FFT). Inspired by \cite{GHHH17}, here we propose to decompose the weight $G(l,m)$ as follows:
\begin{equation} \label{lowrank}
G(l,m)\approx\sum_{p=1}^{N_p}\alpha_p(l+m)\beta_p(m),
\end{equation}
where $\alpha_p$ and $\beta_p$ are some functions to be determined and the number of terms $N_p$ in the expansion is small. Then (\ref{sum}) becomes
\begin{equation}
\hat{Q}_k\approx \sum_{p=1}^{N_p}\alpha_p(k)\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l  \left(\beta_p(m)\hat{f}_m\right),
\end{equation} 
where the inner summation is a pure convolution of two functions $\hat{f}_l$ and $\beta_p(m)\hat{f}_m$. Hence the total complexity to evaluate $\hat{Q}_k$ (for all $k$) is brought down to $O(N_pN^d\log N)$, i.e., a couple of FFTs.

A close inspection of (\ref{weight}) suggests that it is not difficult to generate an expansion of the form (\ref{lowrank}). The key is to approximate the integral in $g$ using a fixed quadrature rule. Specifically, we only precompute the part inside the bracket
\begin{equation} \label{FF}
F(l+m,\rho,\hat{g}):=\int_{S^{d-1}}B_{\sigma}(\rho,\sigma\cdot \hat{g})\left(e^{i\frac{\pi}{L}\rho \frac{1+e}{4}(l+m)\cdot (\hat{g}-\sigma)}-1\right)\,\rd{\sigma},
\end{equation}
and approximate $G(l,m)$ ``on the fly" as
\begin{equation} \label{lowrank1}
G(l,m)\approx \sum_{\rho, \hat{g}}w_{\rho}w_{\hat{g}}\, \rho^{d-1} e^{-i\frac{\pi}{L}\rho\, m \cdot \hat{g}} F(l+m,\rho,\hat{g}),
\end{equation}
where the integral in $g$ is split into a radial part $\rho:=|g|\in [0,R]$ and a spherical part $\hat{g}\in S^{d-1}$, and $w_{\rho}$ and $w_{\hat{g}}$ are the corresponding quadrature weights. The radial direction $\rho$ can be approximated by the Gauss-Legendre quadrature. Since the integrand in (\ref{lowrank1}) is oscillatory on the scale of $O(N)$, the number of quadrature points needed for $\rho$ should be $O(N)$. The spherical direction $\hat{g}$ can be treated as follows: in 2D, the integrand is $2\pi$-periodic, so the simple rectangular rule is expected to yield spectral accuracy; in 3D, we propose to use the spherical design \cite{Womersley}, which is the optimal quadrature on the sphere up to date and performs much better than the tensor product based Gauss quadrature \cite{Beentjes15}. It should be pointed out that it is hard to obtain a priori estimate of the number of points $M$ needed on $S^2$, but numerical experiments show that typically $M\ll N^2$, see Section 4.

In summary, we have obtained the decomposition (\ref{lowrank1}), which is in the desired form (\ref{lowrank}) with $N_p=O(MN)$. So the total cost to evaluate $\hat{Q}_k$ (for all $k$) is reduced to $O(MN^{d+1}\log N)$. Furthermore, the only quantity that needs to be precomputed and stored is (\ref{FF}), which in the worst scenario requires $O(MN^{d+1})$ memory. In fact, no precomputation is needed for the VHS model (\ref{VHS}), as will be shown below.

The above formulation treats the gain term and loss term (positive and negative parts) in (\ref{weight}) in a unified way. Note that the loss term is readily a function of $m$, hence no approximation/decomposition is actually needed. This suggests an alternative way to evaluate the loss term of the collision operator: precompute
\begin{equation}
G(m):=\int_{B_R}e^{-i\frac{\pi}{L}m \cdot g}\left[\int_{S^{d-1}}B_{\sigma}(|g|,\sigma\cdot \hat{g})\,\rd{\sigma}\right]\,\rd{g},
\end{equation}
and compute
\begin{equation}
\hat{Q}^-_k=\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l\left(G(m)\hat{f}_m\right)
\end{equation}
by FFT. In our numerical experiments, we test both approaches carefully. Surprisingly, even it seems quite natural to approximate gain and loss terms together starting from the weak formulation, approximating them separately yields more accurate results. For details, see Section 4.

\begin{rmk}
The fast algorithm proposed here works for both constant and non-constant restitution coefficient. If $e$ is a function of impact velocity (\cite{BP}), i.e., $e=e(|g\cdot \omega|)=e(|g-|g|\sigma|/2)$, it merely complicates the precomputation of $F$, everything else remains the same.
\end{rmk}

To close this section, we give explicitly the formulas for the VHS model (\ref{VHS}) in both 2D and 3D. These will be used in Section 4 to validate our proposed method.

\subsection{2D VHS model}

For the VHS model (\ref{VHS}) in 2D, we have
\begin{align}
 F(k,\rho,\hat{g})&= \int_{S^1}C_{\lambda}\,\rho^{\lambda}\left(e^{i\frac{\pi}{L} \rho \frac{1+e}{4}k\cdot (\hat{g}-\sigma)}-1\right)\,\rd{\sigma}\nonumber\\
 &=2\pi \,C_{\lambda}\,\rho^{\lambda}\left[ e^{i\frac{\pi}{L}\rho \frac{1+e}{4}k\cdot \hat{g}}\,\text{J}_0\left(\frac{\pi}{L}\rho \frac{1+e}{4}|k|\right)-1\right],
\end{align}
where $\text{J}_0$ is zero-th order Bessel function of the first kind. Therefore,
\begin{equation} \label{full}
  \hat{Q}_k\approx\sum_{\rho,\hat{g}}w_{\rho}w_{\hat{g}}\,\rho \,F(k,\rho,\hat{g})\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[e^{-i\frac{\pi}{L}\rho\, m \cdot \hat{g}}\hat{f}_m\right],
\end{equation} 
or the loss term can be computed separately as
\begin{equation} \label{separate}
\hat{Q}_k^-\approx\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[\sum_{\rho}w_{\rho}\,4\pi^2\, C_{\lambda}\,\rho^{\lambda+1}\, \text{J}_0\left(\frac{\pi}{L}\rho |m|\right)\hat{f}_m\right].
\end{equation}

\subsection{3D VHS model}

For the VHS model (\ref{VHS}) in 3D, we have
\begin{align}
  F(k,\rho,\hat{g})&=\int_{S^2}C_{\lambda}\,\rho^{\lambda}\left(e^{i\frac{\pi}{L}\rho \frac{1+e}{4}k\cdot (\hat{g}-\sigma)}-1\right)\,\rd{\sigma}\nonumber\\
  &=4\pi \,C_{\lambda}\,\rho^{\lambda}\left[ e^{i\frac{\pi}{L}\rho \frac{1+e}{4}k\cdot \hat{g}}\,\text{Sinc}\left(\frac{\pi}{L}\rho\frac{1+e}{4}|k|\right)-1\right],
\end{align}
where $\text{Sinc}(x)=\sin x/x$ is the Sinc function. Therefore,
\begin{equation} 
\hat{Q}_k\approx\sum_{\rho,\hat{g}}w_{\rho}w_{\hat{g}}\, \rho^2\,F(k,\rho,\hat{g})\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[e^{-i\frac{\pi}{L}\rho \,m \cdot \hat{g}}\hat{f}_m\right],
\end{equation} 
or the loss term can be computed separately as
\begin{equation} \label{separate3}
\hat{Q}_k^-\approx\sum_{\substack{l,m=-\frac{N}{2}\\l+m=k}}^{\frac{N}{2}-1}\hat{f}_l \left[\sum_{\rho}w_{\rho}\,16\pi^2 \,C_{\lambda}\,\rho^{\lambda+2} \,\text{Sinc}\left(\frac{\pi}{L}\rho |m|\right)\hat{f}_m\right].
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Numerical examples}

In this section, we verify the accuracy and efficiency of the proposed method with extensive numerical studies in both two and three dimensions. As mentioned previously, unlike the elastic collision operator where an analytical solution $f$ is available, the inelastic case, however, only has analytical formulas for macroscopic quantities, such as the temperature $T$ as shown in \eqref{soln:T}. This means that in order to provide a trustworthy verification of our solver, one needs an accurate scheme to solve the inelastic Boltzmann equation \eqref{inBoltz}.

In the following numerical simulations, we apply our fast spectral solver to evaluate the collision operator, paired with a standard Fourier spectral method to treat the diffusion term $\varepsilon\Delta_v f$ and a third-order Runge-Kutta scheme for time discretization. For the ODE system $\partial_t f=L(f)$, the RK3 scheme we use is given by
\begin{align}
  k_1 &= L(f^n), \notag \\
  k_2 &= L(f^n + \frac{1}{2}k_1\Delta t), \notag \\
  k_3 &= L(f^n - k_1\Delta t + 2k_2\Delta t), \notag \\
  f^{n+1} &= f^n + \frac{1}{6}(k_1 + 4k_2 + k_3)\Delta t.
\end{align}
We always set $\varepsilon=10^{-6}$ for the diffusion term except the last example where $\varepsilon=0$.

\subsection{2D examples}

In this subsection, we perform several 2D examples for Maxwell molecules, i.e., $B_{\sigma}=\frac{1}{2\pi}$ in (\ref{VHS}). Since it is clear that the simple rectangular rule on the circle will yield spectral accuracy, our focus here is to verify the spectral accuracy in $N$. We will consider two variants, one is to approximate the gain and loss terms together as in (\ref{full}), which we shall denote by ``full" method below; the other is to approximate them separately with the loss term given by (\ref{separate}), which we shall denote by ``separate" method.

\subsubsection{2D inelastic Maxwell molecules --- isotropic solution}

We first consider the isotropic initial condition:
\begin{equation} \label{ext1}
  f_0(v) = \frac{1}{2\pi K^2}\exp\left(-\frac{v^2}{2K}\right)\left(2K-1+\frac{1-K}{2K}v^2\right),
\end{equation}
where $K=1-\exp(-1/16)/2$. One can easily check that $\rho_0 = 1$, $u_0 = 0$ and $T_0 =1$.

We compare the difference of temperature $|T_\text{num} - T_\text{ref}|$ at some given time $t_\text{final}$. The numerical $T_\text{num}$ is obtained by taking the moments of the numerical solution $f_\text{num}$, computed by our fast spectral method. The reference $T_\text{ref}$ is obtained using the exact formula \eqref{soln:T}.

In order to minimize the error due to time discretization, we first perform a convergence test in time of the RK3 scheme. In Figure~\ref{dt_conv_1} we plot the errors versus different $\Delta t$ at $t_\text{final}=2$. Various values of the restitution coefficient $e = 0.2, 0.5$ and $0.8$, and both the ``full" and ``separate" methods are considered. We choose a relatively large $N=64$ and $N_{\rho}=M_{\text{cir}}=30$, where $N$ is the number of discretization points in each velocity dimension, $N_{\rho}$ is the Gauss-Legendre points used in radial direction, and $M_{\text{cir}}$ is the (uniform) quadrature points used on the circle. A third order convergence rate can be clearly observed. Since in this test we fix all the parameters in our spectral method as $\Delta t$ changes, these results also imply that the error of the spectral method can get very small, or at least around $O(10^{-9})$.

\begin{figure}[htp!]
  \centering
   \subfloat[$e = 0.2$, full]{\includegraphics[width = 0.5\linewidth]{dt_2d_bkw_e_full=02}} \hfill
    \subfloat[$e = 0.2$, separate]{\includegraphics[width = 0.5\linewidth]{dt_2d_bkw_e=02}}\\
   \subfloat[$e = 0.5$, full]{\includegraphics[width = 0.5\linewidth]{dt_2d_bkw_e_full=05}} \hfill
    \subfloat[$e = 0.5$, separate]{\includegraphics[width = 0.5\linewidth]{dt_2d_bkw_e=05}}\\
  \subfloat[$e = 0.8$, full]{\includegraphics[width = 0.5\linewidth]{dt_2d_bkw_e_full=08}} \hfill
    \subfloat[$e = 0.8$, separate]{\includegraphics[width = 0.5\linewidth]{dt_2d_bkw_e=08}}
  \caption{Convergence test in time of the 2D Maxwell molecule (isotropic solution) for $e=0.2$, $0.5$ and $0.8$, and for both the ``full" and ``separate" methods. Errors shown here are $|T_\text{num} - T_\text{ref}|$ at $t_\text{final}=2$. $N=64$, $N_{\rho}=30$, $M_{\text{cir}}=30$, $R=7.8$, $L=8.61$.}
  \label{dt_conv_1}
\end{figure}

We then perform the convergence test in $N$. We set $\Delta t = 0.01$ (the previous test suggests that this is sufficient to minimize the time discretization error). In Table~\ref{table 1}, we report the errors for different $N$ at $t_\text{final}=2$. Various values of the restitution coefficient $e = 0.2, 0.5$ and $0.8$, and both the ``full" and ``separate" methods are considered. These results imply that our method can indeed achieve the spectral accuracy, and both the ``full" and ``separate" methods behave similarly for the isotropic solution.

\begin{table}[htp!]
  \centering
  \subfloat[$e = 0.2$]{
  \begin{tabu} to 0.6\linewidth {X[1, c] X[3, c] X[3, c]}
    \toprule
    $N$ &  full & separate  \\
    \midrule
    8  & 9.21116565e-01& 9.21116565e-01 \\
    16 & 1.27640374e-02 & 1.27634481e-02  \\
    32  & 6.79745658e-06 & 6.79544555e-06 \\
    64  & 2.36438646e-10 & 2.34851361e-10 \\
    128  & 6.13890050e-11 & 6.30565600e-11 \\
    \bottomrule
  \end{tabu}
  } \\
  \subfloat[$e = 0.5$]{
  \begin{tabu} to 0.6\linewidth {X[1, c] X[3, c] X[3, c]}
    \toprule
    $N$ & full & separate  \\
    \midrule
    8  & 7.98706096e-01 & 7.98706096e-01 \\
    16  & 6.42644165e-03 & 6.42641236e-03 \\
    32  & 4.55730861e-06 & 4.55713801e-06 \\
    64  & 4.93770580e-11 & 4.93595165e-11 \\
    128  & 3.14279713e-11 & 3.13873372e-11 \\
    \bottomrule
  \end{tabu}
  } \\
  \subfloat[$e = 0.8$]{
  \begin{tabu} to 0.6\linewidth {X[1, c] X[3, c] X[3, c]}
    \toprule
    $N$  & full & separate\\
    \midrule
    8  & 5.52666966e-01 & 5.52666966e-01\\
    16  & 4.88586534e-04 & 4.88821204e-04 \\
    32  & 1.13897201e-07  & 1.14430393e-07\\
    64  & 9.82520731e-11 & 9.82359749e-11 \\
    128 & 1.00117470e-10  & 1.00099595e-10 \\
    \bottomrule
  \end{tabu}
  }
  \caption{Convergence test in $N$ of the 2D Maxwell molecule (isotropic solution) for $e=0.2$, $0.5$ and $0.8$, and for both the ``full" and ``separate" methods. Errors shown here are $|T_\text{num} - T_\text{ref}|$ at $t_\text{final}=2$. $\Delta t = 0.01$, $N_{\rho}=30$, $M_{\text{cir}}=30$, $R=7.8$, $L=8.61$.}
  \label{table 1}
\end{table}

\subsubsection{2D inelastic Maxwell molecules --- anisotropic solution}

To further test the performance of the ``full" and ``separate" methods, we consider the following double-well initial condition
\begin{equation}
  f_0(v) = \frac{0.8}{\pi}\left(\exp(-4|v-2|^2) + \exp(-|v+0.5|^2)\right).
\end{equation}
This function is anisotropic, and $\rho_0=1$, $u_0=0$, $T_0=1.425$.

We choose the same set of parameters as in the previous example and perform the convergence test in $N$. The results are shown in Table~\ref{table 2}. Overall, the accuracy degrades a bit compared to the isotropic solution. Surprisingly, the accuracy of the ``separate" method is much better than the ``full" method when $N$ gets bigger. Although it seems more natural to approximate the gain and loss terms together using the weak formulation, these results suggest that one should use the more accurate formula for the loss term when it is available (intuitively, the smoothing property of the gain term \cite{Lions94} alleviates the error introduced by the quadrature; since the loss term does not have this property, the same quadrature can incur more errors, hence a more accurate formula should be used).

\begin{table}[htp!]
  \centering
  \subfloat[$e = 0.2$]{
  \begin{tabu} to 0.6\linewidth {X[1, c] X[3, c] X[3, c]}
    \toprule
    $N$  & full & separate\\
    \midrule
    8  & 1.25303916e-01 & 1.25303916e-01 \\
    16  & 1.42811856e-02 & 1.41601818e-02 \\
    32  & 8.50383206e-05 & 1.21162093e-04 \\
    64  & 5.75217760e-05  & 8.65618628e-08\\
    128  & 5.74603408e-05 & 2.64749862e-08 \\
    \bottomrule
  \end{tabu}
  } \\
  \subfloat[$e = 0.5$]{
  \begin{tabu} to 0.6\linewidth {X[1, c] X[3, c] X[3, c]}
    \toprule
    $N$ & full & separate \\
    \midrule
    8  & 9.06935081e-02 & 9.06935081e-02 \\
    16  & 2.07345865e-02 & 2.06153352e-02 \\
    32  & 7.68575010e-05  & 1.08598123e-04\\
    64  & 4.58166915e-05 & 3.61540865e-08 \\
    128  & 4.57852460e-05 & 4.84827622e-09 \\
    \bottomrule
  \end{tabu}
  } \\
  \subfloat[$e = 0.8$]{
  \begin{tabu} to 0.6\linewidth {X[1, c] X[3, c] X[3, c]}
    \toprule
    $N$  & full & separate\\
    \midrule
    8 & 4.21932177e-02& 4.21932177e-02  \\
    16 & 2.17989934e-02 & 2.19257970e-02  \\
    32 & 1.17000137e-04 & 1.25546782e-04 \\
    64  & 2.42867607e-05  & 1.55334599e-08\\
    128  & 2.42767854e-05 &  5.91159477e-09 \\
    \bottomrule
  \end{tabu}
  }
  \caption{Convergence test in $N$ of the 2D Maxwell molecule (anisotropic solution) for $e=0.2$, $0.5$ and $0.8$, and for both the ``full" and ``separate" methods. Errors shown here are $|T_\text{num} - T_\text{ref}|$ at $t_\text{final}=2$. $\Delta t = 0.01$, $N_{\rho}=30$, $M_{\text{cir}}=30$, $R=8$, $L=8.83$.}
  \label{table 2}
\end{table}

\subsection{3D examples}

In 3D, the quadrature on the sphere can have a rather nontrivial effect. Therefore, we first test the accuracy of spherical design for Maxwell molecules. Using the same example, we also demonstrate the efficiency of our method by comparing it with the direct Fourier spectral method \cite{FPT05}. After this we numerically verify the physical Haff's cooling law for inelastic hard spheres.

Numerical examples in this subsection are generated using the ``separate" method (the loss term is computed as in (\ref{separate3}).

\subsubsection{3D inelastic Maxwell molecules}

Consider the constant collision kernel $B_{\sigma} = \frac{1}{4\pi}$ and the initial condition:
\begin{equation}\label{3Dbkw}
  f_0(v) = \frac{1}{2(2\pi K)^{3/2}}\exp\left(-\frac{v^2}{2K}\right)\left(\frac{5K-3}{K}+\frac{1-K}{K^2}v^2\right),
\end{equation}
where $K = 1 - \exp(-6.5/6)$. Note that $\rho_0=1$, $u_0=0$ and $T_0=1$.

We perform a convergence test in $M_{\text{sph}}$, the number of spherical design quadrature points used on the unit sphere. We set $e=0.2$, $t_\text{final} = 1$, $\Delta t = 0.01$, $N = 32$, and $N_{\rho}=30$. The results are reported in Table~\ref{table 3}. We find that the error does not decrease further when $M_{\text{sph}}$ reaches $32$. This is due to the fact that the measurement we use is the macroscopic quantity $T$ and other sources of errors such as the truncation of the domain and the discretization in $v$ may pollute the solution. In order to see this error more clearly, we compare the result of our method with that obtained from the direct spectral method (without any speedup strategy), see Table~\ref{table 4}.

\begin{table}[htp!]
  \centering
  \begin{tabu} to 0.8\linewidth {X[1,c] X[2,c]}
    \toprule
    $M_{\text{sph}}$ & $|T_\text{num} - T_\text{ref}|$\\
    \midrule
    6 & 0.0013957739849754791 \\
    12 & 9.9706271716293315e-05 \\
    32 & 2.2499901350947482e-06 \\
    48 & 2.4272557155313734e-06 \\
    70 & 2.4703481364962698e-06 \\
    94 & 2.4703481364962698e-06 \\
    120 & 2.453380453903975e-06 \\
    \bottomrule
  \end{tabu}
  \caption{Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule for $e=0.2$. Errors shown here are $|T_\text{num} - T_\text{ref}|$ at $t_\text{final}=1$. $\Delta t = 0.01$, $N = 32$, $N_{\rho}=30$, $R=7$, $L=7.72$.}
  \label{table 3}
\end{table}

\begin{table}[htp!]
  \centering
  \begin{tabu} to 0.8\linewidth {X[1,c] X[2,c]}
    \toprule
    $M_{\text{sph}}$ & $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$\\
    \midrule
 	6 & 0.00041820767783433107 \\
	12 & 3.1726851245650724e-05 \\
	32 & 6.5752814213618921e-07 \\
	48 & 5.6132233266191489e-07 \\
	70 & 2.6714628356683257e-07 \\
	94 & 1.0508293902503248e-07 \\
	120 & 2.8872928147741922e-08 \\ 
    \bottomrule
  \end{tabu}
  \caption{Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule for $e=0.2$. Errors shown here are $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$ at $t_\text{final}=1$. $\Delta t = 0.01$, $N = 32$, $N_{\rho}=30$, $R=7$, $L=7.72$.}
  \label{table 4}
\end{table}

To demonstrate the efficiency of our method, we show in Table~\ref{table 5} and Figure~\ref{CPUtime} the average running time of both the fast method and the direct method for one time evaluation of the collision operator, from which we see a significant speedup ($N=64$ of the direct method is left out due to the memory constraint).

\begin{table}[H]
	\centering
  \begin{tabu} to 0.6\linewidth {X[1, c] X[3, c] X[3, c]}
    \toprule
    $N$ & direct & fast \\
    \midrule
    8 & 32.3ms & 17.6ms \\
    16 & 602ms & 140ms \\
    32 & 23.7s & 1.18s \\
    64 & -- & 9.7s \\
    \bottomrule
  \end{tabu}
  	\caption{Average running time per evaluation of the collision operator. Comparison between the direct method and the fast method for various $N$ and fixed $N_{\rho}=30$, $M_{\text{sph}}=32$.}
	\label{table 5}
 \end{table}

\begin{figure}[htp!]
  \centering
  \includegraphics[width = .8\linewidth]{CPU_time}
  \caption{Same numbers as in Table~\ref{table 5} plotted in a log-log scale.}
  \label{CPUtime}
\end{figure}

\subsubsection{3D inelastic hard spheres --- Haff's cooling law}

In this last example, we try to observe the famous Haff's cooling law numerically. First proposed by Haff in 1983 in his seminal work about granular flows \cite{Haff83}, Haff's cooling law implies that the temperature in a spatially homogeneous granular gas composed of inelastic hard spheres decays like $O(t^{-2})$ with time. Note the algebraic decay rather than the exponential decay of Maxwell molecules (cf. (\ref{soln:T})).

We take the same initial data \eqref{3Dbkw} as in the previous example but for hard sphere molecules, that is, consider the collision kernel $B_{\sigma} = \frac{1}{4\pi}|g|$. Also we set $\varepsilon=0$ in equation (\ref{inBoltz}) (the Haff's law is for unheated case, i.e., the homogeneous cooling state). We numerically evaluate the temperature up to time $t_\text{final}=3$ with two different values of restitution coefficient $e=0.2$ and $0.8$. The time evolution of $T$ is shown in Figure~\ref{Haff_cooling}, from which one can see that the numerical temperature decays as 
\begin{equation}
T(t) = \frac{T_0}{(1+C_0 t)^2},
\end{equation}
where the constant $C_0$ depends on the value of $e$. This is in perfect agreement of the Haff's formula (see Section 4.1 of \cite{NE}).

\begin{figure}[htp!]
  \centering
  \includegraphics[width = .8\linewidth]{Haff's_cooling}
  \caption{Haff's cooling law. $\Delta t=0.01$, $N=32$, $N_{\rho}=30$, $M_{\text{sph}}=32$, $R=7$, $L=7.72$.}
  \label{Haff_cooling}
\end{figure}

\section{Conclusion}

In this work, we introduced a simple strategy to accelerate the direct Fourier spectral method for the inelastic Boltzmann collision operator, which is an accurate and popular deterministic method for approximating kinetic equations yet has been hindered in real applications due its huge computational cost and memory constraint. Our fast algorithm, by leveraging the low rank and convolutional structure in the collision integral, is able to reduce the computational complexity by orders of magnitude as well as relieve the memory bottleneck in the direct method. A series of numerical experiments has been performed to validate the accuracy and efficiency of the proposed method. In particular, we recovered successfully the Haff's cooling law for inelastic hard spheres. In the future, we will apply the fast solver to the spatially inhomogeneous setting and to simulate more interesting problems in granular materials.

\section*{References}

\bibliography{mybibfile}

\end{document}

### 2018-07-08 src/fast_spec_col_2d.py

@@ -1,0 +1,94 @@
+# fast_spec_col_2d
+
+import numpy as np
+import pyfftw
+from scipy import special
+from math import pi
+
+class FastSpectralCollison2D:
+    
+    def __init__(self, e, gamma, L, N, R, N_R, M=8):
+        # legendre quadrature
+        x, w = np.polynomial.legendre.leggauss(N_R)
+        r = 0.5*(x + 1)*R
+        self.r_w = 0.5*R*w
+        # circular points and weight
+        self.s_w = 2*pi/M
+        m = np.arange(0, 2*pi, self.s_w)
+        # index
+        k = np.fft.fftshift(np.arange(-N/2,N/2))
+        # dot with index
+        rkg = (k[:,None,None]*np.cos(m) + k[:,None]*np.sin(m))[...,None]*r
+        # norm of index
+        k_norm = np.sqrt(k[:,None]**2 + k**2)
+        # gain kernel
+        self.F_k_gain = 2*pi*r**(gamma+1)*(np.exp(0.25*1j*pi*(1+e)*rkg/L)
+                                           *special.jn(0, 0.25*pi*(1+e)*r*k_norm[...,None,None]/L))
+        # loss kernel
+        self.F_k_loss = 2*pi*r**(gamma+1)
+        # exp for fft
+        self.exp = np.exp(-1j*pi*rkg/L)
+        # j0
+        self.j0 = special.jn(0, pi*r*k_norm[...,None]/L)
+        # laplacian
+        self.lapl = -pi**2/L**2*k_norm**2
+        
+        # pyfftw planning of (N, N)
+        array_2d = pyfftw.empty_aligned((N, N), dtype='complex128')
+        self.fft2 = pyfftw.builders.fft2(array_2d, overwrite_input=True,
+                                   planner_effort='FFTW_ESTIMATE', threads=20, avoid_copy=True)
+        self.ifft2 = pyfftw.builders.ifft2(array_2d, overwrite_input=True,
+                                     planner_effort='FFTW_ESTIMATE', threads=20, avoid_copy=True)
+        # pyfftw planning of (N, N, N_R)
+        array_3d = pyfftw.empty_aligned((N, N, N_R), dtype='complex128')
+        self.fft3 = pyfftw.builders.fftn(array_3d, axes=(0,1), overwrite_input=True,
+                                   planner_effort='FFTW_ESTIMATE', threads=20, avoid_copy=True)
[…]

### 2018-07-08 src/fast_spec_col_3d.py

@@ -1,0 +1,99 @@
+# fast_spec_col_3d
+
+import numpy as np
+import pyfftw
+from math import pi
+
+EPS = 1e-8
+def sinc(x):
+    return np.sin(x+EPS)/(x+EPS)
+
+class FastSpectralCollison3D:
+    
+    def __init__(self, e, gamma, L, N, R, N_R, char='ss005.00012.txt'):
+        # legendre quadrature
+        x, w = np.polynomial.legendre.leggauss(N_R)
+        r = 0.5*(x + 1)*R
+        self.r_w = 0.5*R*w
+        # spherical points and weight
+        sigma = np.loadtxt(char)
+        self.s_w = 4*pi/sigma.shape[0]
+        # index
+        k = np.fft.fftshift(np.arange(-N/2,N/2))
+        # dot with index
+        rkg = (k[:,None,None,None]*sigma[:,0] 
+                    + k[:,None,None]*sigma[:,1] 
+                    + k[:,None]*sigma[:,2])[...,None]*r
+        # norm of index
+        k_norm = np.sqrt(k[:,None,None]**2 + k[:,None]**2 + k**2)
+        # gain kernel
+        self.F_k_gain = 4*pi*r**(gamma+2)*(np.exp(0.25*1j*pi*(1+e)*rkg/L)
+                                           *sinc(0.25*pi*(1+e)*r*k_norm[...,None,None]/L))
+        # loss kernel
+        self.F_k_loss = 4*pi*r**(gamma+2)
+        # exp for fft
+        self.exp = np.exp(-1j*pi*rkg/L)
+        # sinc
+        self.sinc = sinc(pi*r*k_norm[...,None]/L)
+        # laplacian
+        self.lapl = -pi**2/L**2*k_norm**2
+        
+        # pyfftw planning of (N, N, N)
+        array_3d = pyfftw.empty_aligned((N, N, N), dtype='complex128')
+        self.fft3 = pyfftw.builders.fftn(array_3d, overwrite_input=True,
+                                   planner_effort='FFTW_ESTIMATE', threads=8, avoid_copy=True)
+        self.ifft3 = pyfftw.builders.ifftn(array_3d, overwrite_input=True,
+                                     planner_effort='FFTW_ESTIMATE', threads=8, avoid_copy=True)
+        # pyfftw planning of (N, N, N, N_R)
+        array_4d = pyfftw.empty_aligned((N, N, N, N_R), dtype='complex128')
[…]

### 2018-08-13 paper/response/Response to Reviewers.tex
Add response

\documentclass[11pt]{article}
\usepackage{latexsym}
\usepackage{amssymb,amsbsy,amsmath,amsfonts,amssymb,amscd}
\usepackage{mathrsfs}
\usepackage{epsfig, graphicx}

\setlength{\oddsidemargin}{0mm} \setlength{\evensidemargin}{0mm}
\setlength{\topmargin}{5mm} \setlength{\textheight}{22cm}
\setlength{\textwidth}{17cm}
\renewcommand{\baselinestretch}{1.0}
\parindent 10pt

\usepackage[margin=1.0in]{geometry}

\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{cases}
\usepackage{color}

\newcommand{\jh}[1]{\textcolor{red}{JH: #1}}

\begin{document}

\title{Revision of manuscript JCOMP-D-18-00770  ``A fast spectral method for the inelastic Boltzmann collision operator and application to heated granular gases''}
\author{Jingwei Hu\footnote{Department of Mathematics, Purdue University, West Lafayette, IN 47907, USA (jingweihu@purdue.edu).}, \  \  
	    Zheng Ma\footnote{Department of Mathematics, Purdue University, West Lafayette, IN 47907, USA (ma531@purdue.edu).}
           }      
	\maketitle

\noindent {We would like to thank the anonymous reviewers for their valuable suggestions and comments, which greatly help us improve the quality and clarity of the manuscript. In the following, {\bf Q} denote the questions or suggestions raised by the reviewers, and {\bf A} refer to our answers. }

\vspace{0.3in}
\centerline{{\large \bf{Response to Reviewer \#1}}}
\bigskip

\begin{itemize}
\item[{\bf Q1}]  {\it Around lines 53 and 54, the authors mentioned that ``the method by Wu et al. only works for 2D Maxwell molecules and 3D hard sphere...". This is absolutely not true as clearly stated in Ref. 16 (see the paragraph after Eq. 22) that the spectral method can handle the Enskog collision term with a general form of collision kernels, which has been demonstrated by the numerical results for hard-sphere, Maxwell, and soft-potential molecules in Fig. 2 therein.}

\item[{\bf A1}]

\item[{\bf Q2}] {\it In lines 60 and 61, evidence on the advantage of `spherical design' over `Gauss quadrature' should be provided.}

\item[{\bf A2}]  \jh{Add a comparsion}

\item[{\bf Q3}] {\it G in Eq. 40 is calculated ``exactly," while the gain term is approximated by quadrature. This may violate the mass conservation. This problem may be removed by calculating G using the same quadrature rule as for the gain term.
}

\item[{\bf A3}]

\item[{\bf Q4}] {\it In Fig. 2 when comparing the direct method to the Fourier spectral method, the software, computer configurations, number of CPU cores used, etc. should be specified.  }

\item[{\bf A4}]

\item[{\bf Q5}] {\it The accuracy of the proposed Fourier spectral method is proved in spatially homogeneous relaxation problems. The paper is a little bit short. I think at least one spatially-inhomogeneous problem should be simulated to test the stability of the present method because some deterministic numerical method fails to solve this problem although they can solve spatially homogeneous relaxation problems. }

\item[{\bf A5}]

\item[{\bf Q6}] {\it The extendibility of the present method to deal with the Enskog collision term where the binary collision is not spatially localized need to be commented in order the compare the fast spectral method developed in Ref. 16. }

\item[{\bf A6}]

\end{itemize}

\vspace{0.3in}
\centerline{{\large \bf{Response to Reviewer \#2}}}
\bigskip

{\bf Critque.} {\it In the reviewer opinion, this work is not a true novelty, because it arrives more than three years after the fast spectral method for the Wu et al. paper [16] (which was in itself a refinement of a paper from the same group about multiple species Boltzmann equation). This method is almost the same than the one introduced in the manuscript, and the reviewer hardly sees then the relevance of such a work, in particular for possible publication in JCP, journal which ``focuses on the computational aspects of physical problems. The scope of the Journal is the presentation of new significantly improved techniques for the numerical solution of problems in all areas of physics." The present manuscript hardly fits in this scope, especially given the fact that the numerical simulations presented are only space homogeneous.

Concerning the overall paper, the most important point for the reviewer is that the authors doesn't compare their methods with the one introduced in [16]. Comparing with the Classical Spectral Method from [11] is nice, and gives strong confidence regarding convergence of the new method, but concerning the accuracy and the computational times, [16] would be more relevant. Moreover, the reviewer thinks that, even if the derivation is different between here and [16], in the constant restitution coefficient case, both methods are the same (this fact is even almost said by the authors p. 4, l. 50). In that regards, the paper would gain a lot of interest to the reviewer if numerical simulations for the non-constant restitution coefficient-case were available.
}

First of all, this reviewer's comment ``This method is almost the same than the one introduced in the manuscript" is absolutely not true! The method in [16] is based on the so-called Carleman representation of the collision operator. This representation is first used in Mouhot and Pareschi (2006) to derive a fast Fourier method for the classical (elastic) Boltzmann collision operator and is in fact the basis of the method proposed in [16]. Our method is based on the $\sigma$-representation of the collision operator. The two representations are equivalent at the continuous level. However, they become quite different when numerical approximation is applied, for example, the truncation of the integral is done on different variables, in our method, it is $g=v-v_*$, the relative velocity is truncated, while in [16], it is $x=(|g|\sigma-g)/2$ and $y=-g-x$ are truncated; furthermore, our numerical quadrature is applied on $g$, while in [16], it is $x$ that is discretized. These will of course result in different approximations of the collision operator. Another difference is that, our method makes no assumptions on the collision kernel and applies to the general form $B_{\sigma}(|g|,\sigma\cdot \hat{g})$. The method in [16], although can treat general kernels, is based on a modification of the kernel so that it is in the separable form. In view of this, our method does not need any special treatment/approximation of the kernel and is much easier to implement. Moreover, we propose to use the sophisticated quadrature, spherical design, on the sphere, in contrast to the tensor product based Gauss quadrature in [16], which results in much better accuracy as demonstrated in our paper.

\jh{Our method provides better accuracy compared to that in [16], do the same numerical test in Section 4.1 of [16] ($10^{-8}$ v.s. $10^{-5}$). I think [16] does not contain a CPU time, please double check.}

The reviewer's comment ``even if the derivation is different between here and [16], in the constant restitution coefficient case, both methods are the same (this fact is even almost said by the authors p. 4, l. 50)" is absolutely wrong as already explained in the previous paragraph. Both methods are different no matter $e$ is a constant or not. The statement in page 4, line 50 concerns the difference between the Boltzmann and Enskog operators and it is by no means related to the numerical methods!

\begin{itemize}

\item[{\bf Q1}] {\it P. 4 l. 70 the reviewer is surprised of the term ``ambiguity" used for the strong form of the collision operator. The series of papers by eg Mischler, Mouhot et al. (JSP 2006 part 1 and 2, CMP 2009, DCDS A 2009) eg Alonso, Lods et al. (CMS 2011, CMP 2013 and SIMA 2015) asserts this expression quite clearly. More work on this topic is needed.}

\item[{\bf A1}] The weak form of the inelastic collision operator is never of an issue and is consistent in the literature. However, one can find different strong forms from various sources and it is because of this reason, in [19], the authors (in particular, J. A. Carrillo is among the first few people to study the mathematical properties of the inelastic Boltzmann equation) spent a few pages clarifying this issue. This part of presentation in our paper is a shorter version and serves as a starting point of the discussion of the inelastic operator.

\item[{\bf Q2}] {\it On p. 7, before (18), the reviewer would like to note that in the inelastic settings, the particles are macroscopic bodies (pollen, meteoroids, beads, ...) and not perfect molecules. This is the reason of the inelasticity of the collision, and as such, the collision kernel should be only of hard sphere type! The variable hard sphere kernel is not intended for granular gases, only molecular ones.}

\item[{\bf A2}] We agree that the physically relevant case is the hard sphere molecules but for theoretical purpose and numerical validation, other types of molecules, for example the Maxwell molecules, are also useful. In fact, the reference [16] also considers the variable hard sphere kernel (see Section 4.1 in their paper).

\item[{\bf Q3}] {\it On p. 9, the computation of the temperature for the constant kernel is done, but the authors could also give some ideas of where the Haff's law comes from (in particular the value of the coefficient $C_0$ in the simulations).
}

\item[{\bf A3}] $C_0=1/(4\pi)$ as mentioned at the bottom of page 21.

\item[{\bf Q4}] {\it On p. 10, a reference about the fact that it suffices to take R = 2S would be needed, because that doesn't seems that easy for the reviewer. Same goes for the value of L latter in this page, because the inelastic case is slightly different on this regard that the elastic one.}

\item[{\bf A4}] \jh{Refer to [11] or briefly add the discussion to the paper.}

\item[{\bf Q5}] {\it When one deals with Fast Spectral methods, reference to the seminal paper of Mouhot and Pareschi, Math. Comp. 2005, would be necessary (for example on p. 11, before (36)).}

\item[{\bf A5}] \jh{Mention that this is a different method....}

\item[{\bf Q6}] {\it The reviewer doesn't understand AT ALL the assertion at the end of p. 11 saying that the ``integrand in (39) is oscillatory on the scale of $O(N)$". What does that means? More details (a lot !) are needed here!}

\item[{\bf A6}]

\item[{\bf Q7}] {\it On p. 12, expressions for the weight $w_{\rho}$ and $w_g$ are missing, as well as more details on the underlying numerical methods used here!}

\item[{\bf A7}]

\item[{\bf Q8}] {\it On the same page, l. 125, some more explication on the numerical complexity is needed, because the order of magnitudes are not the same than in the introduction section. The same goes for the storage requirement: numerical experiments showing that the new method is better on that regard than the classical one is needed.
}

\item[{\bf A8}]

\item[{\bf Q9}] {\it What about the conservations properties of the new scheme? Are there any numerical simulations available?
}

\item[{\bf A9}] The Fourier spectral method does not have the conservation. We did not check the mass and momentum but the extensive results presented regarding the temperature demonstrate that the method can capture the moments fairly accurately.

\item[{\bf Q10}] {\it Have the authors tried to take $e=1$ namely the elastic case, in order to see if the BKW solution is obtained?
}

\item[{\bf A10}] \jh{Maybe include it in Section 4.1.1 as this is quite easy.}

\item[{\bf Q11}] {\it On p. 14, using spectral to compute the laplacian is nice, but what is the effect on the aliasing properties? Should one needs to change the truncation? this should be studied too, because the Laplace operator spreads in velocity, whereas Q concentrates, and the analysis on R and S is then broken.

}

\item[{\bf A11}]

\item[{\bf Q12}] {\it On pp. 15-16, more explanations would be nice regarding the discrepancies between the decoupling approach in 2D and 3D. Moreover, [24] is not the good reference for the smoothing properties of the gain term, because this paper concerns only the elastic case. References toward on of the Mischler paper would be better (and the smoothing properties are by the way quite difference in the inelastic case).
}

\item[{\bf A12}] \jh{There is no rigorous proof except the numerical evidence. However, we do believe this is attributed to the smoothing property of the gain term. Cite the paper ``Cooling Process for Inelastic Boltzmann Equations for Hard Spheres, Part II: Self-Similar Solutions and Tail Behavior" by Mischler and Mouhot for the first regularity proof of the gain term of the inelastic collision operator. In fact, it is mentioned in this paper the regularity result is similar to the elastic case and the proof follows the lines in the elastic case.... See the discussion in Section 1.6 of that paper.}

\item[{\bf Q13}] {\it On p. 5, l. 85, the expression for $(,*)$ would improve the readability, as well as the one of J in (7).}

\item[{\bf A13}] \jh{I guess what he means is to give explicitly the formulas for $\tilde{v}$, $\tilde{v}_*$ and $J$. We can do this but need to mention that this is only possible when $e$ is a constant.}

\item[{\bf Q14}] {\it On p. 6, it seems to the reviewer that it is usually easier to derive the strong form of Q, start- ing from its weak form, rather than the contrary (due to the irreversible microscopic collision dynamics).}

\item[{\bf A14}] This is more of a personal taste. The review paper [4] by Villani also starts with the strong form and then introduces the weak form. In fact, it is quite intuitive to understand the strong form directly: the pre-collisional velocities $\tilde{v}$ and $\tilde{v}_*$ become $v$ and $v_*$ after the collision, hence yields the gain term of the collision operator.

\item[{\bf Q15}] {\it Why does the author decide in the numerical section to use the BKW initial conditions? This has no physical relevancy in the inelastic case (and too much symmetries).}

\item[{\bf A15}] We considered the BKW solution in Section 4.1.1 as a starting example mainly because it is an isotropic solution and becomes an exact solution when $e=1$. Later in Section 4.1.2 we did consider an anisotropic solution (see equation (50)) and observed quite different behavior. \jh{Why didn't we consider an anisotropic solution in 3D?}

\end{itemize}

\end{document}

### 2018-08-13 paper/draft/elsarticle-template0.tex
add refs

--- 2018-07-08 paper/draft/elsarticle_v2.tex
@@
 % personal macros
-\newtheorem{thm}{Theorem} \newtheorem{lem}[thm]{Lemma} \newdefinition{rmk}{Remark} \newproof{pf}{Proof}
-\newproof{pot}{Proof of Theorem \ref{thm2}}
-
 \newcommand*\diff{\mathop{}\!\mathrm{d}}
@@
 where $f(t,x,v)$ is the one-particle distribution function depending on the time $t \geq 0$, position $x \in \mathbb{R}^d$ and velocity $v \in \mathbb{R}^d$ with the dimension $d \geq 1$, and $Q_{\text{in}}$ is the so-called inelastic (or granular) collision operator, whose exact expression is presented in later discussions.
-For convenience, we use $Q$ instead of $Q_{\text{in}}$ in the rest of this paper, whose notation should not be confused with the usual elastic collision operator.
-Another widely used model, first introduced by van Noije and Ernst \cite{}, is the spatial homogeneous inelastic Boltzmann equation based on Enskog-Boltzmann model with a heat source:
+Another widely used model, first introduced by van Noije and Ernst \cite{}, is the spatial homogenous inelastic Boltzmann equation based on Enskog-Boltzmann model with a heat source:
 \begin{equation} \label{inBoltz}
-\partial_tf-\varepsilon \Delta_vf=Q(f,f),
+\partial_tf-\varepsilon \Delta_vf=Q_{\text{in}}(f,f),
 \end{equation}
@@
 Finally in Section 4, a number of numerical tests are performed to test both accuracy and efficiency of our fast algorithm in 2D and 3D cases.
-Also, as can be in the 3D numerical example, the Haff's cooling law is verified.
+Also, as can be in the 3D numerical example, the Haff's coolling law is verified.
 
@@
 \end{equation}
-Now we are ready to define the inelastic collision operator in its weak form
-\begin{equation} \label{col_omega}
-\int Q(f,f)(v)\phi(v)\diff{v}= \frac{1}{2} \int_{\mathbb{R}^d} \int_{\mathbb{R}^d} \int_{\mathbb{S}^{d-1}} B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_*\left(\phi'+\phi_*'-\phi-\phi_* \right)\diff{\omega} \diff{v} \diff{v_*},
-\end{equation}
-where $g := v - v_\ast$, $\hat{g}$ denotes the unit vector in the direction of $g$, $\phi(t,v)$ a regular test function, $B_{\omega}$ the binary collision kernel in the $\omega$-representation, and we use the shorthands $f$ for $f(t,v)$, $f_\ast$ for $f_\ast(t, v_\ast)$ and so on.
-This operator clearly has collisional invariants $1$ and $v_i$ ($i = 1, 2, 3$), namely,
-\begin{equation}
-  \int_{\mathbb{R}^3} Q(f, f)(v) \diff v = \int_{\mathbb{R}^3} Q(f,f)(v)v_i\diff v = 0,
-\end{equation}
-where $v_i$ denotes the $i$-th component of the velocity $v$.
+Now we are ready to define the inelastic collision operator
 
-Note that \eqref{omega} can be also written in the so-called $\sigma$-representation, where $\sigma$ is related to $\omega$ in the following way:
+Note that for numerical purpose, \eqref{omega} can be also written in the so-called $\sigma$-representation, where $\sigma$ is related to $\omega$ in the following way:
 \begin{equation} \label{relation}
@@
 \end{align*}
+
 The transformation between $\omega$ and $\sigma$ representations is given as follows:
-\begin{align*}
-& B_{\omega}(|g|, |\omega\cdot \hat{g}|) \diff{\omega} \diff{v_*}  = B_{\sigma}(|g|,\sigma\cdot \hat{g})\diff{\sigma} \diff{v_*}  , \\
-& B_{\omega}(|g|, |\omega\cdot \hat{g}|) =|2(\omega\cdot \hat{g})|^{d-2}B_{\sigma}(|g|,1-2(\omega\cdot\hat{g})^2),
-\end{align*}
-where $B_\sigma$ denotes the collision kernels in $\sigma$-representations.
+\begin{align}
+\int_{\mathbb{R}^d}\int_{S^{d-1}}\, \cdot  \,\, B_{\sigma}(|g|,\sigma\cdot \hat{g})\,\diff{\sigma}\,\diff{v_*} = \int_{\mathbb{R}^d}\int_{S^{d-1}}\, \cdot  \,\, B_{\omega}(|g|, |\omega\cdot \hat{g}|)\,\diff{\omega}\,\diff{v_*},
+\end{align}
+where
+\begin{equation}
+B_{\omega}(|g|, |\omega\cdot \hat{g}|)=|2(\omega\cdot \hat{g})|^{d-2}B_{\sigma}(|g|,1-2(\omega\cdot\hat{g})^2),
+\end{equation}
+where $B_\omega$ and $B_\sigma$ are the collision kernels in its respective representations.
 Two common cases to keep in mind are the 2D pseudo (no angular dependence) Maxwell molecule, where $B_{\sigma}=B_{\omega}=1$; and the 3D hard sphere, with $B_{\sigma}=|g|$, $B_{\omega}=2|g\cdot\omega|$.
-Hereafter, we shall use the collision operator in its $\sigma$-representation
-\begin{equation} \label{col_sigma}
-\int Q(f,f)(v)\phi(v)\diff{v}= \frac{1}{2} \int_{\mathbb{R}^d} \int_{\mathbb{R}^d} \int_{\mathbb{S}^{d-1}}
-B_{\sigma}(|g|,\sigma\cdot \hat{g})ff_*\left(\phi'+\phi_*'-\phi-\phi_* \right)\diff{\sigma}\diff{v}\diff{v_*}.
+
+We now derive the strong form of the inelastic Boltzmann collision operator. We define $'v$ and $'v_*$ to be the velocities before $v$ and $v_*$, i.e.,
+\begin{equation}
+('v,'v_*) \rightarrow (v,v_*) \rightarrow (v',v_*'),
 \end{equation}
-since its convenience for numerical purposes, which will be discussed with more details later.
+and we can also compute the Jacobian between these transformations:
+\begin{equation}
+\frac{\partial(v,v_*)}{\partial ('v,'v_*)}=\frac{\partial(v',v_*')}{\partial (v,v_*)}=-e, \quad \frac{\partial ('v,'v_*)}{\partial(v,v_*)}=\frac{\partial (v,v_*)}{\partial(v',v_*')}=-\frac{1}{e}.
+\end{equation}
+
+We start with the weak form:
+\begin{equation} \label{weak1}
+\int Q(f,f)(v)\phi(v)\,\diff{v}=\iiint B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_*(\phi'-\phi)\,\diff{\omega}\,\diff{v}\,\diff{v_*},
+\end{equation}
+and take $\phi(v)=\delta(v-v_0)$, then
+\begin{align}
+  Q(f,f)(v_0)&=\iiint B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_*(\delta(v'-v_0)-\delta(v-v_0))\,\diff{\omega}\,\diff{v}\,\diff{v_*}\nonumber\\
+  &=\iiint B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_*\delta(v'-v_0)\,\diff{\omega}\,\diff{v}\,\diff{v_*}-\iint B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_*|_{v=v_0}\,\diff{\omega}\,\diff{v_*}\nonumber\\
+  &=\iiint B_{\omega}(|'g|,|\omega\cdot '\hat{g}|)'f'f_*\delta(v-v_0)\,\diff{\omega}\,\diff{'v}\,\diff{'v_*}-\iint B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_*|_{v=v_0}\,\diff{\omega}\,\diff{v_*}\nonumber\\
+  &=\iiint B_{\omega}(|'g|,|\omega\cdot '\hat{g}|)'f'f_*\delta(v-v_0) \frac{1}{e}\diff{\omega}\,\diff{v}\,\diff{v_*}-\iint B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_*|_{v=v_0}\,\diff{\omega}\,\diff{v_*}\nonumber\\
+  &=\iint B_{\omega}(|'g|,|\omega\cdot '\hat{g}|)'f'f_*|_{v=v_0} \frac{1}{e}\diff{\omega}\,\diff{v_*}-\iint B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_*|_{v=v_0}\,\diff{\omega}\,\diff{v_*},
+\end{align}
+where in the third equality, we changed $(v,v_*)$ to $('v,'v_*)$, correspondingly, $(v',v_*')$ is changed to $(v,v_*)$.
+
+Therefore, we have obtained
+\begin{equation}
+Q(f,f)(v)=\iint B_{\omega}(|'g|,|\omega\cdot '\hat{g}|)'f'f_* \frac{1}{e}\diff{\omega}\,\diff{v_*}-\iint B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_*\,\diff{\omega}\,\diff{v_*}.
+\end{equation}
+Note that
+\begin{equation}
+|g'|^2=|g|^2+(e^2-1)(g\cdot \omega)^2, \quad g\cdot \omega=-\frac{1}{e}g'\cdot\omega,
+\end{equation}
+then
+\begin{equation}
+|g|^2=|g'|^2+(\frac{1}{e^2}-1)(g'\cdot \omega)^2=\frac{|g'|^2}{e^2}[e^2+(1-e^2)(\omega \cdot \hat{g}')^2],
+\end{equation}
+and
+\begin{equation}
+\omega\cdot \hat{g}=-\frac{\omega\cdot\hat{g}'}{\sqrt{e^2+(1-e^2)(\omega\cdot \hat{g}')^2}}.
+\end{equation}
+So
+\begin{equation}
+B_{\omega}(|'g|,|\omega\cdot '\hat{g}|)=B_{\omega}\left(\frac{|g|}{e}\sqrt{e^2+(1-e^2)(\omega\cdot \hat{g})^2}, \frac{\omega \cdot \hat{g}}{\sqrt{e^2+(1-e^2)(\omega\cdot \hat{g})^2}}\right).
+\end{equation}
+
+In most situations, we will just use the weak form (\ref{weak1}), or an equivalent form
+\begin{equation} \label{weak2}
+\int Q(f,f)(v)\phi(v)\,\diff{v}=\iiint B_{\omega}(|g|,|\omega\cdot \hat{g}|)ff_*\frac{\phi'+\phi_*'-\phi-\phi_*}{2}\,\diff{\omega}\,\diff{v}\,\diff{v_*}.
+\end{equation}
+These weak forms written in $\sigma$-representation are
+\begin{equation} \label{weak11}
+\int Q(f,f)(v)\phi(v)\,\diff{v}=\iiint B_{\sigma}(|g|,\sigma\cdot \hat{g})ff_*(\phi'-\phi)\,\diff{\sigma}\,\diff{v}\,\diff{v_*},
+\end{equation}
+\begin{equation} \label{weak22}
+\int Q(f,f)(v)\phi(v)\,\diff{v}=\iiint B_{\sigma}(|g|,\sigma\cdot \hat{g})ff_*\frac{\phi'+\phi_*'-\phi-\phi_*}{2}\,\diff{\sigma}\,\diff{v}\,\diff{v_*}.
+\end{equation}
 
@@
 
-Consider the inelastic Boltzmann equation with a heating source \eqref{inBoltz}.
-We shall revisit a few properties of this equation.
-First, for a solution $f$ to \eqref{inBoltz}, the conservation of mass and momentum becomes straightfoward taking $\phi = 1$ and $\phi = v$ respectively in \eqref{col_sigma}.
-In other words,
+We often consider the inelastic Boltzmann equation with a heating source:
+\begin{equation} \label{inBoltz}
+\partial_tf-\varepsilon \Delta_vf=Q(f,f),
+\end{equation}
+where $\varepsilon$ is a small parameter, $Q(f,f)$ is given via the weak forms (\ref{weak11}) or (\ref{weak22}). Let's derive a few properties of this equation.
+
+Define density, momentum and energy as
 \begin{equation}
-\rho := \int f\diff{v}\equiv \rho_0, \quad  u :=\frac{1}{\rho}\int fv \diff{v}\equiv u_0.
+\rho=\int f\,\diff{v}, \quad  m=\int fv\,\diff{v}, \quad E=\int f\frac{1}{2}v^2\,\diff{v}.
 \end{equation}
-where $\rho$ is the density and $u$ the bulk velocity. Similarly, define the kinetic energy as
+Taking the moments $\int \cdot \, \phi(v)\,\diff{v}$ on both sides of (\ref{inBoltz}), we have
 \begin{equation}
-  E := \frac{1}{2}\int f |v|^2\diff v
+\partial_t \int f\phi\,\diff{v} -\varepsilon \int f \Delta \phi \,\diff{v}=\iiint B_{\sigma}(|g|,\sigma\cdot \hat{g})ff_*\frac{\phi'+\phi_*'-\phi-\phi_*}{2}\,\diff{\sigma}\,\diff{v}\,\diff{v_*}.
 \end{equation}
-and take $\phi = |v|^2$, one obtains the relation of energy dissipation
-\begin{equation}\label{ki_energy}
-\partial_t E -\varepsilon d\rho_0=-\frac{1-e^2}{16}\iiint B_{\sigma}(|g|,\sigma\cdot \hat{g})|g|^2(1-\sigma\cdot \hat{g})ff_*\diff{\sigma}\diff{v}\diff{v_*},
+Therefore, if $\phi=1$ and $v$, we have the conservation of mass and momentum
+\begin{equation}
+\rho \equiv \rho_0, \quad m\equiv m_0.
 \end{equation}
-where we used (\ref{lossenergy}) and (\ref{relation}). Consider the following collision kernel
+If $\phi=\frac{1}{2}v^2$, we have
+\begin{equation}
+\partial_t E -\varepsilon d\rho_0=-\frac{1-e^2}{16}\iiint B_{\sigma}(|g|,\sigma\cdot \hat{g})|g|^2(1-\sigma\cdot \hat{g})ff_*\,\diff{\sigma}\,\diff{v}\,\diff{v_*},
+\end{equation}
+where we used (\ref{lossenergy}) and (\ref{relation}).
+
+Now we consider the following collision kernel
 \begin{equation} \label{kernel}
@@
 \end{equation}
-where $C_{\lambda}$ is some constant and $b_{\lambda}$ is some function. For Maxwell molecule, i.e., $\lambda=0$ in (\ref{kernel}), equation \eqref{ki_energy} becomes
+where $C_{\lambda}$ is some constant and $b_{\lambda}$ is some function. For Maxwell molecule, i.e., $\lambda=0$ in (\ref{kernel}), the above equation becomes
 \begin{equation} 
@@
 \end{equation}
-is a constant regardless of $\hat{g}$ and, without loss of generality, is assumed to be $1$. Then
+is a constant regardless of $\hat{g}$. Assume this constant is 1. Then
 \begin{align} 
-  \partial_t E -\varepsilon d\rho_0 &=-\frac{1-e^2}{16}\iint |g|^{2}ff_*\,\diff{v}\,\diff{v_*} \notag\\
-   & =-\frac{1-e^2}{16}\iint (v^2+v_*^2-2v\cdot v_*)ff_*\,\diff{v}\,\diff{v_*} \notag \\
-   &=-\frac{1-e^2}{8}\left(2\rho_0E - (\rho_0 u_0)^2\right).
+\partial_t E -\varepsilon d\rho_0&=-\frac{1-e^2}{16}\iint |g|^{2}ff_*\,\diff{v}\,\diff{v_*}=-\frac{1-e^2}{16}\iint (v^2+v_*^2-2v\cdot v_*)ff_*\,\diff{v}\,\diff{v_*}\nonumber\\
+&=-\frac{1-e^2}{8}(2\rho_0E-m_0^2).
 \end{align}
-
-Finally, the temperature defined by
-\begin{equation}
-  T = \frac{1}{\rho d}\int_{\mathbb{R}^d}f|u - v|^2\diff v
-\end{equation}
-goes to $0$ as time approaches infinity. In particular, if the initial condition is $\rho_0=1$, $u_0=0$, then $T=\frac{2}{d}E$ satisfies
+In particular, if the initial condition is $\rho_0=1$, $u_0=0$, then $E=\frac{d}{2}T$, the temperature $T$ hence satisfies
 \begin{equation}
@@
 \end{equation}
-\begin{rmk}
-This explicit expression serves as the reference quantity for numerical tests, since in inelastic cases it is notably harder to find a special $f$ such that the collision operator has an analytical expression in contrast to elastic cases.
-\end{rmk}
 
@@
 
-In this section, we construct the fast spectral algorithm.
-Before that, let us first revisit the rather conventional spectral method and its direct solver.
-The idea of the spectral method is rather straightfoward, that is, to exploit the structure of the collision operator in its Fourier space.
-Specificallly, by choosing the test function $\phi(v)=e^{-i\frac{\pi}{L}k\cdot v}$ in the weak form \eqref{col_sigma}, one can obtain the expression of the $k$-th mode in the Fourier expansion of $Q$:
+By choosing the test function $\phi(v)=e^{-i\frac{\pi}{L}k\cdot v}$ in the weak form (\ref{weak11}), we can obtain the Fourier expansion of $Q$:
+\vspace{-0.05in}
 \begin{equation} \label{sum1}
@@
 where the weight $G(l,m)$ is given by
-\begin{equation*}  \label{G}
-G(l,m)=\int_{\mathbb{R}^d}e^{-i\frac{\pi}{L}m \cdot g}\left[\int_{S^{d-1}}B_{\sigma}(|g|,\sigma\cdot \hat{g})\left(e^{-i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot (|g|\sigma-g)}-1\right)\diff{\sigma}\right]\diff{g},
+\begin{equation*} 
+G(l,m)=\int_{\mathbb{R}^d}e^{-i\frac{\pi}{L}m \cdot g}\left[\int_{S^{d-1}}B_{\sigma}(|g|,\sigma\cdot \hat{g})\left(e^{-i\frac{\pi}{L}\frac{1+e}{4}(l+m)\cdot (|g|\sigma-g)}-1\right)\,\diff{\sigma}\right]\diff{g},
 \end{equation*}
-here $g$ needs to be truncated properly as was done for the elastic case \cite{gamba_hu}.
+here $g$ needs to be truncated properly as was done for the elastic case.
 
-For the three-dimensional VHS model \eqref{kernel}, it follows by a straightfoward calculation that the formula for $G$ reduces to
-\begin{equation} \label{direct_G}
-G(l,m)=16\pi^2C_{\gamma}\int_0^R \rho^{\gamma+2}\left[\text{Sinc}\left(\frac{\pi}{L}\left|\frac{1+e}{4}(l+m)-m\right|\rho\right)\text{Sinc}\left(\frac{\pi}{L}\frac{1+e}{4}|l+m|\rho\right)-\text{Sinc}\left(\frac{\pi}{L}\rho|m|\right)\right]\diff\rho.
-\end{equation}
-Hence, one can construct a direct spectral solver by evaluating the numerical quadrature of \eqref{direct_G} and proceed the summation of \eqref{sum1}.
-However, as is mentioned before, this approach results in a high computational cost of $O(N^6)$ and storage of $O(N^6)$, apparently because there is no convolution structure in the summation to follow.
-In applications, the scale of problems that one can compute using this approach may become quickly bottlenecked due to the computational limitations, which motivates the study of a fast spectral algorithm.
-
-We begin this exploration by asking ourselves whether there are any underlying convolution structures to be unveiled.
-While it becomes unclear once using the final analytical expression \eqref{direct_G} of $G(l,m)$, there are in fact hidden convolution structures in equation \eqref{G}, namely, to consider a partial numerical quadrature of this intermidiate form instead of integrating it analytically.
-In particular, the idea of our fast algorithm is to separate the weight $G$ as $G(l,m)\approx\sum_{t=1}^T\alpha_t(l+m)\beta_t(m)$ using quadrature rules.
+The idea of the fast algorithm is to separate the weight $G$ as $G(l,m)\approx\sum_{t=1}^T\alpha_t(l+m)\beta_t(m)$ using quadrature rules.
 
@@
 
-In this section, we first verify the accuracy and efficiency of the proposed method with extensive numerical studies in two dimensional cases, and then provide three dimensional numerical results as well.
-Note that unlike the elastic case where the analytical results of the collision kernel are accessible by choosing a suitable function $f$, the inelastic case, however, only has the analytical formula for its macro quantities, such as the temperature $T$ as is shown in \eqref{soln:T}.
-This means in order to provide a trustworthy verification of our solver, one needs a numerical scheme to solve the inelastic Boltzmann equation \eqref{inBoltz}.
+In this section, we shall verify the accuracy and efficiency of the proposed method with extensive two dimensional numerical studies, and provide simulation examples in two and three dimensional cases.
+Unlike in the elastic case where one can obtain the analytical results of the collision kernel by choosing suitable function $f$, we only have the analytical formula for the macro quantities, such as temperature $T$ as shown in \eqref{soln:T}.
+This means in order to check the accuracy of our method we need a numerical scheme to solve the inelastic Boltzmann equation with heating sources \eqref{inBoltz}
+\begin{equation*}
+  \partial_t f-\varepsilon \Delta_vf=Q(f,f),
+\end{equation*}
+or the equation withour heating sources
+\begin{equation}
+  \partial_t f - \varepsilon \Delta_vf = Q(f, f).
+\end{equation}
 
-In the following numerical simulations, we use our fast spectral solver to evaluate the collision operator, paired with a standard fourier spectral method to treat the diffusion term $\varepsilon\Delta f$ and the SSP-RK3 (the 3rd order Strong Stability-Preserving Runge-Kutta method) \cite{ShuTadmor} for time discretizations.
+In the following numerical tests and simulations, we will use explicit Runge-Kutta methods for time discretization and standard fourier spectral method for the heating term $\varepsilon\Delta f$, and also our fast spectral method for the evaluation of the collision term.
+
 \subsection{2D examples - Maxwell molecules}
@@
 \paragraph{\bf Accuracy of spherical design}
-For 3D
+
 \paragraph{\bf Haff's cooling law}

### 2018-08-14 src/fast_spec_col_2d.py
Add config file and code refactoring.

@@ -8,6 +8,7 @@ from math import pi
 class FastSpectralCollison2D:
     
     def __init__(self, e, gamma, L, N, R, N_R, M=8):
+        gamma, self.e = config.physical_config.gamma, config.physical_config.e
         # legendre quadrature
         x, w = np.polynomial.legendre.leggauss(N_R)
         r = 0.5*(x + 1)*R

### 2018-08-14 src/fast_spec_col_3d.py

@@ -18,6 +18,13 @@ class FastSpectralCollison3D:
         # spherical points and weight
         sigma = np.loadtxt(char)
         self.s_w = 4*pi/sigma.shape[0]
+        # integral on unit sphere using quadrature
+        theta, w_theta = np.polynomial.legendre.leggauss(M_sigma[0])
+        theta = 0.5*(theta + 1)*2*pi
+        w_theta = 0.5*2*pi*w_theta
+        phi, w_phi = np.polynomial.legendre.leggauss(M_sigma[1])
+        phi = 0.5*(phi + 1)*pi
+        w_phi = 0.5*pi*w_phi
         # index
         k = np.fft.fftshift(np.arange(-N/2,N/2))
         # dot with index

### 2018-08-14 src/utility.py

@@ -1,0 +1,101 @@
+# ===================================================================
+# Utility functions for different time marching schemes
+# ===================================================================
+
+def Euler(f, L, eps, dt):
+    return f + dt*L(f, eps)
+
+def RK2(f, L, eps, dt):
+    k1 = L(f, eps)
+    return f + dt*L(f + 0.5*dt*k1, eps)
+
+def RK3(f, L, eps, dt):
+    k1 = L(f, eps)
+    k2 = L(f + 0.5*dt*k1, eps)
+    k3 = L(f - dt*k1 + 2*dt*k2, eps)
+    return f + 1/6*dt*(k1 + 4*k2 + k3)
+
+def RK4(f, L, eps, dt):
+    k1 = L(f, eps)
+    k2 = L(f + 0.5*dt*k1, eps)
+    k3 = L(f + 0.5*dt*k2, eps)
+    k4 = L(f + dt*k3, eps)
+    return f + dt*(k1 + 2*k2 + 2*k3 + k4)/6
+
+# ===================================================================
+# Utility functions for plotting 2d function
+# ===================================================================
+
+# Plotting 2d function f
+def plot_2d(f, cs):
+    N = f.shape[0]
+    dv = 2*L/N
+    v = np.mgrid[-L+dv/2:L+dv/2:dv,-L+dv/2:L+dv/2:dv]
+    v_norm = v[0]**2 + v[1]**2
+
+    fig, ax = plt.subplots()
+    cs = ax.contour(v[0], v[1], f)
+    if cs == 1:
+        ax.clabel(cs, inline=0.5)
+        
+    ax.grid(linestyle=':')
+
+    plt.show()
+
+
+# ===================================================================
+# Utility functions for reading config files
+# ===================================================================
+import json
+
+class DictionaryUtility:
+    """
+    Utility methods for dealing with dictionaries.
+    """
+    @staticmethod
+    def to_object(item):
+        """
+        Convert a dictionary to an object (recursive).
+        """
+        def convert(item):
+            if isinstance(item, dict):
+                return type('jo', (), {k: convert(v) for k, v in item.items()})
+            if isinstance(item, list):
+                def yield_convert(item):
+                    for index, value in enumerate(item):
+                        yield convert(value)
+                return list(yield_convert(item))
+            else:
+                return item
+
+        return convert(item)
+
+    def to_dict(obj):
+        """
+         Convert an object to a dictionary (recursive).
+         """
+        def convert(obj):
[…]

### 2018-08-15 src/fast_spec_col_2d.py
Add new config files, refactor codes

@@ -7,89 +7,150 @@ from math import pi
 
 class FastSpectralCollison2D:
     
-    def __init__(self, e, gamma, L, N, R, N_R, M=8):
-        gamma, self.e = config.physical_config.gamma, config.physical_config.e
+    def __init__(self, config, e=None, N=None):
+        # import parameters from config file
+        self._gamma = config.physical_config.gamma
+        if e is None:
+            self._e = config.physical_config.e
+        else:
+            self._e = e
+
+        S = config.domain_config.S
+        self._R = 2*S
+        self._L = eval(config.domain_config.L)
+        if N is None:
+            self._N = config.domain_config.N
+        else:
+            self._N = N
+
+        self._N_R = config.quadrature_config.N_g
+        self._M = config.quadrature_config.N_sigma
+
+        self._dv = None
+        self._v = None
+        self._v_norm = None
+
+        self._fftw_plan()
+        self._precompute()
+
+    @property
+    def dv(self):
+        if self._dv is None:
+            self._dv = 2*self._L/self._N
+        return self._dv
+
+    @property
+    def v(self):
+        if self._v is None:
+            self._v = np.arange(-self._L + self.dv/2, self._L+self.dv/2, self.dv)
+        return self._v
+    
+    @property
+    def v_norm(self):
+        if self._v_norm is None:
+            self._v_norm = (self.v**2)[:,None] + self.v**2
+        return self._v_norm
+
+    @property
+    def N(self):
+        return self._N
+
+    @property
+    def e(self):
+        return self._e
+
+    @property
+    def fft2(self):
+        return self._fft2
+
+    @property
+    def ifft2(self):
+        return self._ifft2
+           
+    def col_full(self, f):
+        # fft of f
+        f_hat = self._fft2(f)
+        # convolution and quadrature
+        Q = self._s_w*np.sum(self._r_w*(self._F_k_gain - self._F_k_loss)
+                            *self._fft4(self._ifft4(self._exp*f_hat[...,None,None])*f[...,None,None]), axis=(-1, -2))
+        return np.real(self._ifft2(Q))
+    
+    def col_sep(self, f):
[…]

### 2018-08-15 src/fast_spec_col_3d.py

@@ -3,104 +3,162 @@
 import numpy as np
 import pyfftw
 from math import pi
+from utility import sinc
 
-EPS = 1e-8
-def sinc(x):
-    return np.sin(x+EPS)/(x+EPS)
+class FastSpectralCollison3D:    
+    def __init__(self, config, e=None, N=None):
+        self._gamma = config.physical_config.gamma
+        if e is None:
+            self._e = config.physical_config.e
+        else:
+            self._e = e
 
-class FastSpectralCollison3D:
+        S = config.domain_config.S
+        self._R = 2*S
+        self._L = eval(config.domain_config.L)
+        if N is None:
+            self._N = config.domain_config.N
+        else:
+            self._N = N
+
+        self._N_R = config.quadrature_config.N_g
+        self._sigma = np.loadtxt(config.quadrature_config.sigma)
+
+        self._dv = None
+        self._v = None
+        self._v_norm = None
+
+        self._precompute()
+        self._fftw_plan()
+
     
-    def __init__(self, e, gamma, L, N, R, N_R, char='ss005.00012.txt'):
-        # legendre quadrature
-        x, w = np.polynomial.legendre.leggauss(N_R)
-        r = 0.5*(x + 1)*R
-        self.r_w = 0.5*R*w
-        # spherical points and weight
-        sigma = np.loadtxt(char)
-        self.s_w = 4*pi/sigma.shape[0]
-        # integral on unit sphere using quadrature
-        theta, w_theta = np.polynomial.legendre.leggauss(M_sigma[0])
-        theta = 0.5*(theta + 1)*2*pi
-        w_theta = 0.5*2*pi*w_theta
-        phi, w_phi = np.polynomial.legendre.leggauss(M_sigma[1])
-        phi = 0.5*(phi + 1)*pi
-        w_phi = 0.5*pi*w_phi
-        # index
-        k = np.fft.fftshift(np.arange(-N/2,N/2))
-        # dot with index
-        rkg = (k[:,None,None,None]*sigma[:,0] 
-                    + k[:,None,None]*sigma[:,1] 
-                    + k[:,None]*sigma[:,2])[...,None]*r
-        # norm of index
-        k_norm = np.sqrt(k[:,None,None]**2 + k[:,None]**2 + k**2)
-        # gain kernel
-        self.F_k_gain = 4*pi*r**(gamma+2)*(np.exp(0.25*1j*pi*(1+e)*rkg/L)
[…]

### 2018-08-15 src/test_2D.py

@@ -1,0 +1,49 @@
+import numpy as np
+from utility import RK3
+from fast_spec_col_2d import FastSpectralCollison2D
+
+
+class TestModule2D:
+    def __init__(self, config, initial, N, e, eps=0):
+        S = config.domain_config.S
+        self.L = eval(config.domain_config.L)
+        self.N = config.domain_config.N
+        self.e, self.eps = e, eps
+        self.f0 = initial
+        self.Q = []
+        for n in N:
+            self.Q.append(FastSpectralCollison2D(config, e=e, N=n))
+            
+    def ext_T(self, N_index, tfinal):
+        Q = self.Q[N_index]
+        v, v_norm, dv = Q.v, Q.v_norm, Q.dv
+        T_0 = 0.5*np.sum(self.f0(v)*v_norm)*dv**2
+#         T_0 = 1.425
+        return (T_0 - 8*self.eps/(1-self.e**2))*np.exp(-(1-self.e**2)*tfinal/4) + 8*self.eps/(1-self.e**2)
+    
+    def solve(self, RK, dt, tfinal, N_index, method):
+        Q = self.Q[N_index]
+        v, v_norm, dv = Q.v, Q.v_norm, Q.dv
+        f_hat = Q.fft2(self.f0(v))
+        
+        for _ in np.arange(0, tfinal, dt):
+            f_hat = RK(f_hat, getattr(Q, 'col_heat_hat_'+ method), self.eps, dt)    
+        f = np.real(Q.ifft2(f_hat))
+        return f, 0.5*np.sum(f*v_norm)*dv**2
+            
+    def dt_test(self, Dt, method='sep', RK=RK3, N_index=3, tfinal=2):
+        num_T, num_f = [], []
+        for dt in Dt:
+            f, T = self.solve(RK, dt, tfinal, N_index, method)
+            num_T.append(T)
+            num_f.append(f)
+        return np.abs(np.asarray(num_T) - self.ext_T(N_index, tfinal)), num_f, num_T
+    
+    def N_test(self, dt, method='sep', RK=RK3, tfinal=2):   
+        num_T, num_f, ext_T = [], [], []
+        for N_index in [0,1,2,3,4]:
+            f, T = self.solve(RK, dt, tfinal, N_index, method)
+            num_T.append(T)
+            num_f.append(f)
+            ext_T.append(self.ext_T(N_index, tfinal))
+        return np.abs(np.asarray(num_T) - np.asarray(ext_T)), num_f, num_T

### 2018-08-15 src/utility.py

@@ -1,3 +1,5 @@
+import numpy as np
+import matplotlib.pyplot as plt
 # ===================================================================
 # Utility functions for different time marching schemes
 # ===================================================================
@@ -27,17 +29,12 @@ def RK4(f, L, eps, dt):
 # ===================================================================
 
 # Plotting 2d function f
-def plot_2d(f, cs):
-    N = f.shape[0]
-    dv = 2*L/N
-    v = np.mgrid[-L+dv/2:L+dv/2:dv,-L+dv/2:L+dv/2:dv]
-    v_norm = v[0]**2 + v[1]**2
-
+def plot_2d(f, v, cs):
+    xv, yv = np.meshgrid(v, v)
     fig, ax = plt.subplots()
-    cs = ax.contour(v[0], v[1], f)
+    cs = ax.contour(xv, yv, f)
     if cs == 1:
-        ax.clabel(cs, inline=0.5)
-        
+        ax.clabel(cs, inline=0.5)   
     ax.grid(linestyle=':')
 
     plt.show()
@@ -99,3 +96,7 @@ def get_config(config_path):
     # convert dict to object recursively for easy call
     config = DictionaryUtility.to_object(config)
     return config
+    
+EPS = 1e-8
+def sinc(x):
+    return np.sin(x+EPS)/(x+EPS)

### 2018-08-15 src/direct_col_3d.py
Add direct method

@@ -1,0 +1,103 @@
+import numpy as np
+from math import pi
+from utility import sinc
+import pyfftw
+
+class DirectCol3D(object):
+    def __init__(self, config, e=None, N=None):
+        self._config = config
+        self._gamma = config.physical_config.gamma
+        if e is None:
+            self._e = config.physical_config.e
+        else:
+            self._e = e
+
+        S = config.domain_config.S
+        self._R = 2*S
+        self._L = eval(config.domain_config.L)
+        if N is None:
+            self._N = config.domain_config.N
+        else:
+            self._N = N
+
+        self._N_R = config.quadrature_config.N_g
+
+        self._Glm = None
+        self._l = None
+        self._m = None
+        
+        self._dv = None
+        self._v = None
+        self._v_norm = None
+
+        self._fftw_plan()
+
+    def _G(self, k, m):
+        x, w = np.polynomial.legendre.leggauss(self._N_R)
+        r = 0.5*(x + 1)*self._R
+        r_w = 0.5*self._R*w
+        km = 0.25*(1 + self._e)*k[:,None] - m
+        k_m_norm = np.sqrt(km[:,None,None,:,None,None]**2 + km[:,None,None,:,None]**2 + km[:,None,None,:]**2)
+        m_norm = np.sqrt(m[:,None,None]**2 + m[:,None]**2 + m**2)
+        k_norm = np.sqrt(k[:,None,None]**2 + k[:,None]**2 + k**2)
+        return 16*pi**2*np.sum(r_w*r**(self._gamma+2)*(sinc(pi*k_m_norm[...,None]*r/self._L)*sinc(pi*0.25*(1+self._e)*k_norm[...,None,None,None,None]*r/self._L) 
+            - sinc(pi*m_norm[...,None]*r/self._L)), axis = (-1))
+
+    def _lm(self):
+        l, m = [], []
+        for i in range(self._N):
+            l.append([])
+            m.append([])
+        N_half = int(self._N/2)    
+        for i in np.arange(-N_half, N_half):
+            for j in np.arange(-N_half, N_half):
+                if i + j > N_half - 1:
+                    l[i+j-N_half].append(i+N_half)
+                    m[i+j-N_half].append(j+N_half)
+                elif i + j < -N_half:
+                    l[i+j+3*N_half].append(i+N_half)
+                    m[i+j+3*N_half].append(j+N_half)
+                else:
[…]

### 2018-08-15 src/fast_spec_col_2d.py

@@ -5,8 +5,7 @@ import pyfftw
 from scipy import special
 from math import pi
 
-class FastSpectralCollison2D:
-    
+class FastSpectralCollison2D(object):
     def __init__(self, config, e=None, N=None):
         # import parameters from config file
         self._gamma = config.physical_config.gamma

### 2018-08-15 src/fast_spec_col_3d.py

@@ -5,8 +5,9 @@ import pyfftw
 from math import pi
 from utility import sinc
 
-class FastSpectralCollison3D:    
+class FastSpectralCollison3D(object):    
     def __init__(self, config, e=None, N=None):
+        self._config = config
         self._gamma = config.physical_config.gamma
         if e is None:
             self._e = config.physical_config.e
@@ -22,8 +23,8 @@ class FastSpectralCollison3D:
             self._N = N
 
         self._N_R = config.quadrature_config.N_g
-        self._sigma = np.loadtxt(config.quadrature_config.sigma)
-
+        self._method = config.quadrature_config.method
+        
         self._dv = None
         self._v = None
         self._v_norm = None
@@ -31,7 +32,6 @@ class FastSpectralCollison3D:
         self._precompute()
         self._fftw_plan()
 
-    
     @property
     def dv(self):
         if self._dv is None:
@@ -71,7 +71,7 @@ class FastSpectralCollison3D:
         # fft of f
         f_hat = self._fft3(f)
         # convolution and quadrature
-        Q = self._s_w*np.sum(self._r_w*(self._F_k_gain - self._F_k_loss)
+        Q = np.sum(self._r_w*self._s_w[:,None]*(self._F_k_gain - self._F_k_loss)
                             *self._fft5(self._ifft5(self._exp*f_hat[...,None,None])*f[...,None,None]),axis=(-1, -2))
         return np.real(self._ifft3(Q))
     
@@ -79,7 +79,7 @@ class FastSpectralCollison3D:
         # fft of f
         f_hat = self._fft3(f)
         # gain term
-        Q_gain = self._s_w*np.sum(self._r_w*self._F_k_gain
+        Q_gain = np.sum(self._s_w[:,None]*self._r_w*self._F_k_gain
                                  *self._fft5(self._ifft5(self._exp*f_hat[...,None,None])*f[...,None,None]),axis=(-1, -2))
         # loss term
         Q_loss = 4*pi*np.sum(self._r_w*self._F_k_loss
@@ -113,15 +113,25 @@ class FastSpectralCollison3D:
         x, w = np.polynomial.legendre.leggauss(self._N_R)
         r = 0.5*(x + 1)*self._R
         self._r_w = 0.5*self._R*w
+        if self._method == "spherical_design":
         # spherical points and weight
[…]

### 2018-08-15 src/direct_col_3d.py
Change direct_col_3d

@@ -32,6 +32,32 @@ class DirectCol3D(object):
 
         self._fftw_plan()
 
+    @property
+    def dv(self):
+        if self._dv is None:
+            self._dv = 2*self._L/self._N
+        return self._dv
+
+    @property
+    def v(self):
+        if self._v is None:
+            self._v = np.arange(-self._L + self.dv/2, self._L+self.dv/2, self.dv)
+        return self._v
+    
+    @property
+    def v_norm(self):
+        if self._v_norm is None:
+            self._v_norm = (self.v**2)[:,None,None] + (self.v**2)[:,None] + self.v**2
+        return self._v_norm
+
+    @property
+    def N(self):
+        return self._N
+
+    @property
+    def e(self):
+        return self._e
+
     def _G(self, k, m):
         x, w = np.polynomial.legendre.leggauss(self._N_R)
         r = 0.5*(x + 1)*self._R

### 2018-09-10 paper/paper_rev/fast_spec_inelastic_rev.tex

--- 2018-07-08 paper/paper_submit_ver/fast_spec_inelastic.tex
@@
-\documentclass[review, times]{elsarticle}
+\documentclass[review,times]{elsarticle}
 
@@
 
-\usepackage{color}
+\usepackage[table]{xcolor}
 \modulolinenumbers[5]
+
+\usepackage{caption}
 
@@
 Now assuming $f$ is a solution to equation \eqref{inBoltz}, the conservation of mass and momentum is straightforward by taking $\varphi = 1$ and $v$ respectively in the weak form \eqref{weak22}, i.e.,
-\begin{equation}
+\begin{equation}\label{conservation}
 \rho \equiv \rho_0, \quad  m\equiv m_0,
@@
 
-\subsubsection{3D inelastic Maxwell molecules}
+\subsubsection{3D inelastic Maxwell molecules \color{blue} --- isotropic solution}
 
@@
   \end{tabu}
-  \caption{Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule for $e=0.2$.
+  \caption{Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule {\color{blue} with isotropic initial} for $e=0.2$.
 Errors shown here are $|T_\text{num} - T_\text{ref}|$ at $t_\text{final}=1$.
@@
   \end{tabu}
-  \caption{Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule for $e=0.2$.
+  \caption{Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule {\color{blue} with isotropic initial} for $e=0.2$.
 Errors shown here are $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$ at $t_\text{final}=1$.
@@
   \includegraphics[width = .8\linewidth]{CPU_time}
-  \caption{Same numbers as in Table~\ref{table 5} plotted in a log-log scale.}
+  \caption{Same numbers as in Table~\ref{table 5} plotted in a log-log scale.
+{\color{blue} This test is performed using Python (numpy and scipy modules for array operations; pyfftw module (a Python wrapper of FFTW)for computing FFTs) on a $10$ CPUs workstation (Intel(R) Xeon(R) CPU E5-2660 v3 @ 2.60GHz, $8$ threads per FFTW operation.)}}
   \label{CPUtime}
 \end{figure}
+
+{\color{blue} \subsubsection{3D inelastic Maxwell molecules  --- anisotropic solution}
+
+In previous subsection we perform a convergence test in $M_\text{sph}$ (compared with direct method and the analytical solution of the macroscopic quantity $T$) using an isotropic solution.
+This test shows us by using spherical design we can achieve very high accuracy of the integral on the 3D unit sphere surface, which is a key step in our algorithm for 3D case.
+In order to further investigate the performance of spherical design and to illustrate its necessity in the algorithm, an example with anisotropic initial is presented in this subsection.
+
+For the sake of comparing with isotropic case, this test is done use exact the same configuration as in previous subsection except the initial condition:
+\begin{equation}\label{3D_anisotropic}
+  f_0(v) = \frac{4}{5\pi^{\frac{3}{2}}}\left(\exp(-\sqrt[3]{16}|v-2|^2) + \exp(-|v+0.5|^2)\right).
+\end{equation}
+We still set $e=0.2$, $t_\text{final} = 1$, $\Delta t = 0.01$, $N = 32$, and $N_{\rho}=30$ and first use the macroscopic quantity $T$ as our measurement.
+The results are shown in Table~\ref{3D_T_conv_anisotropic}.
+We can find both the convergence rate and accuracy are not as good as in isotropic case.
+It can only achieves an $O(10^{-4})$ error even using $192$ points of spherical design.
+One can also see this result from Table~\ref{3D_direct_conv_anisotropic} and Table~\ref{table 4}.
+Despite this result, we consider it as acceptable since the problem itself (numerical approximation of an integral on 3D unit sphere) is highly nontrivial, and we believe spherical design is so far the best method to deal with such problem.
+We illustrate this statement by performing another test using the standard gauss quadrature method which is also the method used in~\cite{WU201566} for the computation of the integral on 3D unit sphere, shown in Table~\ref{3D_quadrature_conv_anisotropic}.
+Under roughly the same number of quadrature points on sphere used \big((\# of $\theta$)$\times$(\# of $\phi$)\big), this quadrature method has a very bad performance: slow convergence and low accuracy.
+}
+
+\begin{table}[htpb]
+  \color{blue}
+  \centering
+  \begin{tabu} to 0.8\linewidth {X[1,c] X[2,c]}
+    \toprule
+    $M_{\text{sph}}$ & $|T_\text{num} - T_\text{ref}|$\\
+    \midrule
+    6 & 2.661413377541729e-01 \\
+    12 & 1.6773023124809094e-01 \\
+    32 & 1.0297536867238799e-01 \\
+    48 & 2.584202650005074e-02 \\
+    70 & 2.422376071694199e-02 \\
+    94 & 3.058799838901116e-03 \\
+    120 & 4.413127388460447e-03 \\
+    156 & 1.220110856786194e-03 \\
+    192 & 3.96517639616345e-04 \\
+    \bottomrule
+  \end{tabu}
+  \captionsetup{labelfont={color=blue}}
+  \caption{\color{blue} Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule with anisotropic initial for $e=0.2$.
+Errors shown here are $|T_\text{num} - T_\text{ref}|$ at $t_\text{final}=1$.
+$\Delta t = 0.01$, $N = 32$, $N_{\rho}=30$, $R=7$, $L=7.72$.}
+  \label{3D_T_conv_anisotropic}
+\end{table}
+
+\begin{table}[htpb]
+  \color{blue}
+  \centering
+  \begin{tabu} to 0.8\linewidth {X[1,c] X[2,c]}
+    \toprule
+    $M_{\text{sph}}$ & $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$\\
+    \midrule
+ 	  6 & 1.4674566108660258e-02 \\
+    12 & 9.532325835757156e-03 \\
+    32 & 5.92707599850977e-03 \\
+    48 & 1.952601851310574e-03 \\
+    70 & 1.43696494086e-03 \\
+    94 & 4.354500695785647e-04 \\
+    120 & 3.534377201557734e-04 \\
+    156 & 9.811152052724729e-05 \\
+    192 & 3.747408101774895e-05 \\ 
+    \bottomrule
+  \end{tabu}
+  \captionsetup{labelfont={color=blue}}
+  \caption{\color{blue} Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule with anisotropic initial for $e=0.2$.
+Errors shown here are $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$ at $t_\text{final}=1$.
+$\Delta t = 0.01$, $N = 32$, $N_{\rho}=30$, $R=7$, $L=7.72$.}
+  \label{3D_direct_conv_anisotropic}
+\end{table}
+
+\begin{table}[htpb]
+  \color{blue}
+  \centering
+  \begin{tabu} to 0.8\linewidth {X[1,c] X[2,c]}
+    \toprule
+    \# of $(\theta, \phi)$ & $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$\\
+    \midrule
+    (2,3) & 9.301850859331023e-03 \\
+    (3,4) & 1.4164542178285952e-02 \\
+    (4,8) & 2.148725985732372e-02 \\
+    (6,8) & 3.952940827407407e-03 \\
+    (7,10) & 3.3540095954422665e-03 \\
+    (10,10) & 3.405850043869026e-03 \\
+    (10,12) & 3.016912442998702e-03 \\
+    (11,15) & 1.227818352872178e-03 \\
+    (12,16) & 1.2310868626689182e-03 \\
+    \bottomrule
+  \end{tabu}
+  \captionsetup{labelfont={color=blue}}
+  \caption{\color{blue} Convergence test of gauss quadrature method of the 3D Maxwell molecule with anisotropic solution for $e=0.2$.
+Errors shown here are $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$ at $t_\text{final}=1$.
+$\Delta t = 0.01$, $N = 32$, $N_{\rho}=30$, $R=7$, $L=7.72$.}
+  \label{3D_quadrature_conv_anisotropic}
+\end{table}
 
@@
 
-We take the same initial data \eqref{3Dbkw} as in the previous example but for hard sphere molecules, that is, consider the collision kernel $B_{\sigma} = \frac{1}{4\pi}|g|$.
+{\color{red} We perform two simulations regarding different initial data: The first one we take the same initial data \eqref{3Dbkw} as in the previous example; the second one we use the following initial data which is a Maxwellian with nonzero bulk velocity,
+\begin{equation}\label{3D_maxwellian}
+  f_0(v) = \frac{\rho_0}{(2\pi T_0)^{3/2}}e^{-(v - u_0)^2},
+\end{equation}
+where $\rho_0 = 1$, $T_0 = 2$ and $u_0 = (0.5, -0.5, 0)^T$.
+From~\eqref{conservation} we know $rho$ and $u$ with keep unchanged in time.
+Both simulations are performed} for hard sphere molecules, that is, consider the collision kernel $B_{\sigma} = \frac{1}{4\pi}|g|$.
 Also we set $\varepsilon=0$ in equation (\ref{inBoltz}) (the Haff's law is for unheated case, i.e., the homogeneous cooling state).
 We numerically evaluate the temperature up to time $t_\text{final}=3$ with two different values of restitution coefficient $e=0.2$ and $0.8$.
-The time evolution of $T$ is shown in Figure~\ref{Haff_cooling}, from which one can see that the numerical temperature decays as
+The time evolution of $T$ is shown in Figure~\ref{Haff_cooling} and Figure~\ref{Haff_cooling_maxwellian}, from which one can see that the numerical temperature decays as
 \begin{equation}
@@
 \end{equation}
-where the constant $C_0$ depends on the value of $e$. This is in perfect agreement of the Haff's formula (see Section 4.1 of \cite{NE}).
+where the constant $C_0$ {\color{red} depending on the value of $e$ is chosen by using least squares fitting}. This is in perfect agreement of the Haff's formula (see Section 4.1 of \cite{NE}).
 
@@
   \includegraphics[width = .8\linewidth]{Haff's_cooling}
-  \caption{Haff's cooling law. $\Delta t=0.01$, $N=32$, $N_{\rho}=30$, $M_{\text{sph}}=32$, $R=7$, $L=7.72$.}
+  \caption{Haff's cooling law {\color{red} with~\eqref{3Dbkw} as initial data}. $\Delta t=0.01$, $N=32$, $N_{\rho}=30$, $M_{\text{sph}}=32$, $R=7$, $L=7.72$.}
   \label{Haff_cooling}
+\end{figure}
+
+\begin{figure}[htp!]
+  \centering
+  \includegraphics[width = .8\linewidth]{Haff_cooling_maxwellian}
+  \captionsetup{labelfont={color=blue}}
+  \caption{\color{blue} Haff's cooling law {\color{red} with~\eqref{3D_maxwellian} as initial data}. $\Delta t=0.01$, $N=32$, $N_{\rho}=30$, $M_{\text{sph}}=32$, $R=7$, $L=7.72$.}
+  \label{Haff_cooling_maxwellian}
 \end{figure}

### 2018-09-10 paper/archive/response/Response to Reviewers.tex
New response and paper_rev.

--- 2018-08-13 paper/response/Response to Reviewers.tex
@@
 \usepackage{cases}
-\usepackage{color}
+\usepackage{xcolor}
 
 \newcommand{\jh}[1]{\textcolor{red}{JH: #1}}
+\newcommand{\mz}[1]{\textcolor{blue}{MZ: #1}}
 
@@
 
-\item[{\bf A1}]
+\item[{\bf A1}] \mz{The ``direct'' or ``standard'' spectral method indeed can handle general form of collision kernels.
+However, the method by Wu et al., namely the proposed FAST spectral method based on Carleman representation CANNOT work for general collision kernels.
+We already checked that paper and believe that, only in the case $B(|x|, |y|)$ is a SEPARABLE funtion of $|x|$, $|y|$, the proposed method will work (all the examples shown in that paper are in $B(|x|, |y|) = |x|^a |y|^b$ this form).
+Of course one can argue that you can use this separable function to APPROXIMATE ``general'' collision kernel, however, we suspect this is good for $B(|x|, |y|)$ being an arbitrary function.
+On the contrary, our fast method does not have this constrain since we do not reply on Carleman representation and can work for ``TRUE'' general collision kernel without any appoximation of it.}
 
@@
 
-\item[{\bf A2}]  \jh{Add a comparsion}
+\item[{\bf A2}]  \jh{Add a comparsion} \mz{New numercial test is added to illustrate this conclusion, see Table~8.}
 
@@
 
-\item[{\bf A3}]
+\item[{\bf A3}] \mz{We have explanations for this at line 130 and Section 4. The conservation can be achieved using other approachn (e.g., adding a correction step).}
 
@@
 
-\item[{\bf A4}]
+\item[{\bf A4}] \mz{Added accordingly, see Fig. 2}
 
@@
 
-\item[{\bf A5}]
+\item[{\bf A5}] \mz{This is no reason that this method will fail for spatially in-homogeneous problem......}
 
@@
 
-\item[{\bf A6}]
+\item[{\bf A6}] \mz{...}
 
@@
 $10^{-5}$).
-I think [16] does not contain a CPU time, please double check.}
+I think [16] does not contain a CPU time, please double check (\mz{Confirmed, at least not comparable}).}
 
@@
 
-\item[{\bf A7}]
+\item[{\bf A7}] \mz{They are just standard Gauss-Legendre quardrature weights which are included in any numerical textbook.}
 
@@
 
-\item[{\bf A10}] \jh{Maybe include it in Section 4.1.1 as this is quite easy.}
+\item[{\bf A10}] \jh{Maybe include it in Section 4.1.1 as this is quite easy \mz{Of course we checked our code can recover the elastic case otherwise it means our code is wrong....}.}
 
@@
 
-\item[{\bf A11}]
+\item[{\bf A11}] \mz{This is a good question.
+But since it is just a heating bath with a very small $\varepsilon = 10^{-6}$ in our test, we believe, at least numerical results show that we do not need to change the truncation.}
 
@@
 Later in Section 4.1.2 we did consider an anisotropic solution (see equation (50)) and observed quite different behavior.
-\jh{Why didn't we consider an anisotropic solution in 3D?}
+\mz{We have added an anisotropic solution case in 3D, see the new subsection~4.2.2.}\jh{Why didn't we consider an anisotropic solution in 3D?}
 

### 2018-09-10 paper/response/Response to Reviewers.tex
New response and paper_rev.

@@
 
-\noindent {We would like to thank the anonymous reviewers for their valuable suggestions and comments, which greatly help us improve the quality and clarity of the manuscript.
-In the following, {\bf Q} denote the questions or suggestions raised by the reviewers, and {\bf A} refer to our answers.
+\noindent {We would like to thank the Reviewers for their valuable suggestions and comments, which greatly help us improve the quality and clarity of the manuscript.
+In the following, {\bf Q} denote the questions/comments raised by the Reviewers, and {\bf A} refer to our answers.
 }
@@
 \bigskip
+
+In the revised manuscript, all changes in response to Reviewer \#1 have been highlighted in \textcolor{blue}{Blue}. Below is a detailed response to each comment.
 
@@
 
-\item[{\bf A1}]
+\item[{\bf A1}] In our original manuscript, we wrote ``the method by Wu et al.
+only {READILY} works for 2D Maxwell ....
+To deal with other types of interactions, additional modification of the kernel or parameter fitting is necessary, as was done in [17]." We never said that the method cannot treat more general kernels.
+It can be modified to do so based on parameter fitting.
+But we do want to mention that this additional modification/fitting introduces an error that is hard to quantify.
+Our method, on the other hand, is based on a different representation and can directly apply to general kernel without any approximation of it.
+
+%\mz{The ``direct'' or ``standard'' spectral method indeed can handle general form of collision kernels.
+However, the method by Wu et al., namely the proposed FAST spectral method based on Carleman representation CANNOT work for general collision kernels.
+We already checked that paper and believe that, only in the case $B(|x|, |y|)$ is a SEPARABLE funtion of $|x|$, $|y|$, the proposed method will work (all the examples shown in that paper are in $B(|x|, |y|) = |x|^a |y|^b$ this form).
+Of course one can argue that you can use this separable function to APPROXIMATE ``general'' collision kernel, however, we suspect this is good for $B(|x|, |y|)$ being an arbitrary function.
+On the contrary, our fast method does not have this constrain since we do not reply on Carleman representation and can work for ``TRUE'' general collision kernel without any appoximation of it.}
 
@@
 
-\item[{\bf A2}]  \jh{Add a comparsion}
+\item[{\bf A2}] The optimal performance of the spherical design can be found in reference [24].
+For better illustration, a comparison of the {\it spherical design} v.s.
+{\it Gauss quadrature} is provided in the revised version, see Tables 7 and 8, where one can see that using roughly the same number of quadrature points on the sphere, the former provides better accuracy and convergence than the latter.
 
@@
 
-\item[{\bf A3}]
+\item[{\bf A3}] This is a valid point.
+In fact, we have checked both versions in our numerical tests.
+The conclusion we obtained is that computing both terms separately provides better accuracy, even in capturing the macroscopic quantities such as the temperature (see in particular Table 2).
+Some explanations are given at the end of Section 4.1.
+We also would like to point out that the spectral method itself (without any additional approximation) cannot preserve the moments except the mass.
+Therefore, if the conservation (in the inelastic case, only mass and momentum are conserved) is strictly required, one can do a correction step [12] as post processing.
 
@@
 
-\item[{\bf A4}]
+\item[{\bf A4}] These are provided now in Table 5.
 
@@
 
-\item[{\bf A5}]
+\item[{\bf A5}] We have carefully validated our solver in both 2D and 3D for the spatially homogeneous case.
+In fact, more tests are provided in the revised version.
+We believe the solver should work well in the inhomogeneous case.
+If there exist such examples that the deterministic method fails due to the instability of the homogeneous part, we would be happy to know the details.
 
@@
 
-\item[{\bf A6}]
-
+\item[{\bf A6}] Mathematically, the Enskog equation is not justified.
+It is more of a phenomenological model used for engineering purpose.
+Also there exist many different versions of the Enskog equation in the literature.
+For these reasons, we chose not to discuss the Enskog equation to avoid ambiguities.
+On the other hand, our method readily generalizes to the Enskog model discussed in [16], since there the spatial variable only plays the role as a parameter.
 \end{itemize}
@@
 
-{\bf Critque.} {\it In the reviewer opinion, this work is not a true novelty, because it arrives more than three years after the fast spectral method for the Wu et al.
+{\bf Critique.} {\it In the reviewer opinion, this work is not a true novelty, because it arrives more than three years after the fast spectral method for the Wu et al.
 paper [16] (which was in itself a refinement of a paper from the same group about multiple species Boltzmann equation).
@@
 
-First of all, this reviewer's comment ``This method is almost the same than the one introduced in the manuscript" is absolutely not true!
-The method in [16] is based on the so-called Carleman representation of the collision operator.
-This representation is first used in Mouhot and Pareschi (2006) to derive a fast Fourier method for the classical (elastic) Boltzmann collision operator and is in fact the basis of the method proposed in [16].
+{\bf Our response.} First of all, this reviewer's comment ``This method is almost the same than the one introduced in the manuscript" is absolutely not true.
+The method in [16] is based on the so-called Carleman representation of the collision operator, which is first used in [17] (citation number in the revised version) to derive a fast spectral method for the classical (elastic) Boltzmann collision operator and is in fact the basis of the method in [16].
 Our method is based on the $\sigma$-representation of the collision operator.
 The two representations are equivalent at the continuous level.
-However, they become quite different when numerical approximation is applied, for example, the truncation of the integral is done on different variables, in our method, it is $g=v-v_*$, the relative velocity is truncated, while in [16], it is $x=(|g|\sigma-g)/2$ and $y=-g-x$ are truncated; furthermore, our numerical quadrature is applied on $g$, while in [16], it is $x$ that is discretized.
+However, they become quite different when numerical approximation is applied, for example, the truncation of the integral is done on different variables, in our method, it is $g=v-v_*$, the relative velocity that is truncated, while in [16], it is $x=(|g|\sigma-g)/2$ and $y=-g-x$ that are truncated; furthermore, our numerical quadrature is applied to $g$, while in [16], it is $x$ that is discretized.
 These will of course result in different approximations of the collision operator.
-Another difference is that, our method makes no assumptions on the collision kernel and applies to the general form $B_{\sigma}(|g|,\sigma\cdot \hat{g})$.
-The method in [16], although can treat general kernels, is based on a modification of the kernel so that it is in the separable form.
-In view of this, our method does not need any special treatment/approximation of the kernel and is much easier to implement.
-Moreover, we propose to use the sophisticated quadrature, spherical design, on the sphere, in contrast to the tensor product based Gauss quadrature in [16], which results in much better accuracy as demonstrated in our paper.
+Another key difference is that, by using the $\sigma$-representation our method requires no assumption on the collision kernel and applies directly to the general form $B_{\sigma}(|g|,\sigma\cdot \hat{g})$.
+The method in [16], although can treat general kernels, is based on a modification of the kernel so that it is in the separable form, hence FFT can be used to accelerate the computation.
+The error introduced in this procedure is not clear.
+Moreover, we propose to use the sophisticated quadrature {\it spherical design} on the sphere, in contrast to the tensor product based Gauss quadrature in [16].
+The former provides much better accuracy as demonstrated in our paper.
+Last but not least, we tested carefully the ``full" and ``separated" approaches and observed that the ``separated" approximation can actually yield much better results even when capturing the moments (we in fact leverage the smoothing property of the gain term).
+The method in [16] discretizes the gain and loss terms simultaneously and the accuracy of doing them separately has never been evaluated.
 
-\jh{Our method provides better accuracy compared to that in [16], do the same numerical test in Section 4.1 of [16] ($10^{-8}$ v.s.
+Regarding the comparison, since our method directly accelerates the original spectral method in [11], it makes a lot of sense to compare the accuracy and efficiency of these two methods.
+As mentioned already, the method in [16] is based on a different approximation than ours (even the criteria of the truncation of the computational domain are different), hence we don't see a fair way to compare them directly.
+Furthermore, careful numerical validation (convergence test with respect to different discretization parameters) in [16] is quite limited except for one example of 2D Maxwell molecules, neither the CPU time is provided, which make it very hard to do the comparison.
+
+%\jh{Our method provides better accuracy compared to that in [16], do the same numerical test in Section 4.1 of [16] ($10^{-8}$ v.s.
 $10^{-5}$).
-I think [16] does not contain a CPU time, please double check.}
+I think [16] does not contain a CPU time, please double check (\mz{Confirmed, at least not comparable}).}
 
@@
 4, l.
-50)" is absolutely wrong as already explained in the previous paragraph.
-Both methods are different no matter $e$ is a constant or not.
-The statement in page 4, line 50 concerns the difference between the Boltzmann and Enskog operators and it is by no means related to the numerical methods!
+50)" is absolutely wrong as already explained above.
+The two methods are different no matter $e$ is a constant or not.
+The statement on page 4, line 50 concerns the difference between the Boltzmann and Enskog operators and has nothing to do with the numerical methods.
+
+\bigskip
+In the revised manuscript, all changes in response to Reviewer \#2 have been highlighted in \textcolor{red}{Red}. Below is a detailed response to each comment.
 
@@
 \item[{\bf A1}] The weak form of the inelastic collision operator is never of an issue and is consistent in the literature.
-However, one can find different strong forms from various sources and it is because of this reason, in [19], the authors (in particular, J.
+However, one can find different strong forms from various sources especially for the case of non-constant restitution coefficient and it is because of this reason, in [20], the authors (in particular, J.
 A.
@@
 
-\item[{\bf A2}] We agree that the physically relevant case is the hard sphere molecules but for theoretical purpose and numerical validation, other types of molecules, for example the Maxwell molecules, are also useful.
+\item[{\bf A2}] We agree that the physically relevant case is the hard sphere molecules but for theoretical purpose and numerical validation, other types of molecules, for example, the Maxwell molecules, are also useful.
 In fact, the reference [16] also considers the variable hard sphere kernel (see Section 4.1 in their paper).
@@
 
-\item[{\bf A3}] $C_0=1/(4\pi)$ as mentioned at the bottom of page 21.
+\item[{\bf A3}] This is described in many books on granular materials and requires several levels of approximations, see for example Section 4.1 in [27].
 
@@
 
-\item[{\bf A4}] \jh{Refer to [11] or briefly add the discussion to the paper.}
+\item[{\bf A4}] This can be seen from the strong form (6) of the collision operator and is discussed clearly in [11].
+As the argument is quite standard, we chose not to repeat the details.
+It is also worth to mention that the choice of the computational domain in the Carleman representation is determined differently and a larger $L$ is required.
 
@@
 
-\item[{\bf A5}] \jh{Mention that this is a different method....}
+\item[{\bf A5}] Our method starts from the $\sigma$-representation and the fast spectral method in Mouhot and Pareschi ([17] in the revised version) is based on Carleman representation.
+Since two methods follow different approximations, it would cause confusion to mention [17] in the description of our method.
+We have added it to the Introduction part when discussing the method in [16].
 
@@
 
-\item[{\bf A6}]
+\item[{\bf A6}] The integrand in (39) contains a complex exponential whose exponent varies on the order of $N$ (the magnitude of $m$ is $\sim O(N)$).
+According to the well-known Shannon-Nyquist sampling theorem, one needs at least $N$ points to approximate an oscillatory function like this.
 
@@
 
-\item[{\bf A7}]
+\item[{\bf A7}] They are just standard Gauss-Legendre quadrature points and weights that can be found in many textbooks on numerical analysis.
 
@@
 
-\item[{\bf A8}]
+\item[{\bf A8}] We don't think there is any inconsistency in the discussion of the numerical complexity and storage requirement.
 
@@
 
-\item[{\bf A9}] The Fourier spectral method does not have the conservation.
-We did not check the mass and momentum but the extensive results presented regarding the temperature demonstrate that the method can capture the moments fairly accurately.
+\item[{\bf A9}] It is known that the Fourier spectral method does not have the conservation.
+We did not check specifically on mass and momentum but extensive results presented regarding the temperature demonstrate that the method can capture the moments fairly accurately.
 
@@
 
-\item[{\bf A10}] \jh{Maybe include it in Section 4.1.1 as this is quite easy.}
+\item[{\bf A10}] We obviously checked this case. It is also clear from the derivation that the method reduces automatically to the elastic collision solver that was already considered in [14].
 
@@
 
-\item[{\bf A11}]
+\item[{\bf A11}] We have chosen a relatively large computational domain $L\approx 2.2 S$, where $S$ is the compact support of the function.
+In addition, the diffusion parameter $\varepsilon$ is very small ($10^{-6}$), hence the domain should be safe.
+The fact that numerical errors of our method when comparing with the analytical temperature exhibit spectral decay (see Table 2 for example) is a clear evidence.
+
+%\mz{This is a good question.
+But since it is just a heating bath with a very small $\varepsilon = 10^{-6}$ in our test, we believe, at least numerical results show that we do not need to change the truncation.}
 
@@
 
-\item[{\bf A12}] \jh{There is no rigorous proof except the numerical evidence.
-However, we do believe this is attributed to the smoothing property of the gain term.
-Cite the paper ``Cooling Process for Inelastic Boltzmann Equations for Hard Spheres, Part II: Self-Similar Solutions and Tail Behavior" by Mischler and Mouhot for the first regularity proof of the gain term of the inelastic collision operator.
-In fact, it is mentioned in this paper the regularity result is similar to the elastic case and the proof follows the lines in the elastic case....
-See the discussion in Section 1.6 of that paper.}
+\item[{\bf A12}] There is no rigorous proof except the numerical evidence.
+However, we do believe this is attributed to the smoothing property of the gain term (one can check the accuracy of the gain and loss terms separately and will see that the same discretization applied to the loss term would result in more error).
+We have cited the paper by Mischler and Mouhot ([26]) for the smoothing effect of the inelastic operator.
+In fact, it is mentioned in this paper the regularity result is similar to the elastic case and the proof follows the lines there, see the discussion in Section 1.6 of [26].
 
@@
 
-\item[{\bf A13}] \jh{I guess what he means is to give explicitly the formulas for $\tilde{v}$, $\tilde{v}_*$ and $J$.
-We can do this but need to mention that this is only possible when $e$ is a constant.}
+\item[{\bf A13}] These formulas can be given explicitly only for the case of constant restitution coefficient.
+Since our method works for general $e$ (which does not have to be constant), we chose not to provide them to avoid ambiguity.
 
@@
 Later in Section 4.1.2 we did consider an anisotropic solution (see equation (50)) and observed quite different behavior.
-\jh{Why didn't we consider an anisotropic solution in 3D?}
+In the revised version, we have also added the anisotropic solutions in 3D, see both sections 4.2.2 and 4.2.3.
 

### 2018-09-10 paper/response/Response to the Editor.tex
New response and paper_rev.

\documentclass[11pt]{article}
\usepackage{latexsym}
\usepackage{amssymb,amsbsy,amsmath,amsfonts,amssymb,amscd}
\usepackage{mathrsfs}
\usepackage{epsfig, graphicx}

\setlength{\oddsidemargin}{0mm} \setlength{\evensidemargin}{0mm}
\setlength{\topmargin}{5mm} \setlength{\textheight}{22cm}
\setlength{\textwidth}{17cm}
\renewcommand{\baselinestretch}{1.0}
\parindent 10pt

\usepackage[margin=1.0in]{geometry}

\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{graphicx}
\usepackage{subfigure}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsmath}
\usepackage{bm}
\usepackage{cases}
\usepackage{xcolor}

\newcommand{\jh}[1]{\textcolor{red}{JH: #1}}
\newcommand{\mz}[1]{\textcolor{blue}{MZ: #1}}

\begin{document}

\title{Revision of manuscript JCOMP-D-18-00770  ``A fast spectral method for the inelastic Boltzmann collision operator and application to heated granular gases''}
\author{Jingwei Hu\footnote{Department of Mathematics, Purdue University, West Lafayette, IN 47907, USA (jingweihu@purdue.edu).}, \  \  
	    Zheng Ma\footnote{Department of Mathematics, Purdue University, West Lafayette, IN 47907, USA (ma531@purdue.edu).}
           }      
	\maketitle

\noindent{Dear Editor,}\\

In addition to the Response to Reviewers, we think it is beneficial to address specifically the questions that you raised earlier in the revision stage. We would be happy to provide additional details if needed. Thank you for your time and consideration.

\bigskip
Let us begin by stating that we believe we already gave a fair description of our contribution and the related work [16] in the original version of the paper (there was no mistake or inaccurate statement). In the revised version, we have expanded a bit by providing more details which hopefully will give readers a better idea. Below we further elaborate on these.

\bigskip
1 ``{\it How does this contribution compare with [16]? Please describe the assumptions and applicability of the algorithms in [16] accurately.}"

First of all, the two methods are based on different representations of the collision operator. Our method is based on the so-called $\sigma$-representation, while the method in [16] is based on the Carleman representation. Simply speaking, these are two different coordinate systems to represent the collision integral which upon truncation of the domain and discretization (the variables discretized are of course different in two methods) would result in different approximations of the original collision operator.

Second, the Carleman representation as used in [16] follows an earlier paper ([17] in the revised version) and only readily works for 2D Maxwell molecules and 3D hard spheres as in these two cases the collision kernel is in a separable form so that the resulting sum is a convolution and FFT can be applied. To treat other kernels, [16] modifies the kernel directly to a decoupled form by introducing additional parameters and then determines the values by parameter fitting. The error introduced in this procedure is not clear. The key advantage of our method is that, by using a different representation/discretization, the method requires no assumption on the collision kernel and applies directly to the general form.

Third, we propose to use the {\it spherical design} which is the optimal quadrature on the sphere up to date. In contrast, [16] used the standard tensor product based Gauss quadrature whose accuracy and convergence properties are not comparable to the former.

\bigskip
2 ``{\it What is the novelty of this paper, in regards to [16] and related work? The claim that the algorithm reduces the cost from $O(N^6)$ is misleading since in fact [16] has a computational cost comparable to the proposed algorithm. This should be correctly reflected in the abstract and introduction.}

The novelty of our method has been explained in the previous question. We did point out in the Introduction of the paper that the method [16] has the same order of complexity as ours. However, it should be noted that our method is based on the $\sigma$-representation of the collision operator. This is the one used in the original, direct spectral method [11]. Our fast strategy allows us to accelerate the direct method significantly. Therefore, we can do a clear comparison of the direct and fast methods in terms of the accuracy and efficiency (see Table 4 and Figure 2). The method in [16] does not have a direct, accurate version to compare with. Hence, it is hard to do a careful validation. In fact, in [16], only the macroscopic quantities such as the temperature was quantitatively checked (convergence and accuracy test) and was only done for the 2D Maxwell molecules.

\bigskip
3 ``{\it Please address Reviewer 2's ``Critique" section in details.}

This part is included in our Response to Reviewers.

\end{document}

### 2018-09-14 paper/paper_rev/fast_spec_inelastic_rev.tex
Revised version submitted.

--- 2018-07-08 paper/paper_submit_ver/fast_spec_inelastic.tex
@@
 As a result, the final computational cost is comparable to our method.
-However, we point out that the method in \cite{WZR15} only readily works for 2D Maxwell molecules and 3D hard spheres.
+However, we point out that the method in \cite{WZR15} only readily works for 2D Maxwell molecules and 3D hard spheres \textcolor{red}{as in these two cases the collision kernel is in a separable form so that the resulting sum is a convolution and the fast Fourier transform can be applied (see \cite{MP06} where the Carleman form was first used in the numerical approximation of the Boltzmann collision operator)}.
 To deal with other types of interactions (even 3D Maxwell molecules, say), additional modification of the kernel or parameter fitting is necessary, as was done in \cite{WWSRZ13}.
@@
 \end{equation}
-is a positive constant regardless of $\hat{g}$ and, without loss of generality, is assumed to be $1$. Then
+is a positive constant regardless of $\hat{g}$ and, without loss of generality, is assumed to be~$1$. Then
 \begin{align} 
@@
 \end{equation}
-It is then assumed that $f$ has a compact support: $\text{Supp}_v(f) \approx B_S$, where $B_S$ is a ball centered at the origin with radius $S$ (in practice, the support can be chosen as some multiple of the initial thermal speed $v_{\text{th}}=\sqrt{T_0}$).
+It is then assumed that $f$ has a compact support: $\text{Supp}_v(f) \approx B_S$, where $B_S$ is a ball centered at the origin with radius $S$ (in practice, $S$ can be chosen roughly as  $\max(u_0\pm c\sqrt{T_0})$, where $c$ is some constant around 3).
 Hence it suffices to truncate the infinite integral in $g$ to a larger ball $B_R$ with radius $R=2S$:
@@
 Note that $k=(k_1,\dots,k_d)$ is a multidimensional index, and $\sum_{k=-N/2}^{N/2-1}:=\sum_{k_1,\dots,k_d=-N/2}^{N/2-1}$.
-The choice of $L$ can be made similarly as in the classical case $L\geq (3+\sqrt{2})S/2$, see \cite{PR00} for more details.
+The choice of $L$ can be made similarly as in the classical case $L\geq (3+\sqrt{2})S/2$, see \cite{PR00, FPT05} for more details.
 Now substituting $f_N$ into (\ref{weak}) and choosing $\varphi(v)=e^{-i\frac{\pi}{L}k\cdot v}$, we can obtain the $k$-th mode of the collision operator as
@@
 Surprisingly, the accuracy of the ``separate" method is much better than the ``full" method when $N$ gets bigger.
-Although it seems more natural to approximate the gain and loss terms together using the weak formulation, these results suggest that one should use the more accurate formula for the loss term when it is available (intuitively, the smoothing property of the gain term \cite{Lions94} alleviates the error introduced by the quadrature; since the loss term does not have this property, the same quadrature can incur more errors, hence a more accurate formula should be used).
+Although it seems more natural to approximate the gain and loss terms together using the weak formulation, these results suggest that one should use the more accurate formula for the loss term when it is available.
+Intuitively, the smoothing property of the gain term \textcolor{red}{(see \cite{Lions94} for the classical collision operator and \cite{MM06} for the inelastic case)} alleviates the error introduced by the quadrature; since the loss term does not have this property, the same quadrature can incur more errors, hence a more accurate formula should be used.
 
@@
 
-Numerical examples in this subsection are generated using the ``separate" method (the loss term is computed as in (\ref{separate3}).
+Numerical examples in this subsection are generated using the ``separate" method (the loss term is computed as in equation (\ref{separate3})).
 
-\subsubsection{3D inelastic Maxwell molecules}
+\subsubsection{3D inelastic Maxwell molecules \color{red} --- isotropic solution}
 
@@
   \end{tabu}
-  \caption{Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule for $e=0.2$.
+  \caption{Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule {\color{red} (isotropic solution)} for $e=0.2$.
 Errors shown here are $|T_\text{num} - T_\text{ref}|$ at $t_\text{final}=1$.
@@
   \end{tabu}
-  \caption{Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule for $e=0.2$.
+  \caption{Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule {\color{red} (isotropic solution)} for $e=0.2$.
 Errors shown here are $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$ at $t_\text{final}=1$.
@@
   \end{tabu}
-  	\caption{Average running time per evaluation of the collision operator. Comparison between the direct method and the fast method for various $N$ and fixed $N_{\rho}=30$, $M_{\text{sph}}=32$.}
+  	\caption{Average running time per evaluation of the collision operator.
+Comparison between the direct method and the fast method for various $N$ and fixed $N_{\rho}=30$, $M_{\text{sph}}=32$.
+{\color{blue} This test is performed using Python (numpy and scipy modules for array operations; pyfftw module (a Python wrapper of FFTW) for computing FFTs) on a $10$ CPUs workstation (Intel(R) Xeon(R) CPU E5-2660 v3 @ 2.60GHz, $8$ threads per FFTW operation).}}
 	\label{table 5}
@@
   \includegraphics[width = .8\linewidth]{CPU_time}
-  \caption{Same numbers as in Table~\ref{table 5} plotted in a log-log scale.}
+  \caption{Same numbers as in Table~\ref{table 5} plotted in a log-log scale. }
   \label{CPUtime}
 \end{figure}
+
+{\color{red} \subsubsection{3D inelastic Maxwell molecules  --- anisotropic solution}
+
+In previous subsection we performed a convergence test in $M_\text{sph}$ (compared with the direct method and the analytical solution of the temperature) by considering an isotropic solution.
+This test shows that by using spherical design one can achieve very high accuracy to approximate the integral on the unit sphere, which is a key step in our 3D algorithm.
+In order to further investigate the performance of spherical design and to illustrate its necessity in the algorithm, an example with anisotropic solution is presented in this subsection.
+
+For the sake of comparing with isotropic case, this test is done using exactly the same configuration as in previous subsection except the initial condition:
+\begin{equation}\label{3D_anisotropic}
+  f_0(v) = \frac{4}{5\pi^{\frac{3}{2}}}\left(\exp(-\sqrt[3]{16}|v-2|^2) + \exp(-|v+0.5|^2)\right).
+\end{equation}
+The results are given in Table~\ref{3D_T_conv_anisotropic} and Table~\ref{3D_direct_conv_anisotropic}.
+The error decays with increasing $M_{\text{sph}}$ as expected.
+However, it is also evident that the accuracy is not as good as in the isotropic case.
+Despite this result, we consider it as acceptable since the problem itself (numerical approximation of an integral on 3D unit sphere) is highly nontrivial and is an active research field.
+We believe spherical design is so far among the best methods to deal with such problems.} \textcolor{blue}{We illustrate this statement by performing another test using the standard tensor product based Gauss quadrature, which is also the method used in~\cite{WZR15} for the computation of the integral on 3D unit sphere.
+The results are shown in Table~\ref{3D_quadrature_conv_anisotropic}.
+Under roughly the same total number of quadrature points, this method has very bad performance: slow convergence and low accuracy.}
+
+\begin{table}[htpb]
+  \color{red}
+  \centering
+  \begin{tabu} to 0.8\linewidth {X[1,c] X[2,c]}
+    \toprule
+    $M_{\text{sph}}$ & $|T_\text{num} - T_\text{ref}|$\\
+    \midrule
+    6 & 2.661413377541729e-01 \\
+    12 & 1.6773023124809094e-01 \\
+    32 & 1.0297536867238799e-01 \\
+    48 & 2.584202650005074e-02 \\
+    70 & 2.422376071694199e-02 \\
+    94 & 3.058799838901116e-03 \\
+    120 & 4.413127388460447e-03 \\
+    156 & 1.220110856786194e-03 \\
+    192 & 3.96517639616345e-04 \\
+    \bottomrule
+  \end{tabu}
+  \captionsetup{labelfont={color=red}}
+  \caption{\color{red} Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule (anisotropic solution) for $e=0.2$.
+Errors shown here are $|T_\text{num} - T_\text{ref}|$ at $t_\text{final}=1$.
+$\Delta t = 0.01$, $N = 32$, $N_{\rho}=30$, $R=7$, $L=7.72$.}
+  \label{3D_T_conv_anisotropic}
+\end{table}
+
+\begin{table}[htpb]
+  \color{red}
+  \centering
+  \begin{tabu} to 0.8\linewidth {X[1,c] X[2,c]}
+    \toprule
+    $M_{\text{sph}}$ & $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$\\
+    \midrule
+ 	  6 & 1.4674566108660258e-02 \\
+    12 & 9.532325835757156e-03 \\
+    32 & 5.92707599850977e-03 \\
+    48 & 1.952601851310574e-03 \\
+    70 & 1.43696494086e-03 \\
+    94 & 4.354500695785647e-04 \\
+    120 & 3.534377201557734e-04 \\
+    156 & 9.811152052724729e-05 \\
+    192 & 3.747408101774895e-05 \\ 
+    \bottomrule
+  \end{tabu}
+  \captionsetup{labelfont={color=red}}
+  \caption{\color{red} Convergence test in $M_{\text{sph}}$ of the 3D Maxwell molecule (anisotropic solution) for $e=0.2$.
+Errors shown here are $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$ at $t_\text{final}=1$.
+$\Delta t = 0.01$, $N = 32$, $N_{\rho}=30$, $R=7$, $L=7.72$.}
+  \label{3D_direct_conv_anisotropic}
+\end{table}
+
+\begin{table}[htpb]
+  \color{blue}
+  \centering
+  \begin{tabu} to 0.8\linewidth {X[1,c] X[2,c]}
+    \toprule
+    \# of $(\theta, \phi)$ & $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$\\
+    \midrule
+    (2,3) & 9.301850859331023e-03 \\
+    (3,4) & 1.4164542178285952e-02 \\
+    (4,8) & 2.148725985732372e-02 \\
+    (6,8) & 3.952940827407407e-03 \\
+    (7,10) & 3.3540095954422665e-03 \\
+    (10,10) & 3.405850043869026e-03 \\
+    (10,12) & 3.016912442998702e-03 \\
+    (11,15) & 1.227818352872178e-03 \\
+    (12,16) & 1.2310868626689182e-03 \\
+    \bottomrule
+  \end{tabu}
+  \captionsetup{labelfont={color=blue}}
+  \caption{\color{blue} Convergence test of Gauss quadrature of the 3D Maxwell molecule (anisotropic solution) for $e=0.2$.
+Errors shown here are $|Q_\text{num} - Q_\text{direct}|_{L^{\infty}}$ at $t_\text{final}=1$.
+$\Delta t = 0.01$, $N = 32$, $N_{\rho}=30$, $R=7$, $L=7.72$.}
+  \label{3D_quadrature_conv_anisotropic}
+\end{table}
 
@@
 Note the algebraic decay rather than the exponential decay of Maxwell molecules (cf.
-(\ref{soln:T})).
+equation (\ref{soln:T})).
 
-We take the same initial data \eqref{3Dbkw} as in the previous example but for hard sphere molecules, that is, consider the collision kernel $B_{\sigma} = \frac{1}{4\pi}|g|$.
+{\color{red} We perform two simulations regarding different initial data: the first one we take the same initial data as in \eqref{3Dbkw}; the second one we consider the following initial data which is a Maxwellian with nonzero bulk velocity:
+\begin{equation}\label{3D_maxwellian}
+  f_0(v) = \frac{\rho_0}{(2\pi T_0)^{3/2}}e^{-(v - u_0)^2},
+\end{equation}
+where $\rho_0 = 1$, $T_0 = 2$ and $u_0 = (0.5, -0.5, 0)^T$.
+From~\eqref{conservation} we know $\rho$ and $u$ remain constant in time.
+Both simulations are performed} for hard sphere molecules, that is, for the collision kernel $B_{\sigma} = \frac{1}{4\pi}|g|$.
 Also we set $\varepsilon=0$ in equation (\ref{inBoltz}) (the Haff's law is for unheated case, i.e., the homogeneous cooling state).
 We numerically evaluate the temperature up to time $t_\text{final}=3$ with two different values of restitution coefficient $e=0.2$ and $0.8$.
-The time evolution of $T$ is shown in Figure~\ref{Haff_cooling}, from which one can see that the numerical temperature decays as
+The time evolution of $T$ is shown in Figure~\ref{Haff_cooling} and Figure~\ref{Haff_cooling_maxwellian}, from which one can see that the numerical temperature decays as
 \begin{equation}
-T(t) = \frac{T_0}{(1+C_0 t)^2},
+T(t) = \frac{T_0}{(1+\gamma_0 t)^2},
 \end{equation}
-where the constant $C_0$ depends on the value of $e$. This is in perfect agreement of the Haff's formula (see Section 4.1 of \cite{NE}).
+where the constant $\gamma_0$ {\color{red} depends on the value of $e$ and is estimated by the least squares fitting}.
+This is in perfect agreement with the Haff's formula (see Section 4.1 of \cite{NE}).
 
@@
   \includegraphics[width = .8\linewidth]{Haff's_cooling}
-  \caption{Haff's cooling law. $\Delta t=0.01$, $N=32$, $N_{\rho}=30$, $M_{\text{sph}}=32$, $R=7$, $L=7.72$.}
+  \caption{Haff's cooling law {\color{red} with~\eqref{3Dbkw} as initial data}. $\Delta t=0.01$, $N=32$, $N_{\rho}=30$, $M_{\text{sph}}=32$, $R=7$, $L=7.72$.}
   \label{Haff_cooling}
+\end{figure}
+
+\begin{figure}[htp!]
+  \centering
+  \includegraphics[width = .8\linewidth]{Haff_cooling_maxwellian}
+  \captionsetup{labelfont={color=red}}
+  \caption{\color{red} Haff's cooling law {\color{red} with~\eqref{3D_maxwellian} as initial data}. $\Delta t=0.01$, $N=32$, $N_{\rho}=30$, $M_{\text{sph}}=32$, $R=7$, $L=7.72$.}
+  \label{Haff_cooling_maxwellian}
 \end{figure}
